---
title: "Showing data work (for supervisory purposes)"
format: html
editor: visual
---

Importing packages

```{r echo=FALSE, include=FALSE}
library(tidyverse)
library(systemfit)
library(ggplot2)
library(purrr)
library(knitr)
library(lmtest)
library(sandwich)
library(car)

# functions
source("functions.R")

# actlines and directories
source("UKTUS_params.R")

# matrices defined
source("restriction_matrices.R")

# objects
source("datacleaning_UKTUS_2015.R")

# setwd
setwd(wd)
```

# The (baseline) framework

## Brief summary of setup in Arduini (2025):

-   An individual $i$ belongs to a household $h$. Each individual has a person type $t$. Households are in the same category $g$ if they have the same number of household members of each type. For instance, one household category would be one that contains a wife $t=1$, a husband $t=2$, and children $t=3$.
-   Individual-level characteristics: $\pi_{i,h}$
-   Household-level characteristics: $\zeta_{i,h}$
-   The household purchases private material goods $c_{i,h}$ and public material goods $X$. Private good $j \epsilon \Omega^c$ has price $p_j$. Public good $j \epsilon \Omega^X$ has price $r_j$.
-   Each individual has time-endowment normalised to 1.
-   Four types of time-use: private leisure $l_{i,h}$, joint leisure $jt_{i,h}$, market work $m_{i,h}$ (hours worked, does not enter utility function) and domestic work $d_{i,h}$
-   Joint leisure and domestic work are inputs to public goods $D = f^D(\mathbf{d})$ and $JT = f^{JT}(\mathbf{jt})$ enjoyed by the household Private leisure $l_{i,h}$ includes leisure activities enjoyed by an individual without the co-presence of other household members. It is therefore a private good, which enters the utility function directly.
-   $Q_h$ for the vector of public goods including both material and time-use public goods ($X_h$, $JT_h$, $D_h$) and $R_h$ for the associated price vector. $R_h$ includes wages $w_{i,h}$ of each of its members.
-   Each individual has labour income endowment equal to $w_{i,h}$. An individual’s overall endowment is $y_{i,h} = y_{i,h}^{NL} + w_{i,h}$.
-   Each member can be endowed with nonlabour income $y_{i,h}^{NL}$. The household as a whole can have non-labour income $y_{h}^{NL}$.
-   The household’s endowment (or household budget) is $y_h = \sum_{i \epsilon h} y_{i,h}$.
-   Each individual i of type t living in a household category g has utility function $u_{t,g}$. There is preference heterogeneity across types and household compositions. The form of the bargaining solution of each household depends on the number and types of members, so that the sharing rule for different household categories must be estimated separately.
-   Household individuals bargain over how to divide resources. Relative bargaining power will in general depend both on (i) market variables such as prices (including wages), and (ii) distribution factors, which enter the household’s optimisation problem only indirectly through the distribution of bargaining power $\pi_{i,h}, i \epsilon h$ and $\zeta_{i,h}$. The vector of variables that affect bargaining power is $z_h$.
-   Distribution factors are variables that affect the allocation of resources within households but do not affect the tastes of individuals

## Browning, Chiappori and Weiss (2014):

-   The problem solved by any collective household, regardless of the underlying bargaining process and outside options, can be represented as an optimisation problem where the maximand is the weighted sum of the members’ utility functions. Each member’s utility function is weighed by their Pareto weight $µ_{i,h} (zh)$.
-   Maximise $\sum_{i \epsilon h} (\mu_{i,h}(z_h) u_{t,g}(c_{i,h}, X_h, l_{i,h}, JT_h, D_h))$. Each person's utility depends on individual consumption of private material goods, all public goods, private leisure, joint leisure, and domestic work.

## Arduini (2025): household's problem

-   Stage 1: The household chooses expenditure on market and time-produced (converted to wage units) public goods $R_{h}Q_{h} = \sum_{j \epsilon \Omega^X} r^j X_{h}^j + \sum_{i \epsilon h} w_{i,h} (d_{i,h} + jt_{i,h})$, where $Q_h = (X_h, JT_h, D_h)$, and how to divide the remaining household budget into individual budgets $\rho_{i,h} = (y_h − R_{h}Q_{h}) \eta_{i,h}$. The individual achieves indirect utility $v_{t,g}$ under $\rho_{i,h}$.

$$\text{max}_{\rho,X,d,jt} \sum_{i \epsilon h} \mu_{i,h}(z_h) v_{t,g}(\rho_{i,h}, X_h, JT_h, D_h), \newline
\text{subject to} \newline
y_h = \sum_{i \epsilon h} w_{i,h} (d_{i,h} + jt_{i,h}) + \sum_{j \epsilon \Omega^X} r^j X_{h}^j + \sum_{i \epsilon h} \rho_{i,h}, \newline
d_{i,h} + jt_{i,h} \leq 1, \newline
X_{h}^j, \rho_{i,h}, d_{i,h}, jt_{i,h} \leq 0, \newline
D = f^D(\mathbf{d}),
JT = f^{JT}(\mathbf{jt})$$

-   Sharing rule: determines the share of household budget net of public good expenditure assigned to each member (the individual’s resource share $\eta_{i,h}$).
-   Stage 2: Members decide how to allocate their individual budgets $\rho_{i,h}$ to private expenditure. That is, $$\text{max}_{c_{i,h},m_{i,h},l_{i,h}} u_{t,g}(c_{i,h}, X_h, l_{i,h}, JT_h, D_h), \newline
    \text{subject to} \newline
    \rho_{i,h} = \sum_{j \epsilon \Omega^c} p^j c_{i,h}^j + w_{i,h} l_{i,h}, \newline
    l_{i,h} + m_{i,h} = 1 - (d_{i,h} + jt_{i,h}), \newline
    c_{i,h},m_{i,h},l_{i,h} \leq 0$$
-   In general, second-stage demand depends on the prices of all private goods, including leisure (but not directly on the prices of public goods). Unless preferences are separable in private and public goods, it depends on public good consumption as determined in the first stage. Finally, it depends on the individual’s second-stage budget, which is their resource share multiplied by the household budget net of public good expenditure (determined in the first stage) $\rho_{i,h} = (y_h − R_{h}Q_{h}) \eta_{i,h}$.

## Arduini (2025): specification and identification

### The private assignable good

-   The private assignable good: $c^a$ is private and assignable if we observe individual-level demand or expenditure for it. That is, it is consumed by a specific person and we can clearly attribute demand or expenditure to that individual.
-   Individual i’s (of type t) second-stage demand for the assignable good is given by $c_{i,h}^a(p, w_{i,h}, Q_h, (y_h − R_{h}Q_{h}) \eta_{i,h})$

### Other assignable goods:

-   Clothing and non-market work
-   Calvi et al. (2022) find that estimates of women’s resource shares (from material assignable goods) are closely related to those women’s satisfaction with their availability of leisure time, suggesting that the collective model is right to conceptualise the same sharing rule being applied both to time-use and material goods.

### Cobb-Douglas preferences

-   What she uses: Cobb-Douglas preferences, leisure as the assignable good, an identifying assumption restricting preference heterogeneity, and (iv) an additional restriction on preference heterogeneity allowing estimation of the sharing rule from time-use data alone.
-   What she needs: The variation in assignable good expenditure with household budget is taken to be informative about how household budget is split between individual members’ second stage budgets. This relies on the strong assumption that men and women have the same preference parameter for the assignable good, or else that we already know the preference parameters of men and women, e.g. from singles data.
-   When using private leisure as the assignable good, this should not be interpreted as assuming that expenditure on any specific material good is split between members in the same manner as leisure. Men and women have a similar preference parameter for private leisure specifically. This means that, if they have the same wages and the same individual budget, then they will have the same hours of private leisure.

$$u_{t,g} = \sum_{j \epsilon \Omega^c} (\alpha_{t,g}^{cj} \ln{(c_{i,h}^j)}) + \sum_{j \epsilon \Omega^X} (\alpha_{t,g}^{Xj} \ln{(X_{h}^j)}) + \sum_{st \epsilon h} (\alpha_{t,g}^{Tst,g} \ln{(jt_{s,h})}) + \sum_{st \epsilon h} (\alpha_{t,g}^{Dst,g} \ln{(d_{s,h})})$$ - Resource shares are equal to Pareto weights scaled by a composite preference parameter. For instance, a significant effect of wages on the sharing rule implies that Pareto weights vary with wages

### Private leisure C-D SRAT:

-   Private leisure C-D SAP: The ratio between men and women’s preference parameter for the private assignable good is stable across household compositions. Cobb-Douglas preferences over the private assignable good: homotheticity of private leisure. Everyone, regardless of their type (in a given household category) is assumed to have the same Cobb-Douglas preference parameter for private leisure: $\alpha_{t,g}^l = \alpha_{g}^l$. Preference heterogeneity across types is maintained for all other preference parameters, and across household compositions. Men and women spend the same share of their individual second-stage budget on leisure, but will split the remaining budget differently between other private goods.
-   Private leisure C-D SRAT: Instead of requiring the level of preference parameters for the assignable good to be stable across household composition, she requires only that the ratio between men and women’s preference parameter for the assignable good is stable across household compositions: $\frac{\alpha_{t,g}^l}{\alpha_{st,g}^l} = \frac{\alpha_{t}^l}{\alpha_{st}^l}$
-   In particular, Cobb-Douglas preferences impose homotheticity, which may seem to strong an assumption for certain goods. It is important to note that I do not require the whole utility function to be Cobb-Douglas: more specifically, I require preferences that are Cobb-Douglas over the assignable good and some aggregate good. Therefore, I am imposing homotheticity on the assignable good, but not on other specific goods.

### Public good C-D SAP:

-   Public Goods SAP: Using family data probably violates assumption that total expenditure on public goods does not vary with the distribution of bargaining power in the household. This assumption does not amount to a unitary model in that the distribution of bargaining power still affects the composition of public good (and private good) expenditure; it simply does not affect the overall split between private and public expenditure. For example, female empowerment could lead to more spending on housing and less on television subscriptions.
-   With Cobb-Douglas preferences, we can achieve the desired restriction with public good C-D SAP. Public good C-D SAP imposes that the sum of all public good preference parameters is the same for all individuals of the same household category g:

$$\sum_{j \epsilon \Omega^X} \alpha_{t,g}^{j} + \sum_{st \epsilon h} \alpha_{t,g}^{Tst,g} + \sum_{st \epsilon h} \alpha_{t,g}^{Dst,g} = a_{g}^Q$$ - As a result of the public good SAP assumption, it is possible to estimate the sharing rule from time-use data alone, if using private leisure as the assignable good.

## Arduini (2025): estimating

-   Resource shares will generally depend on household budget, prices (including wages of all household members) and Pareto weights of household members: $\eta_{i,h}(y_h, p, r, w_h, µ_h)$.
-   Recall that Pareto weights µh are an unknown function of a vector of variables zh that determine relative bargaining power in the household, and that the collective model does not pin down a specific bargaining solution. Since zh in general includes the other variables which affect $\eta_{i,h}$, we can write $\eta_{i,h}(z_h)$
-   To estimate resource shares by linear regression, we linearly approximate them: $$\eta_{i,h} = \eta_{i,h}^0 + \sum_{z} \eta_{t,g}^{z} (z_h - \bar{z})$$
-   $\eta_{i,h}^0$ is the average resource share of a type (living in a specific household category). By definition, $\sum_{t \epsilon h} \eta_{t,g}^0 = 1$. For instance, a higher-than-average wage for the woman might increase the woman’s resource share, so that she would have a higher-than-average-for-women resource share.
-   $\eta_{t,g}^{z}$ captures the impact on sharing of a household’s characteristic $z_h$ deviating from the sample average $(z_h - \bar{z})$
-   An equation for each type of person in the form: $$w_{i,h}l_{i,h} = \alpha_{t,g}^{l} y_h (\eta_{t,g}^0 + \sum_{z} \eta_{t,g}^{z}(z_h - \bar{z})) + \epsilon_{i,h}$$
-   If $z_h$ includes all key characteristics that affect bargaining (or at least all those that are correlated with regressors) then estimates should be unbiased. The errors are negatively correlated within household (with a correlation weaker in magnitude than -1), so that a SURE estimation approach is recommended.
-   Equation for women: $$w_{i,h}l_{i,h} = \beta_{f,g}^0 y_h + \sum_{z} \beta_{f,g}^{z} y_h (z_h - \bar{z})) + \epsilon_{i,h}$$
-   Equation for men: $$w_{i,h}l_{i,h} = \beta_{m,g}^0 y_h + \sum_{z} \beta_{m,g}^{z} y_h (z_h - \bar{z})) + \epsilon_{i,h}$$
-   Then, $\beta_{t}^0 = \alpha_{t,g}^{l} \eta_{t,g}^0$ and $\beta_{t}^z = \alpha_{t,g}^{l} \eta_{t,g}^z$.
-   First, linear SURE regressions of leisure expenditure, one for men and one for women, with the restriction that $β_{f}^z + β_{m}^z = 0$. This is because $\beta_{t}^z = \alpha_{t}^{l} \eta_{t}^z$ and from private leisure C-D SAP and the definition of the linear approximation parameters, $\sum_{t \epsilon h} \beta_{t}^z = \alpha_{t}^{l} \sum_{t \epsilon h} \eta_{t}^z = 0$.
-   Next, estimate each type’s average resource share as $\hat{\eta}_{t}^0 = \frac{\hat{\beta_{t}^0}}{\hat{\beta_{f}^0} + \hat{\beta_{m}^0}}$
-   Next, estimate $\hat{\alpha_{t}^{l}} = \frac{\hat{\beta_{t}^0}}{\hat{\eta}_{t}^0}$, since $\beta_{t}^0 = \alpha_{t}^{l} \eta_{t}^0$.
-   Next, estimate $\hat{\eta_{t}^{z}} = \frac{\hat{\beta_{t}^z}}{\hat{\alpha_{t}^{l}}}$, since $\beta_{t}^z = \alpha_{t}^{l} \eta_{t}^z$.
-   Armed with the sharing rule, one can estimate individual-specific resource shares in the dataset used for estimation (the time-use dataset in this case): $\hat{\eta_{i,h}} = \hat{\eta}_{t}^0 + \sum_{z} \hat{\eta_{t}^{z}} (z_h - \bar{z})$.

### Issues

-   The identification problem we face is that the man’s expenditure on private leisure may be more or less responsive than the woman’s to changes in the household budget for one of two reasons, or a combination of them. The first possibility is that the man receives a larger proportion of the household budget (net of public good expenditure), i.e. he has a higher resource share than the woman. The second possibility is that the man’s preference for private leisure is stronger relative to the woman’s, so that he spends a larger proportion of his individual budget on private leisure. In order to disentangle these two channels, she shuts down the preference channel with the identifying assumption to identify the sharing rule.
-   Under CD-SAP, private leisure expenditure for individuals of any type (within a given household category) $l_i w_i = α^l y_h η_{i,h}$ responds in the same way to the same increase in individual budget $y_h η_{i,h}$, so differences in responsiveness to changes in $y_h$ identify differences in sharing.
-   All changes in assignable-good expenditure induced by distribution factors operate through the sharing rule.
-   Private leisure C-D SRAT rules out distribution factors (wages, education, demographics, etc.) changing relative tastes for leisure between men and women, and household composition or bargaining environment altering how much men vs women intrinsically value leisure.
-   Public goods C-D SAP says individuals may differ in how much they care about the public good, but the total weight placed on public goods is fixed for a given household type (e.g. heterosexual couples with children).

# ONLY ONE HOUSEHOLD MEMBER COMPLETES THE TIME DIARY IN THE ATUS!! I CANNOT DIRECTLY STUDY WITHIN-HOUSEHOLD SUBSTITUTION

-   I can’t observe whether a female respondent increase in wage causes the husband's childcare to fall. I don't observe the spouse’s full time allocation or the spouse’s private leisure directly
-   In multi-diary data, I would estimate the system by SURE because the unobservables entering the two assignable-good equations share household shocks and are mechanically linked by budget splitting, inducing cross-equation correlation.
-   In ATUS, only one spouse reports a diary, so the two equations are estimated on disjoint household samples and cross-equation covariance is not identified. In the paper, the estimating system is (conceptually) two equations per household, a male assignable-good equation (male leisure expenditure) and a female assignable-good equation (female leisure expenditure) for the same household
-   ATUS does invalidate the specific identification strategy in the paper (the one that uses both spouses’ assignable-good demands in the same household plus within-household adding-up restrictions and SUR correlation).
-   With the ATUS, I do not observe the spouse’s diary outcomes. Rather, I get two different sets of households. This does not invalidate the idea of predicting shares themselves, but what I get is not the same object as the paper’s within-household resource shares.

# The public good of childcare, children as public goods themselves

# Finding the sharing rule

-   State and time dummies: unobserved heterogeneity that we address relates to the individual-specific intercept terms and the common time-varying component of the error term

### No restrictions on coefficients: deviations from average with own sex
```{r}
# leisure including eating, drinking, washing, spirituality, volunteering
fit_unres_within = systemfit(eqns_within, method = "SUR", data = sharing_est_data_2015)
# relaxing, socialising, sport, entertainment, and hobbies only
fit_unres_r_within = systemfit(eqns_r_within, method = "SUR", data = sharing_est_data_2015)
```

### No restrictions on coefficients: deviations from average of men AND women
```{r}
# leisure including eating, drinking, washing, spirituality, volunteering
fit_unres_between = systemfit(eqns_between, method = "SUR", data = sharing_est_data_2015)
# relaxing, socialising, sport, entertainment, and hobbies only
fit_unres_r_between = systemfit(eqns_r_between, method = "SUR", data = sharing_est_data_2015)
```

### No restrictions on coefficients: deviations from average of opposite sex
```{r}
# leisure including eating, drinking, washing, spirituality, volunteering
fit_unres_opposite = systemfit(eqns_opposite, method = "SUR", data = sharing_est_data_2015)
# relaxing, socialising, sport, entertainment, and hobbies only
fit_unres_r_opposite = systemfit(eqns_opposite, method = "SUR", data = sharing_est_data_2015)
```

### Restricted coefficients, deviations from average with own sex
```{r}
# leisure including eating, drinking, washing, spirituality, volunteering
fit_res_within = systemfit(eqns_within, method = "SUR", data = sharing_est_data_2015,
                           restrict.regMat = modReg_within)
# relaxing, socialising, sport, entertainment, and hobbies only
fit_res_r_within = systemfit(eqns_r_within, method = "SUR", data = sharing_est_data_2015, 
                             restrict.regMat = modReg_within)
```

### Restricted coefficients, deviations from average of men AND women
```{r}
# leisure including eating, drinking, washing, spirituality, volunteering
fit_res_between = systemfit(eqns_between, method = "SUR", data = sharing_est_data_2015,
                            restrict.regMat = modReg_between)
# relaxing, socialising, sport, entertainment, and hobbies only
fit_res_r_between = systemfit(eqns_r_between, method = "SUR", data = sharing_est_data_2015,
                              restrict.regMat = modReg_between)
```

### Restricted coefficients, deviations from average of opposite sex
```{r}
# leisure including eating, drinking, washing, spirituality, volunteering
fit_res_opposite = systemfit(eqns_opposite, method = "SUR", data = sharing_est_data_2015,
                             restrict.regMat = modReg_opposite)
# relaxing, socialising, sport, entertainment, and hobbies only
fit_res_r_opposite = systemfit(eqns_r_opposite, method = "SUR", data = sharing_est_data_2015,
                               restrict.regMat = modReg_opposite)
```

### Finding the resource shares

```{r echo=FALSE}
shares_fit_bothsex = add_shares_from_lm(fit_res_between, data=sharing_est_data_2015, dev_type = "all")
shares_fit_bothsex_r = add_shares_from_lm(fit_res_r_between, data=sharing_est_data_2015, dev_type = "all")
shares_fit_ownsex = add_shares_from_lm(fit_res_within, data=sharing_est_data_2015, dev_type = "own")
shares_fit_ownsex_r = add_shares_from_lm(fit_res_r_within, data=sharing_est_data_2015, dev_type = "own")
shares_fit_oppsex = add_shares_from_lm(fit_res_opposite, data=sharing_est_data_2015, dev_type = "opp")
shares_fit_oppsex_r = add_shares_from_lm(fit_res_r_opposite, data=sharing_est_data_2015, dev_type = "opp")
```

# Differences between and within sexes: a closer look at distribution factors

## Some density plots

-   Arduini (2025) uses the following example for deviation: A higher-than-average wage for the woman might increase the woman’s resource share, so that she would have a higher-than-average-for-women resource share.
-   Bargaining power may come from strength relative to one's spouse or strength relative to a spouse's options within one's own gender. Say in 2006 there is a leisure gap when women contribute more to the household. This may result in smaller increases in women’s leisure than if it were men. 
- It's worth seeing the women-vs-women/men-vs-men and women-vs-all/men-vs-all comparison though? Doesn't look very different 
- You can also think about someone's strength relative to their own other options, although I don't really know the interpretation for this

```{r echo=FALSE}
plot_share_densities(shares_fit_bothsex_r, dev_type = "all")
plot_share_densities(shares_fit_ownsex_r, dev_type = "own")
plot_share_densities(shares_fit_oppsex_r, dev_type = "opp")
```
- There's more of a difference when you look at the broader leisure definition though:
```{r echo=FALSE}
plot_share_densities(shares_fit_bothsex, dev_type = "all", restrict_leisure = FALSE)
plot_share_densities(shares_fit_ownsex, dev_type = "own", restrict_leisure = FALSE)
plot_share_densities(shares_fit_oppsex, dev_type = "opp", restrict_leisure = FALSE)
```

## How would households with characteristics $z_h$ be treated by a different sharing rule?

-   Changes over time: take income and matchings from 2004, and take coefficients from 2024. apply the betas to the observed behaviour of 2024. Would I get different counterfactual leisures?
-   For two periods, I estimate $\beta_t^0 = \alpha_t^\ell \eta_t^0,$, $\beta_t^z = \alpha_t^\ell \eta_t^z$, average resource shares $\hat\eta_t^0 = \frac{\hat\beta_t^0}{\hat\beta_m^0 + \hat\beta_f^0}$, the leisure preference parameter $\hat\alpha_t^\ell = \frac{\hat\beta_t^0}{\hat\eta_t^0}$, and each marginal sharing rule $\hat\eta_t^z = \frac{\hat\beta_t^z}{\hat\alpha_t^\ell}$.
-   So for each year (say 2004 and 2024), I'd get: $$\hat\eta_{t,h}^{(2004)} = \hat\eta_t^{0,(2004)} + \sum_z \hat\eta_t^{z,(2004)} (z_h - \bar z), \newline
    \hat\eta_{t,h}^{(2024)} = \hat\eta_t^{0,(2024)} + \sum_z \hat\eta_t^{z,(2024)} (z_h - \bar z)$$
-   Then, why not take households observed in 2004 (their $z_h$), but apply the newer sharing rule: $$\hat\eta_{i,h}^{CF} = \hat\eta_t^{0,(2024)} + \sum_z \hat\eta_t^{z,(2024)} (z_h^{2004} - \bar z^{2024})$$
-   I'd want to compare $\hat\eta_{t,h}^{CF} \quad$ and $\quad \hat\eta_{t,h}^{(2004)}$.
-   If we have $\hat\eta_{f,h}^{CF} > \hat\eta_{f,h}^{(2004)}$ for the observed 2004 household characteristics, the sharing rule has shifted in a way that allocates a larger share of resources to women.
-   When the deviations of a household's characteristics are relative to averages across both males and females, I'd propose that a higher \hat\eta\_t\^{z,(2024)} than \hat\eta\_t\^{z,(2004)} for some $z$ is capturing increased importance of someone’s absolute wage/education/whatever is in determining allocation, more like “my wife’s human capital carries more weight in allocation than it used to” than “my wife's position relative to my other options carries more weight than it used to”.
-   When the deviations of a household's characteristics are relative to averages of only the respective gender associated with the characteristic (for instance, the deviation of the man's wage from the wage of other men), I'd propose that a higher $\hat\eta_t^{z,(2024)}$ than $\hat\eta_t^{z,(2004)}$ for some $z$ is capturing increased importance of someone’s absolute wage/education/whatever is in determining allocation, more like how they compare to the opposing spouse's outside options became more important for intrahousehold allocation. So, it's like the other way around from the previous point.
-   For a counterfactual analysis, I can provide tables with two sets of estimated average resource shares and marginal contributions, each corresponding to the older or newer years, as well as the distribution of estimated counterfactual resource shares in the older data.

### An example

-   Suppose in 2004, the woman's wage relative to all other wages mattered *less* for the sharing rule, i.e. $\hat\eta_t^{w_f,(2004)}$ is small.
-   In 2024, maybe $\hat\eta_t^{w_f,(2024)}$ is larger. This would mean that household allocations are more affected by the woman's wage relative to all other wages. Women’s human capital carries more weight in allocation than it used to
-   Applying the 2024 coefficients to 2004 households: even though 2004 women earned less on average, the same relative wage difference would buy them more bargaining power under the 2024 rule.
-   The identifying assumption is: conditional on $z_h$, the estimated sharing rule captures a stable mapping from characteristics to allocation, which is the same assumption Arduini (2025) relies on.

### 

# To do:

-   data cleaning for the UKTUS! or find something else
-   if women get richer does time spent with child increase or decrease? extra dollar with woman’s wage = more time with baby for the woman and maybe less time for the man? would it be the other way around for the man? if both household members have the same wage increase, does it mean the woman will still be the one taking care of the kid or will there be a shift
-   child gender differences: having boys vs girls (this is plausibly exogenous) and some results in literature find that men tend to want to spend time with sons but not daughters. (check gender ratios of births in states) I think one way to do this would be to use the plausible (at least in an American context) exogeneity of having more sons versus more daughters in the household. Ultimately, I think I want to find the predicted resource shares using the sharing rule and the SURE regressions and then find the effect of resource shares on childcare provision for men and women using son and daughter ratios(?) as a source of plausible exogeneity.
-   if women get richer does time spent with child increase or decrease? extra dollar with woman’s wage = more time with baby for the woman and maybe less time for the man? would it be the other way around for the man? if both household members have the same wage increase, does it mean the woman will still be the one taking care of the kid or will there be a shift

# thought jail slash random archive

My project (roughly): What is the impact of women’s control over their earnings (their bargaining power in the home) on their "child-rearing" supply? Using method as in Arduini to find the sharing rule, say who has more bargaining power based on private leisure, and then see if the person with more power spends more or less time on childcare. Developed from literature on the income effects of household fertility choice and intra-household bargaining, I use the American Time Use Survey (ATUS) to extend existing collective household models estimating intra-household inequality. In line with existing methodology, my identification relies on private assignable goods, goods that cannot be shared between more than one person and are observed to be consumed by only one person in the household. This information is typically found in household expenditure surveys, but such goods are typically durable in the sense that they appear sparingly in expenditure data. To address the resulting power limitations with this approach, I adapt the approach of Arduini (2025) to a setting with parents and children, using private leisure in time use data to estimate bargaining power in the household. I then take this sharing rule to further discussions of childcare provision based on the bargaining power of individual parents. I both estimate and test for heterogeneity of childcare provision. I want to see the differences between male- or female-"dominated" households? Some existing papers use earnings/education differentials as a measure of bargaining power or fertility rates as a way to talk about family planning, but culture and gender roles may play a role in within-household childcare decisions and time spent on childcare is very tractable considering the breadth of time use data available for many countries.

## The literature

Literature on household inequality generally tends to and focuses on working couples *without* children, or models them as public goods. \[Arduini (2024)\]\[https://ifs.org.uk/sites/default/files/2025-05/WP202517-Estimating-intra-household-sharing-from-time-use-data.pdf\] uses the concept of private leisure time (time spent on leisure activities without other members of the household present) to estimate intra-household sharing. She defines the "assignable good", which is a private good. These parts are useful:

*"The main limitation of private leisure is that its price is only known for those who work, for whom it coincides with the observed wage. For non-participants, the price of time is higher than their (unobserved) wage and we would have to estimate that in order to use time-use data to identify sharing. That extension is left for future work: in my application to UK data, I focus on working couples."*

\[Blundell et al (2019)\]\[https://www.sfu.ca/\~pendakur/papers/BKLP16_longtitle.pdf\] still need to read this lol

\[Hyee and Robledo (2010)\]\[https://www.nottingham.ac.uk/economics/documents/discussion-papers/10-06.pdf\]

Calvi (2002): Under some mild assumptions, there exists a monotonic correspondence between Pareto weights and resource shares. Consequently, resource shares may also be interpreted as a measure of intrahousehold bargaining power. In effect, these are more tractable measures of power, as they are invariant to unobservable car- dinalizations of the utility functions.

\[Andrew et al (2021)\]\[https://ifs.org.uk/inequality/wp-content/uploads/2021/03/BN-The-careers-and-time-use-of-mothers-and-fathers.pdf\] show that heteronormative households, mothers are more likely to leave the labour force after having children regardless of the relative wages between mother and father before childbirth. (Much of the recent literature in this area focuses on COVID-19, maybe look into this.) It's difficult to really prove who in a relationship is doing the colloquially-named "mental load" of parenting (stuff like keeping appointments, planning meals, remembering when the laundry was last done), although anecdotally this is usually the wife or mother's job in this kind of household. (There may be a time use paper somewhere that tries to deal with this, maybe look into this). The ATUS records time spent doing activities, but not the thoughts that go on in someone's head. It also records the people present during an activity, but not, as far as I've seen, who is a active participant, or if someone needs to be actively prompted to do the groceries or pick up after a child.

\[Dunbar et al (2013)\]\[https://www.sfu.ca/\~pendakur/CollectiveChildren30-with-appendices.pdf\]: Geoff is looking at parents' and childrens' resource shares using expenditure data, I think I'd want to stick to the inequality of the couple. These parts are useful:

*"In models where goods are purely private or purely public, what we call private, assignable goods are known as exclusive goods. See, e.g. Bourguignon, Browning, and Chiappori (2009). Chiappori and Ekeland (2008) and Cherchye, De Rock and Vermeulen (2008) among others show how assignable goods can aid in the identification of resource shares. Our strategy follows this line in assuming the presence, and observability, of a small number of private assignable goods, and uses these to identify childrens’ resource shares. The end result is that we identify how total household expenditures on all goods are divided up among household members, just by observing how family expenditures on each member’s single private good (like clothing) vary with total expenditures."*

*"We find that children command a reasonably large share of resources – roughly 20 percent for the first child – and that this share rises with the number of children – 5-10 percentage points per additional child. Moreover, fathers command a larger share of resources than mothers, and mothers seem to sacrifice more resources than fathers to their children."*

*"A good is defined to be private if it cannot be shared or consumed jointly by more than one person, and is defined to be assignable if it is consumed by one individual household member that is known to the researcher. Examples could include toys and diapers which are private goods assignable to children, or alcohol and tobacco which are private goods assignable to adults. In our empirical application we use clothing"*

Childcare-related activities from the ATUS documentation:

```         
030101  Physical care for household children
030102  Reading to or with household children
030103  Playing with household children, not sports
030104  Arts and crafts with household children
030105  Playing sports with household children
030106  Talking with or listening to household children
030107  Helping or teaching household children (not related to education) (2003)
030108  Organization and planning for household children
030109  Looking after household children (as a primary activity)
030110  Attending household children's events
030111  Waiting for or with household children
030112  Picking up or dropping off household children
030199  Caring for and helping household children, n.e.c.
030201  Homework (household children)
030202  Meetings and school conferences (household children)
030203  Home schooling of household children
030204  Waiting associated with household children's education
030299  Activities related to household child's education, n.e.c.
030301  Providing medical care to household children
030302  Obtaining medical care for household children
030303  Waiting associated with household children's health
030399  Activities related to household child's health, n.e.c.
080101  Using paid childcare services
080102  Waiting associated with purchasing childcare services
080199  Using paid childcare services, n.e.c.
160103  Telephone calls to or from education services providers
160107  Telephone calls to or from paid child or adult care providers
180301  Travel related to caring for and helping household children, inclusive (2003, 2004)
180302  Travel related to caring for and helping household children (2005+)
180303  Travel related to household children's education (2005+)
180304  Travel related to household children's health (2005+)
180801  Travel related to using childcare services
```

Private leisure activities from the ATUS documentation: needs to be done without the spouse or other household members present This includes all types of self care including sleeping and eating!

```         
General
010101  Sleeping    
010102  Sleeplessness   
010199  Sleeping, n.e.c.
010201  Washing, dressing, and grooming oneself
010299  Grooming, n.e.c.
010301  Health-related self care    
010399  Self care, n.e.c.
010401  Personal or private activities  
010499  Personal activities, n.e.c.
019999  Personal care, n.e.c.
050201  Socializing, relaxing, and leisure as part of job
020603  Walking, exercising, playing with animals (2008+)
120201  Attending or hosting parties, receptions, or ceremonies
120202  Attending meetings for personal interest (not volunteering)
120299  Attending or hosting social events, n.e.c.
120301  Relaxing, thinking
120302  Tobacco and drug use
120303  Television and movies (not religious)
120304  Television (religious)
120305  Listening to the radio
120306  Listening to or playing music (not radio)
120307  Playing games
120308  Computer use for leisure (excluding games)
120309  Arts and crafts as a hobby
120310  Collecting as a hobby
120311  Hobbies, except arts and crafts and collecting
120312  Reading for personal interest
120313  Writing for personal interest
120399  Relaxing and leisure, n.e.c.
120401  Attending performing arts
120402  Attending museums
120403  Attending movies or film
120404  Attending gambling establishments
120405  Security procedures related to arts and entertainment
120499  Arts and entertainment, n.e.c.
120501  Waiting associated with socializing and communicating
120502  Waiting associated with attending or hosting social events
120503  Waiting associated with relaxing or leisure
120504  Waiting associated with arts and entertainment
120599  Waiting associated with socializing, n.e.c.
129999  Socializing, relaxing, and leisure, n.e.c.
060102  Taking class for personal interest
060201  Extracurricular club activities
060202  Extracurricular music and performance activities
060203  Extracurricular student government activities
060204  Waiting associated with extracurricular activities (2004+)
060299  Education-related extracurricular activities, n.e.c.
060302  Research or homework for class (for personal interest)
060402  Administrative activities: class for personal interest
070104  Shopping, except groceries, food, and gas
110101  Eating and drinking
050202  Eating and drinking as part of job
110201  Waiting associated with eating and drinking
110299  Waiting associated with eating and drinking, n.e.c.
119999  Eating and drinking, n.e.c.
120101  Socializing and communicating with others
120199  Socializing and communicating, n.e.c.
160101  Telephone calls to or from family members
160102  Telephone calls to or from friends, neighbors, or acquaintances
080501  Using personal care services
080502  Waiting associated with personal care services
080599  Using personal care services, n.e.c.

Sport
130101  Doing aerobics
130102  Playing baseball
130103  Playing basketball
130104  Biking
130105  Playing billiards
130106  Boating
130107  Bowling
130108  Climbing, spelunking, caving
130109  Dancing
130110  Participating in equestrian sports
130111  Fencing
130112  Fishing
130113  Playing football
130114  Golfing
130115  Doing gymnastics
130116  Hiking
130117  Playing hockey
130118  Hunting
130119  Participating in martial arts
130120  Playing racquet sports
130121  Participating in rodeo competitions
130122  Rollerblading
130123  Playing rugby
130124  Running
130125  Skiing, ice skating, snowboarding
130126  Playing soccer
130127  Playing softball
130128  Using cardiovascular equipment
130129  Vehicle touring or racing
130130  Playing volleyball
130131  Walking
130132  Participating in water sports
130133  Weightlifting or strength training
130134  Working out, unspecified
130135  Wrestling
130136  Doing yoga
130199  Playing sports, n.e.c.
130201  Watching aerobics
130202  Watching baseball
130203  Watching basketball
130204  Watching biking
130205  Watching billiards
130206  Watching boating
130207  Watching bowling
130208  Watching climbing, spelunking, caving
130209  Watching dancing
130210  Watching equestrian sports
130211  Watching fencing
130212  Watching fishing
130213  Watching football
130214  Watching golfing
130215  Watching gymnastics
130216  Watching hockey
130217  Watching martial arts
130218  Watching racquet sports
130219  Watching rodeo competitions
130220  Watching rollerblading
130221  Watching rugby
130222  Watching running
130223  Watching skiing, ice skating, snowboarding
130224  Watching soccer
130225  Watching softball
130226  Watching vehicle touring or racing
130227  Watching volleyball
130228  Watching walking
130229  Watching water sports
130230  Watching weightlifting or strength training
130231  Watching people working out, unspecified
130232  Watching wrestling
130299  Attending sporting events, n.e.c.
130301  Waiting related to playing sports or exercising
130302  Waiting related to attending sporting events
130399  Waiting associated with sports, exercise, and recreation, n.e.c.

Religion
140101  Attending religious services    
140102  Participation in religious practices    
140103  Waiting associated with religious and spiritual activities  
140104  Security procedures related to religious and spiritual activities
140105  Religious education activities (2007+)
149999  Religious and spiritual activities, n.e.c.

Volunteering
150101  Computer use    
150102  Organizing and preparing    
150103  Reading 
150104  Telephone calls (except hotline counseling) 
150105  Writing 
150106  Fundraising 
150199  Administrative and support activities, n.e.c.
150201  Food preparation, presentation, clean-up
150202  Collecting and delivering clothing and other goods
150203  Providing care  
150204  Teaching, leading, counseling, mentoring    
150299  Social service and care activities, n.e.c.
150301  Building houses, wildlife sites, and other structures
150302  Indoor and outdoor maintenance, repair, and clean-up
150399  Indoor and outdoor maintenance, building, and clean-up activities, n.e.c.
150402  Serving at volunteer events and cultural activities
150499  Participating in performance and cultural activities, n.e.c.
150500  Attending Meetings, Conferences, and Training
150501  Attending meetings, conferences, and training   
150599  Attending meetings, conferences, and training, n.e.c.
150601  Public health activities
150602  Public safety activities    
150699  Public health and safety activities, n.e.c. 
150701  Waiting associated with volunteer activities (2004+)
150799  Waiting associated with volunteer activities, n.e.c. (2004+)
150801  Security procedures related to volunteer activities (2007+)
150899  Security proecdures related to voluteer activities, n.e.c. (2007+)
159999  Volunteer activities, n.e.c.

Travel
180805  Travel related to using personal care services
181201  Travel related to socializing and communicating 
181202  Travel related to attending or hosting social events    
181101  Travel related to eating and drinking
181199  Travel related to eating and drinking, n.e.c.
181203  Travel related to relaxing and leisure (2003, 2004)
181204  Travel related to arts and entertainment    
181205  Travel as a form of entertainment (2005+)
181206  Travel related to relaxing and leisure (2005+)
181299  Travel related to socializing, relaxing, and leisure, n.e.c.
181300  Travel Related to Sports, Exercise, and Recreation
181301  Travel related to participating in sports, exercise, or recreation  
181302  Travel related to attending sporting or recreational events
181401  Travel related to religious or spiritual practices
181499  Travel related to religious or spiritual activities, n.e.c.
181501  Travel related to volunteering
181599  Travel related to volunteer activities, n.e.c.
```

People do activities with...

```         
0100    Alone   
0200    Spouse  
0201    Unmarried partner   
0202    Own household child 
0203    Grandchild  
0204    Parent  
0205    Brother sister  
0206    Other related person    
0207    Foster child    
0208    Housemate, roommate 
0209    Roomer, boarder 
0210    Other nonrelative   
0300    Own non-household child under 18    
0400    Parents (not living in household)   
0401    Other non-household family members under 18 
0402    Other non-household family members 18 and older (including parents-in-law)  
0403    Friends 
0404    Co-workers, colleagues, clients (non-work activities only)
0405    Neighbors, acquaintances    
0406    Other non-household children under 18   
0407    Other non-household adults 18 and older 
0408    Boss or manager
0409    People whom I supervise
0410    Co-workers
0411    Customers
9996    Refused
9997    Don't know  
9998    Blank   
```
