---
title: "Showing data work (for supervisory purposes)"
format: html
editor: visual
---

Importing packages
```{r echo=FALSE}
library(tidyverse)
library(systemfit)
library(ggplot2)
library(purrr)
library(knitr)

# functions
source("functions.R")

# matrices defined
source("restriction_matrices.R")
```

Import data

```{r}
individual = read.csv("atus_working_parents_act.csv")
hh = read.csv("atus_working_parents_act_collapse.csv")
```

Table printing:
```{r}
# sharing rule table for coefficients
rows_within = c("y" = "Budget",
                "Bx_dev_wage_f_only" = "Budget * dev. fem. hourly pay",
                "Bx_dev_wage_m_only" = "Budget * dev. mal. hourly pay",
                "Bx_dev_educ_f_only" = "Budget * dev. fem. qualification",
                "Bx_dev_educ_m_only" = "Budget * dev. mal. qualification",
                "Bx_dev_avgage" = "Budget * dev. average age",
                "Bx_dev_agegap" = "Budget * dev. age gap")

rows_between = c("y" = "Budget",
                 "Bx_dev_wage_f_all" = "Budget * dev. all hourly pay",
                 "Bx_dev_wage_m_all" = "Budget * dev. all hourly pay",
                 "Bx_dev_educ_f_all" = "Budget * dev. all qualification",
                 "Bx_dev_educ_m_all" = "Budget * dev. all qualification",
                 "Bx_dev_avgage" = "Budget * dev. average age",
                 "Bx_dev_agegap" = "Budget * dev. age gap")

```

# The framework 

## Brief summary of setup in Arduini (2025):
- An individual $i$ belongs to a household $h$. Each individual has a person type $t$. Households are in the same category $g$ if they have the same number of household members of each type. 
- Individual-level characteristics: $\pi_{i,h}$
- Household-level characteristics: $\zeta_{i,h}$
- The household purchases private material goods $c_{i,h}$ and public material goods $X$. Private good $j \epsilon \Omega^c$ has price $p_j$. Public good $j \epsilon \Omega^X$ has price $r_j$.
- Each individual has time-endowment normalised to 1.
- Four types of time-use: private leisure $l_{i,h}$, joint leisure $jt_{i,h}$, market work $m_{i,h}$ (hours worked, does not enter utility function) and domestic work $d_{i,h}$
- Joint leisure and domestic work are inputs to public goods $D = f^D(\mathbf{d})$ and $JT = f^{JT}(\mathbf{jt})$ enjoyed by the household
 Private leisure $l_{i,h}$ includes leisure activities enjoyed by an individual without the co-presence of other household members. It is therefore a private good, which enters the utility function directly.
- $Q_h$ for the vector of public goods including both material and time-use public goods ($X_h$, $JT_h$, $D_h$) and $R_h$ for the associated price vector. $R_h$ includes wages $w_{i,h}$ of each of its members.
- Each individual has labour income endowment equal to $w_{i,h}$. An individual’s overall endowment is $y_{i,h} = y_{i,h}^{NL} + w_{i,h}$. 
- Each member can be endowed with nonlabour income $y_{i,h}^{NL}$. The household as a whole can have non-labour income $y_{h}^{NL}$. 
- The household’s endowment (or household budget) is $y_h = \sum_{i \epsilon h} y_{i,h}$. 
- Each individual i of type t living in a household category g has utility function $u_{t,g}$. There is preference heterogeneity across types and household compositions. The form of the bargaining solution of each household depends on the number and types of members, so that the sharing rule for different household categories must be estimated separately.
- Household individuals bargain over how to divide resources. Relative bargaining power will in general depend both on (i) market variables such as prices (including wages), and (ii) distribution factors, which enter the household’s optimisation problem only indirectly through the distribution of bargaining power $\pi_{i,h}, i \epsilon h$ and $\zeta_{i,h}$. The vector of variables that affect bargaining power is $z_h$.

## Browning, Chiappori and Weiss (2014): 
- The problem solved by any collective household, regardless of the underlying bargaining process and outside options, can be represented as an optimisation problem where the maximand is the weighted sum of the members’ utility functions. Each member’s utility function is weighed by their Pareto weight $µ_{i,h} (zh)$.
- Maximise $\sum_{i \epsilon h} (\mu_{i,h}(z_h) u_{t,g}(c_{i,h}, X_h, l_{i,h}, JT_h, D_h))$. Each person's utility depends on individual consumption of private material goods, all public goods, private leisure, joint leisure, and domestic work. 

## Arduini (2025) household's problem:
- Stage 1: The household chooses expenditure on market and time-produced (converted to wage units) public goods $R_{h}Q_{h} = \sum_{j \epsilon \Omega^X} r^j X_{h}^j + \sum_{i \epsilon h} w_{i,h} (d_{i,h} + jt_{i,h})$, where $Q_h = (X_h, JT_h, D_h)$, and how to divide the remaining household budget into individual budgets $\rho_{i,h} = (y_h − R_{h}Q_{h}) \eta_{i,h}$. The individual achieves indirect utility $v_{t,g}$ under $\rho_{i,h}$.

$$\text{max}_{\rho,X,d,jt} \sum_{i \epsilon h} \mu_{i,h}(z_h) v_{t,g}(\rho_{i,h}, X_h, l_{i,h}, JT_h, D_h)$$

$$\text{subject to} \newline
y_h = \sum_{i \epsilon h} w_{i,h} (d_{i,h} + jt_{i,h}) +\sum_{j \epsilon \Omega^X} r^j X_{h}^j + \sum_{i \epsilon h} \rho_{i,h}, \newline
d_{i,h} + jt_{i,h} \leq 1, \newline
X_{h}^j, \rho_{i,h}, d_{i,h}, jt_{i,h} \leq 0, \newline
D = f^D(\mathbf{d}),
JT = f^{JT}(\mathbf{jt})$$

- Sharing rule: determines the share of household budget net of public good expenditure assigned to each member (the individual’s resource share $\eta_{i,h}$)
- Stage 2: Members decide how to allocate their individual budgets $\rho_{i,h}$ to private expenditure.

Everyone, regardless of their type (in a given household category) is assumed to have the same Cobb-Douglas preference parameter for private leisure: $\alpha_{t,g}^l = \alpha_{g}^l$. Preference heterogeneity across types is maintained for all other preference parameters, and across household compositions.

## Arduini (2025) identifying assumptions:
Private leisure C-D SRAT: the ratio between men and women’s preference parameter for the private assignable good is stable across household compositions. Cubb-Douglas preferences over the private assignable good: homotheticity of private leisure.

Using family data probably violates assumption that total expenditure on public goods does not vary with the distribution of bargaining power in the household (what Arduini calls public good SAP). "With Cobb-Douglas preferences, we can achieve the desired restriction with public good C-D SAP. Under public good C-D SAP, the resource share and Pareto weight coincide. Public good C-D SAP imposes that the sum of all public good preference parameters is the same for all individuals of the same household category g."

Since I wanted to study family-related stuff in the first place, to that end, I think it's worth taking a different look at the role children play in their households. What I originally thought to study were income effects on fertility, and some literature in that subject finds evidence for children being "normal goods" in the sense that fertility increases with income. Would it be reasonable to take the public goods approach, that both parents could allocate resources (time) to "producing the public good" (parenting their children so that they grow up to have characteristics that the family wants as a whole)? btw time spent caring for household children can be for non-own children too?

The question of how to deal with children is a big issue. Time spent parenting (maybe focus on solo parenting, i.e. without spouse present, could also do money spent on kids) has some of the same characteristics as private leisure time. Parenting is also an unpaid activity, but its value isn't really known (even for those who work) since children aren't true goods---they are autonomous people and have their own preferences/choices, especially in adulthood. Treating children as a private good, where each parent can spend more time and/or own resources on raising their shared children for some private benefit, not necessarily something that the other spouse can benefit from (like teaching your child traits that are socially undesirable but personally useful), is complicated and I'm not sure if it's defensible. However, it is essentially the other side of the same coin as private leisure time. The more parenting you do, the less private leisure time you usually get. I just think parenting would be an interesting angle to take if I want to focus on families. Even if I can't use solo parenting time as a way to measure household sharing, it could be interesting to try and use it to answer my original question about the normal-ness or inferior-ness of children.

# Find the sharing rule

### Male and female specifications: deviations from average with own sex
```{r}
# leisure including eating, drinking, washing, spirituality, volunteering

# men
eq_m_within = leisure_exp_m ~ 0 + y + Bx_dev_wage_f_only + Bx_dev_wage_m_only + 
  Bx_dev_educ_f_only + Bx_dev_educ_m_only + Bx_dev_avgage + Bx_dev_agegap + Bx_dev_gdppc

# women
eq_f_within = leisure_exp_f ~ 0 + y + Bx_dev_wage_f_only + Bx_dev_wage_m_only + 
  Bx_dev_educ_f_only + Bx_dev_educ_m_only + Bx_dev_avgage + Bx_dev_agegap + Bx_dev_gdppc

# relaxing, socialising, sport, entertainment, and hobbies only

# men
eq_m_r_within = leisure_exp_m_r ~ 0 + y + Bx_dev_wage_f_only + Bx_dev_wage_m_only + 
  Bx_dev_educ_f_only + Bx_dev_educ_m_only + Bx_dev_avgage + Bx_dev_agegap + Bx_dev_gdppc

# women
eq_f_r_within = leisure_exp_f_r ~ 0 + y + Bx_dev_wage_f_only + Bx_dev_wage_m_only + 
  Bx_dev_educ_f_only + Bx_dev_educ_m_only + Bx_dev_avgage + Bx_dev_agegap + Bx_dev_gdppc

# list of regressions
eqns_within = list(male = eq_m_within, female = eq_f_within)
eqns_r_within = list(male = eq_m_r_within, female = eq_f_r_within)
```

### Male and female specifications: deviations from average of men AND women
```{r}
# leisure including eating, drinking, washing, spirituality, volunteering

# men
eq_m_between = leisure_exp_m ~ 0 + y + Bx_dev_wage_f_all + Bx_dev_wage_m_all + 
  Bx_dev_educ_f_all + Bx_dev_educ_m_all + Bx_dev_avgage + Bx_dev_agegap + Bx_dev_gdppc

# women
eq_f_between = leisure_exp_f ~ 0 + y + Bx_dev_wage_f_all + Bx_dev_wage_m_all + 
  Bx_dev_educ_f_all + Bx_dev_educ_m_all + Bx_dev_avgage + Bx_dev_agegap + Bx_dev_gdppc

# relaxing, socialising, sport, entertainment, and hobbies only

# men
eq_m_r_between = leisure_exp_m_r ~ 0 + y + Bx_dev_wage_f_all + Bx_dev_wage_m_all + 
  Bx_dev_educ_f_all + Bx_dev_educ_m_all + Bx_dev_avgage + Bx_dev_agegap + Bx_dev_gdppc

# women
eq_f_r_between = leisure_exp_f_r ~ 0 + y + Bx_dev_wage_f_all + Bx_dev_wage_m_all + 
  Bx_dev_educ_f_all + Bx_dev_educ_m_all + Bx_dev_avgage + Bx_dev_agegap + Bx_dev_gdppc

# list of regressions
eqns_between = list(male = eq_m_between, female = eq_f_between)
eqns_r_between = list(male = eq_m_r_between, female = eq_f_r_between)
```

### No restrictions on coefficients: deviations from average with own sex
```{r}
# leisure including eating, drinking, washing, spirituality, volunteering
fit_unres_within = systemfit(eqns_within, method = "SUR", data = hh)
# relaxing, socialising, sport, entertainment, and hobbies only
fit_unres_r_within = systemfit(eqns_r_within, method = "SUR", data = hh)
```

### No restrictions on coefficients: deviations from average of men AND women
```{r}
# leisure including eating, drinking, washing, spirituality, volunteering
fit_unres_between = systemfit(eqns_between, method = "SUR", data = hh)
# relaxing, socialising, sport, entertainment, and hobbies only
fit_unres_r_between = systemfit(eqns_r_between, method = "SUR", data = hh)
```

### Restricted coefficients, within-sex deviations
```{r}
# leisure including eating, drinking, washing, spirituality, volunteering
fit_res_within = systemfit(eqns_within, method = "SUR", data = hh,
                           restrict.regMat = modReg_within)
# relaxing, socialising, sport, entertainment, and hobbies only
fit_res_r_within = systemfit(eqns_r_within, method = "SUR", data = hh, 
                             restrict.regMat = modReg_within)
```

### Restricted coefficients, deviations from average of men AND women
```{r}
# leisure including eating, drinking, washing, spirituality, volunteering
fit_res_between = systemfit(eqns_between, method = "SUR", data = hh,
                            restrict.regMat = modReg_between)
# relaxing, socialising, sport, entertainment, and hobbies only
fit_res_r_between = systemfit(eqns_r_between, method = "SUR", data = hh,
                              restrict.regMat = modReg_between)
```

Print out some of these coefficients:
```{r}
summary(fit_res_r_within)
```

Unrestricted coefficients:
```{r}
test1 = fit_res_r_between$coefficients[c("female_y")]
test2 = fit_res_r_between$coefficients[c("male_y")]

print(test1 + test2)


```

Estimating the sharing rule:
```{r}
within_interactions <- c(
  "Bx_dev_wage_f_only",
  "Bx_dev_wage_m_only",
  "Bx_dev_educ_f_only",
  "Bx_dev_educ_m_only",
  "Bx_dev_avgage",
  "Bx_dev_agegap",
  "Bx_dev_gdppc"
)

res_within <- estimate_sharing_rule(
  fit = fit_unres_within,   # or fit_res_within if using restrictions
  data  = hh,
  interaction_vars = within_interactions
)

# Household-specific estimated shares:
hh_within <- res_within$hh_shares

# Average shares:
res_within$eta0_f
res_within$eta0_m

# Marginal impacts (on female share):
res_within$eta_z_f

```

# To do: notes

- actually calculate the resource shares lol

- deviations from averages: compare to other women or other men and women? try both of these. what’s my replacement value? bargaining power may come from strength relative to spouse, but strength relative to other options within your gender. say in 2006 there is a leisure gap when both contribute similar to household. over time, women contribute more to the household, but this may result in smaller increases in women’s leisure than if it were men. tease out whether you’re earning more relative to your husband or relative to other women. two different regressions, and including both in the same regression

- if women get richer does time spend with child increase or decrease? extra dollar with woman’s wage = more time with baby for the woman and maybe less time for the man? would it be the other way around for the man? if both household members have the same wage increase, does it mean the woman will still be the one taking care of the kid or will there be a shift

- changes over time: take income and matchings from 2004, and take coefficients from 2024. apply the betas to the observed behaviour of 2024, would i get different counterfactual leisures?

- margins plot??

- child gender differences: having boys vs girls (this is plausibly exogenous) and some results in literature find that men tend to want to spend time with sons but not daughters. (check gender ratios of births in states) I think one way to do this would be to use the plausible (at least in an American context) exogeneity of having more sons versus more daughters in the household. Ultimately, I think I want to find the predicted resource shares using the sharing rule and the SURE regressions and then find the effect of resource shares on childcare provision for men and women using son and daughter ratios(?) as a source of plausible exogeneity.

# Thought jail

My project (roughly): What is the impact of women’s control over their earnings (their bargaining power in the home) on their "child-rearing" supply? Using method as in Arduini to find the sharing rule, say who has more bargaining power based on private leisure, and then see if the person with more power spends more or less time on childcare. Developed from literature on the income effects of household fertility choice and intra-household bargaining, I use the American Time Use Survey (ATUS) to extend existing collective household models estimating intra-household inequality. In line with existing methodology, my Identification relies on private assignable goods, goods that cannot be shared between more than one person and are observed to be consumed by only one person in the household. This information is typically found in household expenditure surveys, but such goods are typically durable in the sense that they appear sparingly in expenditure data. To address the resulting power limitations with this approach, I adapt the approach of Arduini (2025) to a setting with parents and children, using private leisure in time use data to estimate bargaining power in the household. I then take this sharing rule to further discussions of childcare provision based on the bargaining power of individual parents. I both estimate and test for heterogeneity of childcare provision. I want to see the differences between male- or female-"dominated" households? Some existing papers use earnings/education differentials as a measure of bargaining power or fertility rates as a way to talk about family planning, but culture and gender roles may play a role in within-household childcare decisions and time spent on childcare is very tractable considering the breadth of time use data available for many countries.

For the purpose of this exercise, I'm not too sure what I want to focus on yet. A good place to start could be nuclear heterosexual households (i.e. two parents, a male partner, a female partner, and children) since I want to look at inequality within the heteronormative household.

## The literature

Literature on household inequality generally tends to and focuses on working couples *without* children, or models them as public goods. \[Arduini (2024)\]\[https://ifs.org.uk/sites/default/files/2025-05/WP202517-Estimating-intra-household-sharing-from-time-use-data.pdf\] uses the concept of private leisure time (time spent on leisure activities without other members of the household present) to estimate intra-household sharing. She defines the "assignable good", which is a private good. These parts are useful:

*"The main limitation of private leisure is that its price is only known for those who work, for whom it coincides with the observed wage. For non-participants, the price of time is higher than their (unobserved) wage and we would have to estimate that in order to use time-use data to identify sharing. That extension is left for future work: in my application to UK data, I focus on working couples."*

\[Blundell et al (2019)\]\[https://www.sfu.ca/\~pendakur/papers/BKLP16_longtitle.pdf\] still need to read this lol

\[Hyee and Robledo (2010)\]\[https://www.nottingham.ac.uk/economics/documents/discussion-papers/10-06.pdf\]

Calvi (2002): Under some mild assumptions, there exists a monotonic correspondence between Pareto weights and resource shares. Consequently, resource shares may also be interpreted as a measure of intrahousehold bargaining power. In effect, these are more tractable measures of power, as they are invariant to unobservable car- dinalizations of the utility functions.

\[Andrew et al (2021)\]\[https://ifs.org.uk/inequality/wp-content/uploads/2021/03/BN-The-careers-and-time-use-of-mothers-and-fathers.pdf\] show that heteronormative households, mothers are more likely to leave the labour force after having children regardless of the relative wages between mother and father before childbirth. (Much of the recent literature in this area focuses on COVID-19, maybe look into this.) It's difficult to really prove who in a relationship is doing the colloquially-named "mental load" of parenting (stuff like keeping appointments, planning meals, remembering when the laundry was last done), although anecdotally this is usually the wife or mother's job in this kind of household. (There may be a time use paper somewhere that tries to deal with this, maybe look into this). The ATUS records time spent doing activities, but not the thoughts that go on in someone's head. It also records the people present during an activity, but not, as far as I've seen, who is a active participant, or if someone needs to be actively prompted to do the groceries or pick up after a child.

\[Dunbar et al (2013)\]\[https://www.sfu.ca/\~pendakur/CollectiveChildren30-with-appendices.pdf\]: Geoff is looking at parents' and childrens' resource shares using expenditure data, I think I'd want to stick to the inequality of the couple. These parts are useful:

*"In models where goods are purely private or purely public, what we call private, assignable goods are known as exclusive goods. See, e.g. Bourguignon, Browning, and Chiappori (2009). Chiappori and Ekeland (2008) and Cherchye, De Rock and Vermeulen (2008) among others show how assignable goods can aid in the identification of resource shares. Our strategy follows this line in assuming the presence, and observability, of a small number of private assignable goods, and uses these to identify childrens’ resource shares. The end result is that we identify how total household expenditures on all goods are divided up among household members, just by observing how family expenditures on each member’s single private good (like clothing) vary with total expenditures."*

*"We find that children command a reasonably large share of resources – roughly 20 percent for the first child – and that this share rises with the number of children – 5-10 percentage points per additional child. Moreover, fathers command a larger share of resources than mothers, and mothers seem to sacrifice more resources than fathers to their children."*

*"A good is defined to be private if it cannot be shared or consumed jointly by more than one person, and is defined to be assignable if it is consumed by one individual household member that is known to the researcher. Examples could include toys and diapers which are private goods assignable to children, or alcohol and tobacco which are private goods assignable to adults. In our empirical application we use clothing"*

# What I'm doing with data

Household data: By splitting up the dataset by types of observations and then doing an inner merge myself, information collected at the household level can be used to populate the individual level data and vice versa. I think I'll have to construct my own data out of each type of observation because right now they all show up on separate lines. If I separate each type and merge them back together, I can fix this. Individual-level data: I create a dataset with observations for each person. Each household has a distinct household serial number, and all members of the household have the same serial number. I have ages for all household members. Activity and Who data: The ATUS has information about detailed activities and the people present during those activities. It also includes information on time spent taking care of children during those activities (secondary childcare)

Children and parents: I might want to look at a setup similar to DLP where there are father, mother, and child-type. The following activities are of interest:

Childcare-related activities from the ATUS documentation:

```         
030101  Physical care for household children
030102  Reading to or with household children
030103  Playing with household children, not sports
030104  Arts and crafts with household children
030105  Playing sports with household children
030106  Talking with or listening to household children
030107  Helping or teaching household children (not related to education) (2003)
030108  Organization and planning for household children
030109  Looking after household children (as a primary activity)
030110  Attending household children's events
030111  Waiting for or with household children
030112  Picking up or dropping off household children
030199  Caring for and helping household children, n.e.c.
030201  Homework (household children)
030202  Meetings and school conferences (household children)
030203  Home schooling of household children
030204  Waiting associated with household children's education
030299  Activities related to household child's education, n.e.c.
030301  Providing medical care to household children
030302  Obtaining medical care for household children
030303  Waiting associated with household children's health
030399  Activities related to household child's health, n.e.c.
080101  Using paid childcare services
080102  Waiting associated with purchasing childcare services
080199  Using paid childcare services, n.e.c.
160103  Telephone calls to or from education services providers
160107  Telephone calls to or from paid child or adult care providers
180301  Travel related to caring for and helping household children, inclusive (2003, 2004)
180302  Travel related to caring for and helping household children (2005+)
180303  Travel related to household children's education (2005+)
180304  Travel related to household children's health (2005+)
180801  Travel related to using childcare services
```

Private leisure activities from the ATUS documentation: needs to be done without the spouse or other household members present This includes all types of self care including sleeping and eating!

```         
General
010101  Sleeping    
010102  Sleeplessness   
010199  Sleeping, n.e.c.
010201  Washing, dressing, and grooming oneself
010299  Grooming, n.e.c.
010301  Health-related self care    
010399  Self care, n.e.c.
010401  Personal or private activities  
010499  Personal activities, n.e.c.
019999  Personal care, n.e.c.
050201  Socializing, relaxing, and leisure as part of job
020603  Walking, exercising, playing with animals (2008+)
120201  Attending or hosting parties, receptions, or ceremonies
120202  Attending meetings for personal interest (not volunteering)
120299  Attending or hosting social events, n.e.c.
120301  Relaxing, thinking
120302  Tobacco and drug use
120303  Television and movies (not religious)
120304  Television (religious)
120305  Listening to the radio
120306  Listening to or playing music (not radio)
120307  Playing games
120308  Computer use for leisure (excluding games)
120309  Arts and crafts as a hobby
120310  Collecting as a hobby
120311  Hobbies, except arts and crafts and collecting
120312  Reading for personal interest
120313  Writing for personal interest
120399  Relaxing and leisure, n.e.c.
120401  Attending performing arts
120402  Attending museums
120403  Attending movies or film
120404  Attending gambling establishments
120405  Security procedures related to arts and entertainment
120499  Arts and entertainment, n.e.c.
120501  Waiting associated with socializing and communicating
120502  Waiting associated with attending or hosting social events
120503  Waiting associated with relaxing or leisure
120504  Waiting associated with arts and entertainment
120599  Waiting associated with socializing, n.e.c.
129999  Socializing, relaxing, and leisure, n.e.c.
060102  Taking class for personal interest
060201  Extracurricular club activities
060202  Extracurricular music and performance activities
060203  Extracurricular student government activities
060204  Waiting associated with extracurricular activities (2004+)
060299  Education-related extracurricular activities, n.e.c.
060302  Research or homework for class (for personal interest)
060402  Administrative activities: class for personal interest
070104  Shopping, except groceries, food, and gas
110101  Eating and drinking
050202  Eating and drinking as part of job
110201  Waiting associated with eating and drinking
110299  Waiting associated with eating and drinking, n.e.c.
119999  Eating and drinking, n.e.c.
120101  Socializing and communicating with others
120199  Socializing and communicating, n.e.c.
160101  Telephone calls to or from family members
160102  Telephone calls to or from friends, neighbors, or acquaintances
080501  Using personal care services
080502  Waiting associated with personal care services
080599  Using personal care services, n.e.c.

Sport
130101  Doing aerobics
130102  Playing baseball
130103  Playing basketball
130104  Biking
130105  Playing billiards
130106  Boating
130107  Bowling
130108  Climbing, spelunking, caving
130109  Dancing
130110  Participating in equestrian sports
130111  Fencing
130112  Fishing
130113  Playing football
130114  Golfing
130115  Doing gymnastics
130116  Hiking
130117  Playing hockey
130118  Hunting
130119  Participating in martial arts
130120  Playing racquet sports
130121  Participating in rodeo competitions
130122  Rollerblading
130123  Playing rugby
130124  Running
130125  Skiing, ice skating, snowboarding
130126  Playing soccer
130127  Playing softball
130128  Using cardiovascular equipment
130129  Vehicle touring or racing
130130  Playing volleyball
130131  Walking
130132  Participating in water sports
130133  Weightlifting or strength training
130134  Working out, unspecified
130135  Wrestling
130136  Doing yoga
130199  Playing sports, n.e.c.
130201  Watching aerobics
130202  Watching baseball
130203  Watching basketball
130204  Watching biking
130205  Watching billiards
130206  Watching boating
130207  Watching bowling
130208  Watching climbing, spelunking, caving
130209  Watching dancing
130210  Watching equestrian sports
130211  Watching fencing
130212  Watching fishing
130213  Watching football
130214  Watching golfing
130215  Watching gymnastics
130216  Watching hockey
130217  Watching martial arts
130218  Watching racquet sports
130219  Watching rodeo competitions
130220  Watching rollerblading
130221  Watching rugby
130222  Watching running
130223  Watching skiing, ice skating, snowboarding
130224  Watching soccer
130225  Watching softball
130226  Watching vehicle touring or racing
130227  Watching volleyball
130228  Watching walking
130229  Watching water sports
130230  Watching weightlifting or strength training
130231  Watching people working out, unspecified
130232  Watching wrestling
130299  Attending sporting events, n.e.c.
130301  Waiting related to playing sports or exercising
130302  Waiting related to attending sporting events
130399  Waiting associated with sports, exercise, and recreation, n.e.c.

Religion
140101  Attending religious services    
140102  Participation in religious practices    
140103  Waiting associated with religious and spiritual activities  
140104  Security procedures related to religious and spiritual activities
140105  Religious education activities (2007+)
149999  Religious and spiritual activities, n.e.c.

Volunteering
150101  Computer use    
150102  Organizing and preparing    
150103  Reading 
150104  Telephone calls (except hotline counseling) 
150105  Writing 
150106  Fundraising 
150199  Administrative and support activities, n.e.c.
150201  Food preparation, presentation, clean-up
150202  Collecting and delivering clothing and other goods
150203  Providing care  
150204  Teaching, leading, counseling, mentoring    
150299  Social service and care activities, n.e.c.
150301  Building houses, wildlife sites, and other structures
150302  Indoor and outdoor maintenance, repair, and clean-up
150399  Indoor and outdoor maintenance, building, and clean-up activities, n.e.c.
150402  Serving at volunteer events and cultural activities
150499  Participating in performance and cultural activities, n.e.c.
150500  Attending Meetings, Conferences, and Training
150501  Attending meetings, conferences, and training   
150599  Attending meetings, conferences, and training, n.e.c.
150601  Public health activities
150602  Public safety activities    
150699  Public health and safety activities, n.e.c. 
150701  Waiting associated with volunteer activities (2004+)
150799  Waiting associated with volunteer activities, n.e.c. (2004+)
150801  Security procedures related to volunteer activities (2007+)
150899  Security proecdures related to voluteer activities, n.e.c. (2007+)
159999  Volunteer activities, n.e.c.

Travel
180805  Travel related to using personal care services
181201  Travel related to socializing and communicating 
181202  Travel related to attending or hosting social events    
181101  Travel related to eating and drinking
181199  Travel related to eating and drinking, n.e.c.
181203  Travel related to relaxing and leisure (2003, 2004)
181204  Travel related to arts and entertainment    
181205  Travel as a form of entertainment (2005+)
181206  Travel related to relaxing and leisure (2005+)
181299  Travel related to socializing, relaxing, and leisure, n.e.c.
181300  Travel Related to Sports, Exercise, and Recreation
181301  Travel related to participating in sports, exercise, or recreation  
181302  Travel related to attending sporting or recreational events
181401  Travel related to religious or spiritual practices
181499  Travel related to religious or spiritual activities, n.e.c.
181501  Travel related to volunteering
181599  Travel related to volunteer activities, n.e.c.
```

People do activities with...

```         
0100    Alone   
0200    Spouse  
0201    Unmarried partner   
0202    Own household child 
0203    Grandchild  
0204    Parent  
0205    Brother sister  
0206    Other related person    
0207    Foster child    
0208    Housemate, roommate 
0209    Roomer, boarder 
0210    Other nonrelative   
0300    Own non-household child under 18    
0400    Parents (not living in household)   
0401    Other non-household family members under 18 
0402    Other non-household family members 18 and older (including parents-in-law)  
0403    Friends 
0404    Co-workers, colleagues, clients (non-work activities only)
0405    Neighbors, acquaintances    
0406    Other non-household children under 18   
0407    Other non-household adults 18 and older 
0408    Boss or manager
0409    People whom I supervise
0410    Co-workers
0411    Customers
9996    Refused
9997    Don't know  
9998    Blank   
```

# Random summary statistics and stuff

## Weird age gaps?

```{r}
jay_and_gloria = individual |> filter(age_gap > 30)
print(jay_and_gloria[c("YEAR", "SERIAL", "SEX", "AGE")])
```

## Number of households in each year

```{r}
hh |>
  count(YEAR, name = "n_households")
```

## Zeros in leisure and childcare

Something that is supposed to be advantageous about private leisure is that people tend to "consume" it more than other private assignable goods like clothing (which is durable and not purchased as often as one would watch TV). However, it looks like ATUS respondents just don't tend to have a lot of fun. 

```{r}
zeros_leisure_childcare = individual |>
  group_by(YEAR) |>
  # find people who don't have any private leisure time use
  # or no-spouse-present childcare time use
  summarise(
    # at least one person in hh with no private leisure
    pct_any_zero_private_leisure  = mean(any_zero_private_leisure),
    # both people in hh with no private leisure
    pct_both_zero_private_leisure = mean(both_zero_private_leisure),
    
    # at least one person in hh with no private leisure (restricted def)
    pct_any_zero_private_leisure_r  = mean(any_zero_private_leisure_r),
    # both people in hh with no private leisure (restricted def)
    pct_both_zero_private_leisure_r = mean(both_zero_private_leisure_r),
    
    # at least one person in hh with no no-spouse-present childcare
    pct_any_zero_childcare  = mean(any_zero_childcare),
    # both people in hh with no no-spouse-present childcare
    pct_both_zero_childcare = mean(both_zero_childcare))

zeros_leisure_childcare
```

## Comparing leisure definitions

Expenditure refers to hours spent on leisure valued at the person’s hourly wage. I currently have two definitions of leisure: one very broadly counts sleeping, eating, and washing/grooming (as defined by Arduini) and one only includes recreation/social activities. Women generally command more leisure expenditure than men in the broader definition of leisure, but less in the narrower definition, but this varies a lot for both men and women.

```{r}
hh |>
  pivot_longer(
    cols = c(leisure_exp_m, leisure_exp_f, leisure_exp_m_r, leisure_exp_f_r),
    names_to = "var",
    values_to = "leisure_exp") |>
  mutate(sex = case_when(grepl("_m", var) ~ "Men", grepl("_f", var) ~ "Women"),
         type = case_when(grepl("_r$", var) ~ "Restricted leisure expenditure",
                          TRUE ~ "Total leisure expenditure")) |>
  group_by(sex, type) |>
  summarise(mean = mean(leisure_exp, na.rm = TRUE),
            sd = sd(leisure_exp, na.rm = TRUE),
            q1 = quantile(leisure_exp, 0.25, na.rm = TRUE),
            q2 = quantile(leisure_exp, 0.50, na.rm = TRUE),
            q3 = quantile(leisure_exp, 0.75, na.rm = TRUE),
            .groups = "drop") |>
  arrange(type, sex)
```

## Comparing age and ability

Men tend to be older and less educated, but earn more than women, although there is also a lot of variation in earnings.

```{r}
hh |>
  pivot_longer(cols = c(educ_f, educ_m, wage_f, wage_m, age_gap),
               names_to = "var",
               values_to = "value") |>
  mutate(sex = case_when(grepl("educ_m", var) ~ "Education men (levels 0,1,2)",
                         grepl("educ_f", var) ~ "Education women (levels 0,1,2)",
                         grepl("wage_m", var) ~ "Hourly earnings men ($)",
                         grepl("wage_f", var) ~ "Hourly earnings women ($)",
                         grepl("age_gap", var) ~ "Age gap (men - women)")) |>
  group_by(sex) |>
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE),
            q1 = quantile(value, 0.25, na.rm = TRUE),
            q2 = quantile(value, 0.50, na.rm = TRUE),
            q3 = quantile(value, 0.75, na.rm = TRUE),
            .groups = "drop") |>
  arrange(sex)
```

## Comparing childcare and working hours

I compare working hours between men and women, and childcare performed by one spouse without the other present. Men tend to work more hours, while childcare undertaken by one parent varies a lot within the sample. This makes sense as I haven’t taken any restrictions on child ages at this point, but women tend to spend more time on childcare at the upper end of the distribution. I’m guessing this is the part of the sample with younger children.

```{r}
hh |>
  pivot_longer(
    cols = c(total_childcare_nospouse_h_f, total_childcare_nospouse_h_m,
             uhrsworkt_f, uhrsworkt_m),
    names_to = "var",
    values_to = "value") |>
  
  # Label sex and category clearly
  mutate(sex = case_when(grepl("_f$", var) ~ "Women", 
                         grepl("_m$", var) ~ "Men"),
    type = case_when(grepl("childcare", var) ~ "Childcare (no spouse present)",
                     grepl("uhrsworkt", var) ~ "Usual weekly hours worked")) |>
  
  # Summary statistics by sex and variable type
  group_by(sex, type) |>
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE),
            q1 = quantile(value, 0.25, na.rm = TRUE),
            q2 = quantile(value, 0.50, na.rm = TRUE),
            q3 = quantile(value, 0.75, na.rm = TRUE),
    .groups = "drop") |>
  arrange(type, sex)
```

## Male/female broad leisure expenditure
There are a few people who regularly work very few hours a week, yet earn a lot. Maybe they are anesthesiologists
```{r}
ggplot(hh) +
  # points
  geom_point(aes(x = leisure_exp_f, y = y), color = "pink") +
  geom_point(aes(x = leisure_exp_m, y = y), color = "lightblue", shape=23) +
  # lines of best fit
  geom_smooth(aes(x = leisure_exp_f, y = y), method = "lm", color = "pink", se = FALSE) +
  geom_smooth(aes(x = leisure_exp_m, y = y), method = "lm", color = "lightblue", se = FALSE) +
  labs(x = "leisure expenditure (broad definition)", y = "household budget") +
  theme_minimal()
```

## Male/female broad leisure expenditure

```{r}
ggplot(hh) +
  # points
  geom_point(aes(x = leisure_exp_f_r, y = y), color = "pink") +
  geom_point(aes(x = leisure_exp_m_r, y = y), color = "lightblue", shape=23) +
  # lines of best fit
  geom_smooth(aes(x = leisure_exp_f_r, y = y), method = "lm", color = "pink", se = FALSE) +
  geom_smooth(aes(x = leisure_exp_m_r, y = y), method = "lm", color = "lightblue", se = FALSE) +
  labs(x = "leisure expenditure (narrow definition)", y = "household budget") +
  theme_minimal()
```
