---
title: "Showing my data work (for supervisory purposes)"
format: html
editor: visual
---

```{r include=FALSE}
library(tidyverse)
library(systemfit)
library(ggplot2)
library(purrr)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(lmtest)
library(sandwich)
library(car)
library(patchwork)
library(fixest)
library(patchwork)
library(marginaleffects)
library(corrplot)

# actlines and directories
source("UKTUS_params.R")

# functions
source("functions.R")

# get all output
source("run_SUR.R")

# for display purposes
options(scipen=1)
options(digits=4)
```


# Estimating the sharing rule using household time use data

```{r}

summary_parents = c("wage_f", "wage_m", "DVAge_f", "DVAge_m", "educ_f", 
                    "educ_m", "num_kids_total", "num_under_5")
summary_nonparents = c("wage_f", "wage_m", "DVAge_f", "DVAge_m", "educ_f", 
                    "educ_m")

nonparents_summary = t(sapply(nonparents_est_data_merged_weekday[summary_nonparents], summarytable))
parents_summary = t(sapply(parents_est_data_merged_weekday[summary_parents], summarytable))

print("UKTUS working heterosexual couples (parents), pooled sample")
print(as.data.frame(parents_summary))
print("UKTUS working heterosexual couples (nonparents), pooled sample")
print(as.data.frame(nonparents_summary))
```

## Baseline results

Weekdays:
```{r}
# Parents, narrow leisure, weekday
print("Regression results (sharing rule) of parents (excluding sleep and personal care, weekday)")
print(make_sysfit_kable_df(parents_res_r_within_merged, p_rows_within))

# Parents, broad leisure, weekday
print("Regression results (sharing rule) of parents (including sleep and personal care, weekday)")
print(make_sysfit_kable_df(parents_res_within_merged, p_rows_within))

# Non-parents, narrow leisure, weekday
print("Regression results (sharing rule) of non-parents (excluding sleep and personal care, weekday)")
print(make_sysfit_kable_df(nonparents_res_r_within_merged, np_rows_within))

# Non-parents, broad leisure, weekday
print("Regression results (sharing rule) of non-parents (including sleep and personal care, weekday)")
print(make_sysfit_kable_df(nonparents_res_within_merged, np_rows_within))
```

Weekends:
```{r}
# Parents, narrow leisure, weekend
print("Regression results (sharing rule) of parents (excluding sleep and personal care, weekend)")
print(make_sysfit_kable_df(parents_res_r_weekend, p_rows_within))

# Parents, broad leisure, weekend 
print("Regression results (sharing rule) of parents (including sleep and personal care, weekend)")
print(make_sysfit_kable_df(parents_res_weekend, p_rows_within))

# Non-parents, narrow leisure, weekend
print("Regression results (sharing rule) of non-parents (excluding sleep and personal care, weekend)")
print(make_sysfit_kable_df(nonparents_res_r_weekend, np_rows_within))

# Non-parents, broad leisure, weekend
print("Regression results (sharing rule) of non-parents (including sleep and personal care, weekend)")
print(make_sysfit_kable_df(nonparents_res_weekend, np_rows_within))

```
Separating sharing rules by year:
```{r}
# Parents, narrow leisure, weekday
print("2000 sharing rule, parents (excluding sleep and personal care, weekday)")
print(make_sysfit_kable_df(parents_res_r_within_2000, p_rows_within))
print("2015 sharing rule, parents (excluding sleep and personal care, weekday)")
print(make_sysfit_kable_df(parents_res_r_within_2015, p_rows_within))

# Parents, broad leisure, weekday
print("2000 sharing rule, parents (including sleep and personal care, weekday)")
print(make_sysfit_kable_df(parents_res_within_2000, p_rows_within))
print("2015 sharing rule, parents (including sleep and personal care, weekday)")
print(make_sysfit_kable_df(parents_res_within_2015, p_rows_within))
```

Equality of parents and non-parents:
```{r}
print(t.test(shares_nonparents_ownsex_merged$data$shareown_etahat_f))

print(t.test(shares_parents_ownsex_merged$data$shareown_etahat_f))

print(t.test(shares_parents_ownsex_merged$data$shareown_etahat_f,
             shares_nonparents_ownsex_merged$data$shareown_etahat_f))
```

```{r}
print(t.test(shares_nonparents_ownsex_r_merged$data$shareown_etahat_f))

print(t.test(shares_parents_ownsex_r_merged$data$shareown_etahat_f))

print(t.test(shares_parents_ownsex_r_merged$data$shareown_etahat_f,
             shares_nonparents_ownsex_r_merged$data$shareown_etahat_f))
```

```{r fig.width=10, fig.height=8}

narrowleisure_shares_p = plot_share_densities(parents_est_data_merged_shares |> filter(is_weekend == 0), 
                                            restrict_leisure = TRUE)

broadleisure_shares_p = plot_share_densities(parents_est_data_merged_shares |> filter(is_weekend == 0), 
                                           restrict_leisure = FALSE)

combined_plot_p =
  narrowleisure_shares_p + broadleisure_shares_p +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Estimated resource shares (parents)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))

narrowleisure_shares_np = plot_share_densities(nonparents_est_data_merged_shares |> filter(is_weekend == 0),
                                               restrict_leisure = TRUE)

broadleisure_shares_np = plot_share_densities(nonparents_est_data_merged_shares |> filter(is_weekend == 0), 
                                           restrict_leisure = FALSE)

combined_plot_np =
  narrowleisure_shares_np + broadleisure_shares_np +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Estimated resource shares (non-parents)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))

combined_plot_baseline = wrap_elements(combined_plot_p) | wrap_elements(combined_plot_np)

print(combined_plot_baseline)

```
Marginal impacts tables
```{r}
marginal_impacts_parents_weekday = marginal_impacts(shares_parents_ownsex_merged, 
                                           parents_est_data_merged, p_rows_within)

marginal_impacts_nonparents_weekday = marginal_impacts(shares_nonparents_ownsex_merged, 
                                           nonparents_est_data_merged, np_rows_within, 
                                           data_type = "nonparents")

marginal_impacts_parents_r_weekday = marginal_impacts(shares_parents_ownsex_r_merged, 
                                           parents_est_data_merged, p_rows_within)

marginal_impacts_nonparents_r_weekday = marginal_impacts(shares_nonparents_ownsex_r_merged, 
                                           nonparents_est_data_merged, np_rows_within, 
                                           data_type = "nonparents")


print("Interpretation of regression coefficients (excluding personal care and sleep, parents, weekday)")
print(marginal_impacts_parents_r_weekday)

print("Interpretation of regression coefficients (including personal care and sleep, parents, weekday)")
print(marginal_impacts_parents_weekday)

print("Interpretation of regression coefficients (excluding personal care and sleep, non-parents, weekday)")
print(marginal_impacts_nonparents_r_weekday)

print("Interpretation of regression coefficients (including personal care and sleep, non-parents, weekday)")
print(marginal_impacts_nonparents_weekday)
```
Weekday vs weekend:
```{r}
marginal_impacts_parents_weekend = marginal_impacts(shares_parents_weekend, 
                                           parents_est_data_merged, p_rows_within)

marginal_impacts_nonparents_weekend = marginal_impacts(shares_nonparents_weekend, 
                                           nonparents_est_data_merged, np_rows_within, 
                                           data_type = "nonparents")

marginal_impacts_parents_r_weekend = marginal_impacts(shares_parents_r_weekend, 
                                           parents_est_data_merged, p_rows_within)

marginal_impacts_nonparents_r_weekend = marginal_impacts(shares_nonparents_r_weekend, 
                                           nonparents_est_data_merged, np_rows_within, 
                                           data_type = "nonparents")


print("Interpretation of regression coefficients (excluding personal care and sleep, parents, weekend)")
print(marginal_impacts_parents_r_weekend)

print("Interpretation of regression coefficients (including personal care and sleep, parents, weekend)")
print(marginal_impacts_parents_weekend)

print("Interpretation of regression coefficients (excluding personal care and sleep, non-parents, weekend)")
print(marginal_impacts_nonparents_r_weekend)

print("Interpretation of regression coefficients (including personal care and sleep, non-parents, weekend)")
print(marginal_impacts_nonparents_weekend)
```

More comparing sharing rules:
```{r}
marginal_impacts_counterfactual_2000 = marginal_impacts(counterfactual_2000sharing_alldata, 
                                           parents_est_data_merged, p_rows_within)

marginal_impacts_counterfactual_2015 = marginal_impacts(counterfactual_2015sharing_alldata, 
                                           parents_est_data_merged, np_rows_within)


marginal_impacts_counterfactual_r_2000 = marginal_impacts(counterfactual_2000sharing_alldata_r, 
                                           parents_est_data_merged, p_rows_within)

marginal_impacts_counterfactual_r_2015 = marginal_impacts(counterfactual_2015sharing_alldata_r, 
                                           parents_est_data_merged, np_rows_within)

print("Interpretation of regression coefficients (combined sample, 2000 sharing rule, including personal care and sleep, weekday)")
print(marginal_impacts_counterfactual_2000)

print("Interpretation of regression coefficients (combined sample, 2015 sharing rule, including personal care and sleep, weekday)")
print(marginal_impacts_counterfactual_2015)

print("Interpretation of regression coefficients (combined sample, 2000 sharing rule, excluding personal care and sleep, weekday)")
print(marginal_impacts_counterfactual_r_2000)

print("Interpretation of regression coefficients (combined sample, 2015 sharing rule, excluding personal care and sleep, weekday)")
print(marginal_impacts_counterfactual_r_2015)
```

Comparing narrow and broad leisure definitions:
```{r}
compare_shares = parents_est_data_merged_shares |>
  mutate(
    diff_share_f = shareown_etahat_f - shareown_etahat_r_f,
    abs_diff_share_f = abs(diff_share_f)
  )

print(compare_shares |> summarise(
    mean_diff = mean(diff_share_f),
    mean_abs_diff = mean(abs_diff_share_f),
    p25 = quantile(abs_diff_share_f, 0.25),
    p50 = quantile(abs_diff_share_f, 0.50),
    p75 = quantile(abs_diff_share_f, 0.75),
    corr = cor(shareown_etahat_f, shareown_etahat_r_f)
  ))

print(compare_shares |>
  mutate(
    q_narrow = ntile(shareown_etahat_r_f, 4),
    q_broad  = ntile(shareown_etahat_f, 4)
  ) |>
  count(q_narrow, q_broad))
```

Comparing estimates across both UKTUS survey years:
```{r fig.width=10, fig.height=8}

narrowleisure_shares_2000 = plot_share_densities(parents_est_data_merged_shares |> filter(sample == "2000", is_weekend == 0), 
                                                 restrict_leisure = TRUE)

broadleisure_shares_2000 = plot_share_densities(parents_est_data_merged_shares |> filter(sample == "2000", is_weekend == 0), 
                                               restrict_leisure = FALSE)

narrowleisure_shares_2015 = plot_share_densities(parents_est_data_merged_shares |> filter(sample == "2015", is_weekend == 0), 
                                                 restrict_leisure = TRUE)

broadleisure_shares_2015 = plot_share_densities(parents_est_data_merged_shares |> filter(sample == "2015", is_weekend == 0),  
                                                restrict_leisure = FALSE)

combined_plot_2000 =
  narrowleisure_shares_2000 + broadleisure_shares_2000 +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Estimated resource shares in 2000 (parents)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))

combined_plot_2015 =
  narrowleisure_shares_2015 + broadleisure_shares_2015 +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Estimated resource shares in 2015 (parents)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))

combined_plot_years = wrap_elements(combined_plot_2000) | wrap_elements(combined_plot_2015)

print(combined_plot_years)
```

```{r fig.width=10, fig.height=8}

narrowleisure_shares_2000 = plot_share_densities(nonparents_est_data_merged_shares |> filter(sample == "2000", is_weekend == 0), 
                                                 restrict_leisure = TRUE)

broadleisure_shares_2000 = plot_share_densities(nonparents_est_data_merged_shares |> filter(sample == "2000", is_weekend == 0),
                                                restrict_leisure = FALSE)

narrowleisure_shares_2015 = plot_share_densities(nonparents_est_data_merged_shares |> filter(sample == "2015", is_weekend == 0),
                                                 restrict_leisure = TRUE)

broadleisure_shares_2015 = plot_share_densities(nonparents_est_data_merged_shares |> filter(sample == "2015", is_weekend == 0),
                                                restrict_leisure = FALSE)

combined_plot_2000 =
  narrowleisure_shares_2000 + broadleisure_shares_2000 +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Estimated resource shares in 2000 (non-parents)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))

combined_plot_2015 =
  narrowleisure_shares_2015 + broadleisure_shares_2015 +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Estimated resource shares in 2015 (non-parents)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))

combined_plot_years = wrap_elements(combined_plot_2000) | wrap_elements(combined_plot_2015)

print(combined_plot_years)
```

Comparing female resource share
```{r fig.width=10, fig.height=5}
plot_data_fshare = bind_rows(
    shares_parents_ownsex_merged$data       |> mutate(source = "Weekday (broad leisure definition)"),
    shares_parents_ownsex_r_merged$data     |> mutate(source = "Weekday"),
    shares_parents_r_weekend$data           |> mutate(source = "Weekend"),
    counterfactual_2000sharing_alldata_r$data |> mutate(source = "2000 sharing rule"),
    counterfactual_2015sharing_alldata_r$data |> mutate(source = "2015 sharing rule")
  )

  ggplot(plot_data_fshare, aes(x = shareown_etahat_f, colour = source)) +                                                             
    geom_density(bw = "nrd0", linewidth = 1, na.rm = TRUE, key_glyph = "point") +                                                     
    scale_colour_manual(                                                                                                              
      values = c(
        "Weekday (broad leisure definition)" = "#920000",                                                                             
        "Weekday"                            = "#490092",                                                                             
        "Weekend"                            = "#ff6db6",
        "2000 sharing rule"                  = "orange",                                                                              
        "2015 sharing rule"                  = "#22cf22"
      ),                                                                                                                              
      breaks = c("Weekday", "Weekday (broad leisure definition)",
                 "Weekend", "2000 sharing rule", "2015 sharing rule")                                                                 
    ) +           
    guides(colour = guide_legend(
      override.aes = list(shape = 15, size = 4)  # 15 = filled square; 16 = filled circle                                             
    )) +
    labs(x = "Estimated female resource share", y = "Density", colour = NULL) +                                                       
    xlim(0, 1) +                                                                                                                      
    theme_minimal() +
    theme(legend.position = "right", plot.title = element_text(face = "bold")) 
```

## Kind of testing assumptions:

- "If private leisure C-D SAP holds then the proportion of household budget spent on overall private leisure (the sum of expenditure on female private leisure and expenditure on male private leisure) should not vary with the distribution of bargaining power"
```{r}
# check budget share on private leisure correlated w/ estimated resource shares

print("test of correlations: estimated resource shares and total private leisure expenditure (including personal care and sleep")
# correlation of estimated resource shares (broad leisure definition)
cor.test(parents_est_data_merged_shares$share_budget_leisure, 
         parents_est_data_merged_shares$shareown_etahat_f)

print("test of correlations: estimated resource shares and total private leisure expenditure (excluding personal care and sleep")
# (narrow leisure definition)
cor.test(parents_est_data_merged_shares$share_budget_leisure_r, 
         parents_est_data_merged_shares$shareown_etahat_r_f)
```

## How private leisure responds to childcare need (constraints?)

Resource shares and child characteristics:
```{r}
# use parents_est_data_merged_shares find average of shareown_etahat_r_f at 
# above or below-median wage_f, educ_f, total_childcare_f 

wage_split_r = parents_est_data_merged_shares |>
  mutate(
    wage_f_high = wage_f > median(wage_f, na.rm = TRUE),
    wage_m_high = wage_m > median(wage_m, na.rm = TRUE)) |>
  group_by(wage_f_high, wage_m_high) |>
  summarise(
    mean_share_f = mean(shareown_etahat_r_f, na.rm = TRUE),
    .groups = "drop") |>
  mutate(
    wage_f_high = if_else(wage_f_high, "Above-median female wage", 
                          "Below-median female wage"),
    wage_m_high = if_else(wage_m_high, "Above-median male wage", 
                          "Below-median male wage")) |>
  pivot_wider(
    names_from = wage_m_high,
    values_from = mean_share_f)

child_split_r = parents_est_data_merged_shares |>
  mutate(
    wage_f_high = wage_f > median(wage_f, na.rm = TRUE),
    young_child = num_under_5 > 0) |>
  group_by(wage_f_high, young_child) |>
  summarise(
    mean_share_f = mean(shareown_etahat_r_f, na.rm = TRUE),
    .groups = "drop") |>
  mutate(
    wage_f_high = if_else(wage_f_high, "Above-median female wage", 
                          "Below-median female wage"),
    young_child = if_else(young_child, "Child under 5 in household", 
                          "No child under 5")) |>
  pivot_wider(
    names_from = young_child,
    values_from = mean_share_f)

num_split_r = parents_est_data_merged_shares |>
  mutate(
    wage_f_high = wage_f > median(wage_f, na.rm = TRUE),
    high_kids = num_kids_total > median(num_kids_total, na.rm=TRUE)) |>
  group_by(wage_f_high, high_kids) |>
  summarise(
    mean_share_f = mean(shareown_etahat_r_f, na.rm = TRUE),
    .groups = "drop") |>
  mutate(
    wage_f_high = if_else(wage_f_high, "Above-median female wage", 
                          "Below-median female wage"),
    high_kids = if_else(high_kids, "Above-median number of children", 
                        "Below-median number of children")) |>
  pivot_wider(
    names_from = high_kids,
    values_from = mean_share_f)

print(wage_split_r)
print(child_split_r)
print(num_split_r)
```

Table of time use data:

```{r}
# compute mean time use within each quartile restricted leisure shares
table_quartiles_r = timeuse_long |>
  group_by(share_quartile_r, day, sex, activity) |>
  summarise(
    mean_hours = mean(hours, na.rm = TRUE),
    sd_hours = sd(hours, na.rm = TRUE),
    n = n(),
    .groups = "drop") |>
  mutate(
    mean_hours = round(mean_hours, 2),
    sd_hours = round(sd_hours, 2)) |>
  select(day, sex, activity, share_quartile_r, mean_hours) |>
  pivot_wider(
    names_from = share_quartile_r,
    values_from = mean_hours) |>
  arrange(day, sex, activity) |>
  filter(activity %in% c(
    "Childcare",
    "Private leisure",
    "Private leisure (excluding sleep and personal care)",
    "Market work",
    "Domestic work excluding childcare"))

print(table_quartiles_r)
```

Non-parent regressions
```{r}
etable(feols(total_work ~ is_weekend * male
      | sample + dgorpaf,
      data = reg_stacked_nonparents,
      cluster = "serial"))

etable(feols(total_private_leisure ~ is_weekend * male
      | sample + dgorpaf,
      data = reg_stacked_nonparents,
      cluster = "serial"))

etable(feols(total_private_leisure_r ~ is_weekend * male
      | sample + dgorpaf,
      data = reg_stacked_nonparents,
      cluster = "serial"))

etable(feols(total_otherdomestic ~ is_weekend * male
      | sample + dgorpaf,
      data = reg_stacked_nonparents,
      cluster = "serial"))
```


```{r}
etable(feols(total_work ~ is_weekend * male
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))

etable(feols(total_childcare ~ is_weekend * male
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))

etable(feols(total_private_leisure ~ is_weekend * male
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))

etable(feols(total_private_leisure_r ~ is_weekend * male
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))

etable(feols(total_otherdomestic ~ is_weekend * male
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))
```

```{r}
etable(feols(total_work ~ is_weekend * male * child_under_5
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))

etable(feols(total_childcare ~ is_weekend * male * child_under_5
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))

etable(feols(total_private_leisure ~ is_weekend * male * child_under_5
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))

etable(feols(total_private_leisure_r ~ is_weekend * male * child_under_5
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))

etable(feols(total_otherdomestic ~ is_weekend * male * child_under_5
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))
```
Difference between male and female weekend adjustment = difference in difference
Compare difference in difference from:
- 25th percentile female wage ratio to 75th percentile female wage ratio
- female gaining one education level over male (educ_gap_f == 0 to educ_gap_f == 1)
- 25th percentile average couple age to 75th percentile average couple age
- no age gap to age gap of 10 years
- poorest region to richest region? go from poorest in 2015 to richest in 2015
- no child under 5 to one child under 5
- no children to 2 children

Also compare difference in difference from:
- 25th percentile female resource share to 75th percentile female resource share

```{r}
parent_wageratio_25 = quantile(reg_stacked$wage_ratio_f, 0.25)
parent_wageratio_75 = quantile(reg_stacked$wage_ratio_f, 0.75)

nonparent_wageratio_25 = quantile(reg_stacked_nonparents$wage_ratio_f, 0.25)
nonparent_wageratio_75 = quantile(reg_stacked_nonparents$wage_ratio_f, 0.75)

parent_avgage_25 = quantile(reg_stacked$avgage, 0.25)
parent_avgage_75 = quantile(reg_stacked$avgage, 0.75)

nonparent_avgage_25 = quantile(reg_stacked$avgage, 0.25)
nonparent_avgage_75 = quantile(reg_stacked$avgage, 0.75)

poorest_region = min(regionalwealth_2014$rgdppc)
richest_region = max(regionalwealth_2014$rgdppc)
```


# side quests

```{r}
# without coefficient restrictions

# deviations of characteristics compared to own sex
print("Leisure expenditure of parents (excluding sleep and personal care)")
print(make_sysfit_kable_df(parents_unres_r_within_merged, p_rows_within))

# deviations of characteristics compared to own sex
print("Leisure expenditure of parents (including sleep and personal care)")
print(make_sysfit_kable_df(parents_unres_within_merged, p_rows_within))
```

### How would households with characteristics $z_h$ be treated by a different sharing rule?
```{r}
# the 2015 sharing rule using the 2000 data
print("female resource shares, 2015 sharing rule using 2000 data, broad leisure")
summary(counterfactual_2015sharing_2000data$data$shareown_etahat_f)

# the main sharing rule (estimated using both years) using the 2000 data
print("female resource shares, main sharing rule using 2000 data, broad leisure")
summary(counterfactual_mainsharing_2000data$data$shareown_etahat_f)

# the 2000 sharing rule using 2000 data
print("female resource shares, 2000 sharing rule using 2000 data, broad leisure")
summary(shares_parents_ownsex_2000$data$shareown_etahat_f)

# the 2000 sharing rule using the 2015 data
print("female resource shares, 2000 sharing rule using 2015 data, broad leisure")
summary(counterfactual_2000sharing_2015data$data$shareown_etahat_f)

# the main sharing rule (estimated using both years) using the 2015 data
print("female resource shares, main sharing rule using 2015 data, broad leisure")
summary(counterfactual_mainsharing_2015data$data$shareown_etahat_f)

# the 2015 sharing rule using 2015 data
print("female resource shares, 2015 sharing rule using 2015 data, broad leisure")
summary(shares_parents_ownsex_2015$data$shareown_etahat_f)

# the 2000 sharing rule using all data
print("female resource shares, 2000 sharing rule using all data, broad leisure")
summary(counterfactual_2000sharing_alldata$data$shareown_etahat_f)

# the 2015 sharing rule using all data
print("female resource shares, 2015 sharing rule using all data, broad leisure")
summary(counterfactual_2015sharing_alldata$data$shareown_etahat_f)

```

```{r}
# the 2015 sharing rule using the 2000 data
print("female resource shares, 2015 sharing rule using 2000 data, narrow leisure")
summary(counterfactual_2015sharing_2000data_r$data$shareown_etahat_f)

# the main sharing rule (estimated using both years) using the 2000 data
print("female resource shares, main sharing rule using 2000 data, narrow leisure")
summary(counterfactual_mainsharing_2000data_r$data$shareown_etahat_f)

# the 2000 sharing rule using 2000 data
print("female resource shares, 2000 sharing rule using 2000 data, narrow leisure")
summary(shares_parents_ownsex_r_2000$data$shareown_etahat_f)

# the 2000 sharing rule using the 2015 data
print("female resource shares, 2000 sharing rule using 2015 data, narrow leisure")
summary(counterfactual_2000sharing_2015data_r$data$shareown_etahat_f)

# the main sharing rule (estimated using both years) using the 2015 data
print("female resource shares, main sharing rule using 2015 data, narrow leisure")
summary(counterfactual_mainsharing_2015data_r$data$shareown_etahat_f)

# the 2015 sharing rule using 2015 data
print("female resource shares, 2015 sharing rule using 2015 data, narrow leisure")
summary(shares_parents_ownsex_r_2015$data$shareown_etahat_f)

# the 2000 sharing rule using all data
print("female resource shares, 2000 sharing rule using all data, narrow leisure")
summary(counterfactual_2000sharing_alldata_r$data$shareown_etahat_f)

# the 2015 sharing rule using all data
print("female resource shares, 2015 sharing rule using all data, narrow leisure")
summary(counterfactual_2015sharing_alldata_r$data$shareown_etahat_f)

```




## papers



\[Blundell et al (2019)\]\[https://www.sfu.ca/\~pendakur/papers/BKLP16_longtitle.pdf\]

\[Hyee and Robledo (2010)\]\[https://www.nottingham.ac.uk/economics/documents/discussion-papers/10-06.pdf\]

\[Gitter and Barham (2010)\]\[https://academic.oup.com/wber/article-abstract/22/2/271/1736679?redirectedFrom=fulltext&login=false\]

Calvi (2002): Under some mild assumptions, there exists a monotonic correspondence between Pareto weights and resource shares. Consequently, resource shares may also be interpreted as a measure of intra-household bargaining power. In effect, these are more tractable measures of power, as they are invariant to unobservable cardinalisations of the utility functions.

Calvi (2020) estimates resource shares and poverty at the individual level in India and finds that women—especially older women—have lower resource shares than men. This then implies that older Indian women have much higher poverty rates than previously thought. Calvi shows that these higher poverty rates (driven by lower resource shares) among older women can explain the finding of Anderson and Ray (2010) that Indian women over the age of 45 have higher mortality rates than do Indian men (a phenomenon they call “missing women”).

\[Andrew et al (2021)\]\[https://ifs.org.uk/inequality/wp-content/uploads/2021/03/BN-The-careers-and-time-use-of-mothers-and-fathers.pdf\] show that heteronormative households, mothers are more likely to leave the labour force after having children regardless of the relative wages between mother and father before childbirth.

\[Dunbar, Lewbel, and Pendakur (2017)\]\[https://www.bankofcanada.ca/wp-content/uploads/2017/10/swp2017-45.pdf\]: Generalizes DLP in two important ways. One is the identification of resource shares without imposing the BCL or DLP restrictions on similarity of preferences across households of different composition, or across people within a household. The other is by allowing for random variation in resource shares across households, presence of unobserved distribution factors. This reflects the fact that the relative bargaining power of individual household members, or the choice of social welfare function for the household, will in general vary across households for unobserved reasons. Assume that resource shares $η_{i,h}$ are independent of the household budget. Lise and Seitz (2011), Lewbel and Pendakur (2008), Bargain and Donni (2009, 2012) and DLP all use this restriction in their identification results, and supply some theoretical arguments for it.

[Bargain, Donni, and Hentati (2022)](https://academic.oup.com/jeea/article/20/6/2468/6608951?login=true#389007362): exclusive goods, that is, goods consumed by specific individuals in the household (e.g. adult male and female clothing), and an assumption of preference stability for these goods. Precisely, our identification hinges on the additional information provided by singles’ expenditure on these exclusive goods and on the assumption that adult preferences—for the exclusive goods only—do not depend on the marital or family status. If the presence of a child implies a decrease in the demand for the exclusive good for one of the spouses, then it is concluded that the latter has transferred a fraction of her/his resources to the child and/or that the shadow price of the composite good has increased. The fraction of child costs borne by each parent is retrieved for the first time. This is relatively important information for welfare evaluation, particularly to determine how reforms that channel resources to women—for example, “wallet to purse” reforms as in Lundberg, Pollak, and Wales (1997)—also improve the condition of children. We suggest a tractable approach whereby only a system of two budget share equations–adult male and female clothing–needs to be estimated. Empirical results show that children command a reasonably large share of resources. Resource shares are relatively balanced between spouses when there are no children. However, women contribute a lot more to children’s resources—their contribution is 36%–41% larger than that of men—which makes their share lower than their husband’s in large families.
