---
title: "Showing data work (for supervisory purposes)"
format: html
editor: visual
---

Importing packages

```{r echo=FALSE, include=FALSE}
library(tidyverse)
library(systemfit)
library(ggplot2)
library(purrr)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(lmtest)
library(sandwich)
library(car)
library(patchwork)
```

Setting parameters

```{r echo=FALSE, include=FALSE}
# actlines and directories
source("UKTUS_params.R")

# functions
source("functions.R")

# get all output
source("run_SUR.R")

# for display purposes
options(scipen=1)
options(digits=4)
```

# The (baseline) framework

## Brief summary of setup in Arduini (2025):

-   An individual $i$ belongs to a household $h$. Each individual has a person type $t$. Households are in the same category $g$ if they have the same number of household members of each type. For instance, one household category would be one that contains a wife $t=1$, a husband $t=2$, and children $t=3$\*.
-   Individual-level characteristics: $\pi_{i,h}$
-   Household-level characteristics: $\zeta_{i,h}$
-   The household purchases private material goods $c_{i,h}$ and public material goods $X$. Private good $j \epsilon \Omega^c$ has price $p_j$. Public good $j \epsilon \Omega^X$ has price $r_j$.
-   Each individual has time-endowment normalised to 1.
-   Four types of time-use: private leisure $l_{i,h}$, joint leisure $jt_{i,h}$, market work $m_{i,h}$ (hours worked, does not enter utility function) and domestic work $d_{i,h}$
-   Joint leisure and domestic work are inputs to public goods $D = f^D(\mathbf{d})$ and $JT = f^{JT}(\mathbf{jt})$ enjoyed by the household Private leisure $l_{i,h}$ includes leisure activities enjoyed by an individual without the co-presence of other household members. It is therefore a private good, which enters the utility function directly.
-   $Q_h$ for the vector of public goods including both material and time-use public goods ($X_h$, $JT_h$, $D_h$) and $R_h$ for the associated price vector. $R_h$ includes wages $w_{i,h}$ of each of its members.
-   Each individual has labour income endowment equal to $w_{i,h}$. An individual’s overall endowment is $y_{i,h} = y_{i,h}^{NL} + w_{i,h}$.
-   Each member can be endowed with non-labour income $y_{i,h}^{NL}$. The household as a whole can have non-labour income $y_{h}^{NL}$.
-   The household’s endowment (or household budget) is $y_h = \sum_{i \epsilon h} y_{i,h}$.
-   Each individual i of type t living in a household category g has utility function $u_{t,g}$. There is preference heterogeneity across types and household compositions. The form of the bargaining solution of each household depends on the number and types of members, so that the sharing rule for different household categories must be estimated separately.
-   Household individuals bargain over how to divide resources. Relative bargaining power will in general depend both on (i) market variables such as prices (including wages), and (ii) distribution factors, which enter the household’s optimisation problem only indirectly through the distribution of bargaining power $\pi_{i,h}, i \epsilon h$ and $\zeta_{i,h}$. The vector of variables that affect bargaining power is $z_h$.
-   Distribution factors are variables that affect the allocation of resources within households but do not affect the tastes of individuals

## Browning, Chiappori and Weiss (2014):

-   The problem solved by any collective household, regardless of the underlying bargaining process and outside options, can be represented as an optimisation problem where the maximand is the weighted sum of the members’ utility functions. Each member’s utility function is weighed by their Pareto weight $µ_{i,h} (zh)$.
-   Maximise $\sum_{i \epsilon h} (\mu_{i,h}(z_h) u_{t,g}(c_{i,h}, X_h, l_{i,h}, JT_h, D_h))$. Each person's utility depends on individual consumption of private material goods, all public goods, private leisure, joint leisure, and domestic work.

## Arduini (2025): household's problem

-   Stage 1: The household chooses expenditure on market and time-produced (converted to wage units) public goods $R_{h}Q_{h} = \sum_{j \epsilon \Omega^X} r^j X_{h}^j + \sum_{i \epsilon h} w_{i,h} (d_{i,h} + jt_{i,h})$, where $Q_h = (X_h, JT_h, D_h)$, and how to divide the remaining household budget into individual budgets $\rho_{i,h} = (y_h − R_{h}Q_{h}) \eta_{i,h}$. The individual achieves indirect utility $v_{t,g}$ under $\rho_{i,h}$.

$$\text{max}_{\rho,X,d,jt} \sum_{i \epsilon h} \mu_{i,h}(z_h) v_{t,g}(\rho_{i,h}, X_h, JT_h, D_h), \newline
\text{subject to} \newline
y_h = \sum_{i \epsilon h} w_{i,h} (d_{i,h} + jt_{i,h}) + \sum_{j \epsilon \Omega^X} r^j X_{h}^j + \sum_{i \epsilon h} \rho_{i,h}, \newline
d_{i,h} + jt_{i,h} \leq 1, \newline
X_{h}^j, \rho_{i,h}, d_{i,h}, jt_{i,h} \geq 0, \newline
D = f^D(\mathbf{d}),
JT = f^{JT}(\mathbf{jt})$$

-   Sharing rule: determines the share of household budget net of public good expenditure assigned to each member (the individual’s resource share $\eta_{i,h}$).
-   Stage 2: Members decide how to allocate their individual budgets $\rho_{i,h}$ to private expenditure. That is, $$\text{max}_{c_{i,h},m_{i,h},l_{i,h}} u_{t,g}(c_{i,h}, X_h, l_{i,h}, JT_h, D_h), \newline
    \text{subject to} \newline
    \rho_{i,h} = \sum_{j \epsilon \Omega^c} p^j c_{i,h}^j + w_{i,h} l_{i,h}, \newline
    l_{i,h} + m_{i,h} = 1 - (d_{i,h} + jt_{i,h}), \newline
    c_{i,h},m_{i,h},l_{i,h} \geq 0$$
-   In general, second-stage demand depends on the prices of all private goods, including leisure (but not directly on the prices of public goods). Unless preferences are separable in private and public goods, it depends on public good consumption as determined in the first stage. Finally, it depends on the individual’s second-stage budget, which is their resource share multiplied by the household budget net of public good expenditure (determined in the first stage) $\rho_{i,h} = (y_h − R_{h}Q_{h}) \eta_{i,h}$.

## Arduini (2025): specification and identification

### The private assignable good

-   The private assignable good: $c^a$ is private and assignable if we observe individual-level demand or expenditure for it. That is, it is consumed by a specific person and we can clearly attribute demand or expenditure to that individual.
-   Individual i’s (of type t) second-stage demand for the assignable good is given by $c_{i,h}^a(p, w_{i,h}, Q_h, (y_h − R_{h}Q_{h}) \eta_{i,h})$

### Other assignable goods:

-   Clothing and non-market work
-   Calvi et al. (2022) find that estimates of women’s resource shares (from material assignable goods) are closely related to those women’s satisfaction with their availability of leisure time, suggesting that the collective model is right to conceptualise the same sharing rule being applied both to time-use and material goods.

### Cobb-Douglas preferences

-   What she uses: Cobb-Douglas preferences, leisure as the assignable good, an identifying assumption restricting preference heterogeneity, and (iv) an additional restriction on preference heterogeneity allowing estimation of the sharing rule from time-use data alone.
-   What she needs: The variation in assignable good expenditure with household budget is taken to be informative about how household budget is split between individual members’ second stage budgets. This relies on the strong assumption that men and women have the same preference parameter for the assignable good, or else that we already know the preference parameters of men and women, e.g. from singles data.
-   When using private leisure as the assignable good, this should not be interpreted as assuming that expenditure on any specific material good is split between members in the same manner as leisure. Men and women have a similar preference parameter for private leisure specifically. This means that, if they have the same wages and the same individual budget, then they will have the same hours of private leisure.

$$u_{t,g} = \sum_{j \epsilon \Omega^c} (\alpha_{t,g}^{cj} \ln{(c_{i,h}^j)}) + \sum_{j \epsilon \Omega^X} (\alpha_{t,g}^{Xj} \ln{(X_{h}^j)}) + \sum_{st \epsilon h} (\alpha_{t,g}^{Tst,g} \ln{(jt_{s,h})}) + \sum_{st \epsilon h} (\alpha_{t,g}^{Dst,g} \ln{(d_{s,h})})$$ 

- Resource shares are equal to Pareto weights scaled by a composite preference parameter. For instance, a significant effect of wages on the sharing rule implies that Pareto weights vary with wages
- Cobb-Duuglas preferences provides the following estimating equation:

$$w_{i,h}l_{i,h} = \alpha_{t,g}^{l} y_h (\eta_{t,g}^0 + \sum_{z} \eta_{t,g}^{z}(z_h - \bar{z}))$$


### Private leisure C-D SAP and private leisure C-D SRAT:

-   Private leisure C-D SAP: The ratio between men and women’s preference parameter for the private assignable good is stable across household compositions. Cobb-Douglas preferences over the private assignable good: homotheticity of private leisure. Everyone, regardless of their type (in a given household category) is assumed to have the same Cobb-Douglas preference parameter for private leisure: $\alpha_{t,g}^l = \alpha_{g}^l$. Preference heterogeneity across types is maintained for all other preference parameters, and across household compositions. Men and women spend the same share of their individual second-stage budget on leisure, but will split the remaining budget differently between other private goods.
-   Private leisure C-D SRAT: The SAT assumption implies $\alpha_{t,g}^l = \alpha_{t}^l$. Instead of requiring the level of preference parameters for the assignable good to be stable across household composition, she requires only that the ratio between men and women’s preference parameter for the assignable good is stable across household compositions: $\frac{\alpha_{t,g}^l}{\alpha_{st,g}^l} = \frac{\alpha_{t}^l}{\alpha_{st}^l}$
-   In particular, Cobb-Douglas preferences impose homotheticity, which may seem to strong an assumption for certain goods. It is important to note that I do not require the whole utility function to be Cobb-Douglas: more specifically, I require preferences that are Cobb-Douglas over the assignable good and some aggregate good. Therefore, I am imposing homotheticity on the assignable good, but not on other specific goods.

### Public good SAP and public good C-D SAP:

-   Public Goods SAP: Using family data probably violates assumption that total expenditure on public goods does not vary with the distribution of bargaining power in the household. This assumption does not amount to a unitary model in that the distribution of bargaining power still affects the composition of public good (and private good) expenditure; it simply does not affect the overall split between private and public expenditure. For example, female empowerment could lead to more spending on housing and less on television subscriptions.
-   Public good C-D SAP imposes that the sum of all public good preference parameters is the same for all individuals of the same household category g:

$$\sum_{j \epsilon \Omega^X} \alpha_{t,g}^{j} + \sum_{st \epsilon h} \alpha_{t,g}^{Tst,g} + \sum_{st \epsilon h} \alpha_{t,g}^{Dst,g} = a_{g}^Q$$

-   As a result of the public good SAP assumption, it is possible to estimate the sharing rule from time-use data alone, if using private leisure as the assignable good. In particular, under public good C-D SAP, the resource share and Pareto weight coincide.
-   Care should be taken extending this assumption to other contexts, especially in households with children if children are modelled as public goods, as gender differences are more likely to emerge in the desired split between private and public expenditure.

## Arduini (2025): estimating

-   Resource shares will generally depend on household budget, prices (including wages of all household members) and Pareto weights of household members: $\eta_{i,h}(y_h, p, r, w_h, µ_h)$.
-   Recall that Pareto weights $\mu_h$ are an unknown function of a vector of variables zh that determine relative bargaining power in the household, and that the collective model does not pin down a specific bargaining solution. Since $z_h$ in general includes the other variables which affect $\eta_{i,h}$, we can write $\eta_{i,h}(z_h)$
-   To estimate resource shares by linear regression, we linearly approximate them: $$\eta_{i,h} = \eta_{i,h}^0 + \sum_{z} \eta_{t,g}^{z} (z_h - \bar{z})$$
-   $\eta_{i,h}^0$ is the average resource share of a type (living in a specific household category). By definition, $\sum_{t \epsilon h} \eta_{t,g}^0 = 1$. For instance, a higher-than-average wage for the woman might increase the woman’s resource share, so that she would have a higher-than-average-for-women resource share.
-   $\eta_{t,g}^{z}$ captures the impact on sharing of a household’s characteristic $z_h$ deviating from the sample average $(z_h - \bar{z})$
-   An equation for each type of person in the form: $$w_{i,h}l_{i,h} = \alpha_{t,g}^{l} y_h (\eta_{t,g}^0 + \sum_{z} \eta_{t,g}^{z}(z_h - \bar{z})) + \epsilon_{i,h}$$
-   If $z_h$ includes all key characteristics that affect bargaining (or at least all those that are correlated with regressors) then estimates should be unbiased. The errors are negatively correlated within household (with a correlation weaker in magnitude than -1), so that a SURE estimation approach is recommended.
-   Equation for women: $$w_{i,h}l_{i,h} = \beta_{f,g}^0 y_h + \sum_{z} \beta_{f,g}^{z} y_h (z_h - \bar{z})) + \epsilon_{i,h}$$
-   Equation for men: $$w_{i,h}l_{i,h} = \beta_{m,g}^0 y_h + \sum_{z} \beta_{m,g}^{z} y_h (z_h - \bar{z})) + \epsilon_{i,h}$$
-   Then, $\beta_{t}^0 = \alpha_{t,g}^{l} \eta_{t,g}^0$ and $\beta_{t}^z = \alpha_{t,g}^{l} \eta_{t,g}^z$.
-   First, linear SURE regressions of leisure expenditure, one for men and one for women, with the restriction that $β_{f}^z + β_{m}^z = 0$. This is because $\beta_{t}^z = \alpha_{t}^{l} \eta_{t}^z$ and from private leisure C-D SAP and the definition of the linear approximation parameters, $\sum_{t \epsilon h} \beta_{t}^z = \alpha_{t}^{l} \sum_{t \epsilon h} \eta_{t}^z = 0$.
-   Next, estimate each type’s average resource share as $\hat{\eta}_{t}^0 = \frac{\hat{\beta_{t}^0}}{\hat{\beta_{f}^0} + \hat{\beta_{m}^0}}$
-   Next, estimate $\hat{\alpha_{t}^{l}} = \frac{\hat{\beta_{t}^0}}{\hat{\eta}_{t}^0}$, since $\beta_{t}^0 = \alpha_{t}^{l} \eta_{t}^0$.
-   Next, estimate $\hat{\eta_{t}^{z}} = \frac{\hat{\beta_{t}^z}}{\hat{\alpha_{t}^{l}}}$, since $\beta_{t}^z = \alpha_{t}^{l} \eta_{t}^z$.
-   Armed with the sharing rule, one can estimate individual-specific resource shares in the dataset used for estimation (the time-use dataset in this case): $\hat{\eta_{i,h}} = \hat{\eta}_{t}^0 + \sum_{z} \hat{\eta_{t}^{z}} (z_h - \bar{z})$.

# How I'm working within this framework:

-   Maybe there could be room for extension of the existing model in her paper to try and make childcare a richer aspect of the model, maybe to try and explain how/why parenthood collides with the private leisure C-D SAP and public good SAP assumptions? The idea is to see if bargaining shifts who supplies the required childcare (mother vs father), it can change which spouse gets squeezed out of leisure, and it can also change the effective private budget.
-   Some collective models of the household model children as public goods that husbands and wives may have different preferences over.
-   In these type of models, changes in distribution factors (due to changes in who income is paid to or changes in divorce laws) will change investments in children.
-   It could be that female empowerment shifts the total desired spending on public goods through a parenting/childcare channel (not just the composition of spending), then overall private/public split moves with bargaining power. Also, childcare time may eat into both private leisure of the mother and father but in different ways. If bargaining shifts who supplies the required childcare (mother vs father), it can change which spouse gets squeezed out of leisure, and it can also change the effective private budget.
-   So, childcare "production" introduces a qualitatively different public-good production technology and valuation structure. In terms of the basic theoretical framework from the lecture this would be:

child human capital = h(mother's time + father's time, parents’ material or financial investments, investments that are provided by the state)

-   This then goes into the parents' utility functions.

-   Children also create a minimum time requirement. Even for people who have childcare assistance, they can't very well never interact with their children. Coordinating childcare takes time too. Childcare can change the feasible set for being alone.

mother's time + father's time >= all childcare time (as a function of household and parent characteristics)

## Why Public Goods C-D SAP is needed

With Cobb–Douglas preferences, individuals spend a constant share of their individual private budget $\rho_{t,h}$ on private leisure:

$$w_{t,h}l_{t,h} = \alpha^l_{t,g},\rho_{t,h}$$

But $\rho_{t,h}$ is the individual’s share of the household budget net of public goods expenditure:

$$\rho_{t,h} ;=; \eta_{t,h},(y_h - R_h Q_h)$$

So the fundamental equation is

$$ w_{t,h}l_{t,h} ;=; \alpha^l_{t,g},\eta_{t,h},(y_h - R_h Q_h)
\tag{A}$$

Public-goods C-D SAP says: for each person type $t$ in category $g$, the sum of Cobb-Douglas preferemce parameters is the same.

$$\sum_{j \epsilon \Omega^X} \alpha_{t,g}^{j} + \sum_{st \epsilon h} \alpha_{t,g}^{Tst,g} + \sum_{st \epsilon h} \alpha_{t,g}^{Dst,g} = a_{g}^Q$$

This also defines the total private weight for person type $t$:
$$a^{priv}_{t,g} \equiv 1 - a_g^Q,$$
which is now the same for everyone in the household category $g$. In a Cobb–Douglas problem, the share of the full budget spent on a block equals the sum of the preference parameters on that block. So (PG) implies:

1. The household spends a constant fraction (a_g^Q) of (y_h) on public goods: $R_h Q_h = a_g^Q, y_h$
2, Therefore the private residual is a constant fraction of (y_h): $y_h - R_h Q_h = (1-a_g^Q),y_h.$

So public-goods C-D SAP gives you:

$$
(y_h - R_h Q_h) = \lambda_g y_h
\quad \text{with } \lambda_g = 1-a_g^Q.
\tag{B}
$$
Substitute (B) into (A):

$$w_{t,h}l_{t,h} = \alpha^l_{t,g} \eta_{t,h} \lambda_g y_h$$

Absorb the category constant $\lambda_g$ into $\alpha^l_{t,g}$ (or into the $\beta$’s). Define $\tilde\alpha^l_{t,g}\equiv \lambda_g \alpha^l_{t,g}$. Then:

$$ w_{t,h}l_{t,h} = \tilde\alpha^l_{t,g} \eta_{t,h} y_h
\tag{C}$$

Public Goods C-D SAP is what justifies replacing $(y_h-RQ_h)$ by resource shares multiplied by $y_h$.

Without it, we would need to find $(y_h-RQ_h)$ directly.

With (C), linearising $\eta_{t,h}$ gives:

$$\eta_{t,h} = \eta_{t,g}^0 + \sum_z \eta_{t,g}^z (z_h-\bar z)$$

Then we end up with 

$$w_{t,h}l_{t,h} = \tilde\alpha^l_{t,g} y_h\left(\eta_{t,g}^0 + 
\sum_z \eta_{t,g}^z (z_h-\bar z)\right) \times \varepsilon_{t,h}$$

So the coefficients are:

$$\beta_{t,g}^0 = \tilde\alpha_{t,g}^l \eta_{t,g}^0,\qquad
\beta_{t,g}^z = \tilde\alpha_{t,g}^l \eta_{t,g}^z.$$

In other words, in the regressions I am really trying to find $\tilde\alpha^l_{t,g} = (1-a_g^Q)\alpha^l_{t,g}$. Public-goods C-D SAP is what makes $(1-a_g^Q)$ a constant within the household category, so it can be absorbed and doesn’t mess up identification.

If public-goods C-D SAP holds, total public spending will be a fixed fraction of $y_h$ for some household so that variation in $y_h$ maps cleanly into variation in the private goods part. Then, variation in leisure expenditure in $y_h$ is informative about $\eta$ (subject to the private-leisure SAP/SRAT restriction to deal with $\alpha^l$.

If it fails, $R_h Q_h$ (and hence $y_h-RQ_h$) can vary with distribution factors $z_h$. Then the slope on $y_h(z_h - z)$ conflates changes in $\eta_{t,h}$ and changes in the size of the private residual $(y_h-RQ_h)$, so the $\beta^z$’s no longer isolate sharing-rule responses. If bargaining power shifts how much the household “wants” to devote to public goods in general, then $(y_h-RQ_h)$ moves with $z_h$ which would mean resource shares are not identified.

## Why Private Leisure C-D SAP is needed
For a household (h) in category (g), individual (t)’s resource share is approximated as

$$ \eta_{t,h} = \eta_{t,g}^0 + \sum_z \eta_{t,g}^z (z_h - \bar z)$$
For every household (h), $\sum_{t \in h} \eta_{t,h} = 1$

Plug in the approximation:

$$\sum_{t\in h}\eta_{t,g}^0 + \sum_z (z_h-\bar z)\sum_{t\in h}\eta_{t,g}^z =1$$

Since this must hold for all $z_h$, we need:

1. $$\sum_{t\in h}\eta_{t,g}^0 = 1$$

2. $$\sum_{t\in h}\eta_{t,g}^z = 0 \quad\forall z$$

With Cobb–Douglas preferences over private leisure and under the two-stage structure, individual $t$’s leisure expenditure satisfies

$$ w_{t,h} l_{t,h} = \alpha_{t,g}^l y_h (\eta_{t,g}^0 + \sum_{z} \eta_{t,g}^{z}(z_h - \bar{z})) \eta_{t,h} $$

This provides estimating equations of the form

$$ w_{t,h} l_{t,h} = \beta_{t,g}^0 y_h + \sum_z \beta_{t,g}^z , y_h (z_h-\bar z) + \varepsilon_{t,h}$$

where

$$\beta_{t,g}^0 = \alpha_{t,g}^l \eta_{t,g}^0 $$
and 

$$ \beta_{t,g}^z = \alpha_{t,g}^l \eta_{t,g}^z. $$

Private leisure C-D SAP states that for a given household category $g$, all individuals have the same Cobb–Douglas parameter for private leisure:
$$\alpha_{t,g}^l = \alpha_g^l \quad \forall t\in h$$
Under SAP, the slope coefficients become $\beta_{t,g}^z = \alpha_g^l \eta_{t,g}^z$ 

Now sum over household members:

$$ \sum_{t\in h}\beta_{t,g}^z = \alpha_g^l \sum_{t\in h}\eta_{t,g}^z$$

But from the adding-up condition above: 
$$\sum_{t\in h}\eta_{t,g}^z = 0 $$

Therefore:
$$ \sum_{t\in h}\beta_{t,g}^z = 0$$ 
Without SAP, this is

$$ \sum_t \beta_{t,g}^z = \sum_t \alpha_{t,g}^l \eta_{t,g}^z$$

which does not in general equal zero unless $\alpha_{t,g}^l$ is constant across (t).

If SAP fails, then for each $z$:

$$\beta_{f}^z = \alpha_f^l \eta_f^z,
\quad
\beta_{m}^z = \alpha_m^l \eta_m^z$$

We observe $\beta_f^z, \beta_m^z$ and two equations with four unknowns:
$$\alpha_f^l,\ \alpha_m^l,\ \eta_f^z,\ \eta_m^z$$
SRAT may be more defensible once I go to a setting with children, as parenthood plausibly shifts the level but not the ranking of leisure preferences between men and women. This could be more believable than men and women having the same level of preference for private leisure within household type. However, it is still a strong assumption as parenting likely changes preferences asymmetrically due to gender norms. Suppose motherhood reduces women’s leisure preference relative to men, that is, women put less weight on private leisure (or face guilt/constraints that behave like a lower effective preference parameter). If women have a lower taste for private leisure, they will choose less leisure even with the same resources.

## A use for weekend and weekday diaries?
We observe time diaries for one weekday and one weekend day. On a weekend versus a weekday, families are likely to perform more childcare and do less market work. In households with children, the assumptions that let us identify η from time use are systematically violated

I could go from a discussion of the collective household model and its motivation to looking at how time use is impacted by childcare for men vs women, using the weekend as a plausible source of exogenous variation. Conditional on working hours, being in a weekend setting shifts childcare in a way that is not chosen by the household and not driven by bargaining power or preferences. Households cannot choose for a weekend or weekday to happen. I have no reason to believe that households with stronger female bargaining power systematically choose different behaviour in anticipation of childcare needs only on the weekend. Female barganing power in a household likely 

## Adapting these identifying assumptions

Once children are in the picture, we could conceivably condition Private Leisure C-D SAP and Public Goods C-D SAP on observable childcare need?

## What are children?

-   Generally, people take the approach of considering children as public goods or modelling them as individual members with their own preferences (like in the Dunbar, Lewbel, and Pendakur papers using expenditure data).
-   The main limitation of private leisure is that its price is only known for those who work, for whom it coincides with the observed wage. For non-participants like children, the price of time is higher than their (unobserved) wage and we would have to estimate that in order to use time-use data to identify sharing. In the Arduini paper, she focuses on working couples and leaves this extension for future work. However, I think children are even more complicated than non-working adults.
-   The way I'm estimating the resource shares suggests that children themselves don't exist as household members in their own right, since the estimated shares add up to one for the male and female spouse/partners, and so falls under the public goods that each spouse/partner provides. I think one of the limitations of child-specific data in general, whether on time use or expenditure of goods in the strict sense, is that children may not have a lot of control over how they spend their time even without parents or other household members present (e.g. specified break times in school). So not only do children not earn wages, they don't even have a reliable way to enforce their own preferences.
-   If I had a third equation linked to a child-assignable good, I'd use $\eta_m + \eta_f + \eta_c = 1$
-   Parents presumably get enjoyment and returns out of parenting and spending time with their children. This is even the case for child-specific activities like homework and a child's private leisure, since children benefit developmentally from them and parents then benefit from this development. I think it would be really challenging to come up with a) a private assignable good and b) a shadow price of private leisure for children.
-   This line of thinking could be extended to anyone's private leisure or even anyone's clothing. If one spouse takes assignable time for private leisure and this gives them something nice to talk about with the other, or if one spouse likes it when the other wears new clothing, then neither are really private.
-   Dauphin et al (2008) and Cherchye, De Rock and Vermeulen (2008) test whether observed household demand functions are consistent with children having separate utility functions, and find evidence in favor of this hypothesis, though they do not estimate children’s shares.
-   The collective household model has been proposed partly in response to the dissatisfaction with the unitary model. The former simply assumes that the family decision process, whatever its exact nature, leads to Pareto efficient outcomes. One interesting feature of the collective model is that it does not require the specification of the bargaining process. Evidently, the generality of the efficiency assumption comes at a price: the model is not very informative about which variables, other than those in the budget constraint, may influence the decision process. In particular, the model is totally silent as to the role children may play in the process.
-   Yet every parent knows that children, even at a very early age, have their own preferences over consumption and their parents’ labour supply. In an experiment, Harbaugh et al. (2001) found that “at age 7 children’s choices about consumption goods show clear evidence of rationality, though also many inconsistencies. By age 11, choices by children... are as rational as choices by adults”. Harbaugh et al. (2003) further showed in another experiment that children display good bargaining skills as early as 7 years of age. Naturally, one may object to the inclusion of young children as decision-makers on several grounds. Within a bargaining framework, the child’s threat point could correspond to his level of well-being in a non-cooperating equilibrium with his parents. Parents could then resort to punishment to reduce their child’s bargaining power or to make it non existent.

# Finding the sharing rule

## A closer look at deviations of average household characteristics

-   Arduini (2025) uses the following example for deviation: A higher-than-average wage for the woman might increase the woman’s resource share, so that she would have a higher-than-average-for-women resource share.
-   Bargaining power may come from strength relative to one's spouse or strength relative to a spouse's options within one's own gender. Say in 2006 there is a leisure gap when women contribute more to the household. This may result in smaller increases in women’s leisure than if it were men.
-   It's worth seeing the women-vs-women/men-vs-men and women-vs-all/men-vs-all comparison (think about someone's strength relative to their spouse's other options or relative to their sponse), but estimated resource shares using any of these definitions are very similar. I've checked that I haven't been calculating them incorrectly
-   So, I'll stick with the within-sex deviations for now, following the definition in Arduini (2025).

### Using just relaxing, socialising, sport, entertainment, and hobbies:

```{r}
# deviations of characteristics compared to own sex
print(make_sysfit_kable_df(fit_res_r_within_merged, rows_within))

# deviations of characteristics across both sexes
print(make_sysfit_kable_df(fit_res_r_between_merged, rows_between))

# deviations of characteristics compared to opposite sex
print(make_sysfit_kable_df(fit_res_r_opposite_merged, rows_opposite))
```

### Including eating, sleeping, and personal care:

```{r}
print(make_sysfit_kable_df(fit_res_within_merged, rows_within))

print(make_sysfit_kable_df(fit_res_between_merged, rows_between))

print(make_sysfit_kable_df(fit_res_opposite_merged, rows_opposite))
```

-   Also, here's a bit of comparison with the unconstrained SURE regressions. I find the magnitudes of coefficients are similar across the two equations

```{r}
# without coefficient restrictions
print(make_sysfit_kable_df(fit_unres_within_merged, rows_within))
# with coefficient restrictions
print(make_sysfit_kable_df(fit_res_within_merged, rows_within))
```

-   Basically, there's more of a difference when we change the definition of leisure. When we exclude personal care and sleep, there's a lot more of an even split on average. Here are some density plots that show this. Most, but not all, estimates fall within the unit interval.

```{r echo=FALSE, fig.width=5, fig.height=8}

narrowleisure_shares = plot_share_densities(shares_fit_ownsex_r_merged$data, 
                                            dev_type = "own")

broadleisure_shares = plot_share_densities(shares_fit_ownsex_merged$data, 
                                           dev_type = "own",
                                           restrict_leisure = FALSE)

combined_plot =
  narrowleisure_shares + broadleisure_shares +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Baseline results using pooled sample",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))

print(combined_plot)
```

## How would households with characteristics $z_h$ be treated by a different sharing rule?

-   Changes over time: take income and matchings from 2000, and take coefficients from 2015. apply the betas to the observed behaviour of 2015. Would I get different counterfactual leisures?
-   For two periods, I estimate $\beta_t^0 = \alpha_t^\ell \eta_t^0,$, $\beta_t^z = \alpha_t^\ell \eta_t^z$, average resource shares $\hat\eta_t^0 = \frac{\hat\beta_t^0}{\hat\beta_m^0 + \hat\beta_f^0}$, the leisure preference parameter $\hat\alpha_t^\ell = \frac{\hat\beta_t^0}{\hat\eta_t^0}$, and each marginal sharing rule $\hat\eta_t^z = \frac{\hat\beta_t^z}{\hat\alpha_t^\ell}$.
-   So for each year (say 2000 and 2015), I'd get: $$\hat\eta_{t,h}^{(2000)} = \hat\eta_t^{0,(2000)} + \sum_z \hat\eta_t^{z,(2000)} (z_h - \bar z), \newline
    \hat\eta_{t,h}^{(2015)} = \hat\eta_t^{0,(2015)} + \sum_z \hat\eta_t^{z,(2015)} (z_h - \bar z)$$
-   Then, why not take households observed in 2000 (their $z_h$), but apply the newer sharing rule: $$\hat\eta_{i,h}^{CF} = \hat\eta_t^{0,(2015)} + \sum_z \hat\eta_t^{z,(2015)} (z_h^{2000} - \bar z^{2015})$$
-   I'd want to compare $\hat\eta_{t,h}^{CF} \quad$ and $\quad \hat\eta_{t,h}^{(2000)}$.
-   If we have $\hat\eta_{f,h}^{CF} > \hat\eta_{f,h}^{(2000)}$ for the observed 2000 household characteristics, the sharing rule has shifted in a way that allocates a larger share of resources to women.
-   When the deviations of a household's characteristics are relative to averages across both males and females, I'd propose that a higher $\hat\eta\_t\^{z,(2015)}$ than $\hat\eta\_t\^{z,(2000)}$ for some $z$ is capturing increased importance of someone’s absolute wage/education/whatever is in determining allocation, more like “my wife’s human capital carries more weight in allocation than it used to” than “my wife's position relative to my other options carries more weight than it used to”.
-   When the deviations of a household's characteristics are relative to averages of only the respective gender associated with the characteristic (for instance, the deviation of the man's wage from the wage of other men), I'd propose that a higher $\hat\eta_t^{z,(2015)}$ than $\hat\eta_t^{z,(2000)}$ for some $z$ is capturing increased importance of someone’s absolute wage/education/whatever is in determining allocation, more like how they compare to the opposing spouse's outside options became more important for intrahousehold allocation. So, it's like the other way around from the previous point.
-   For a counterfactual analysis, I can provide tables with two sets of estimated average resource shares and marginal contributions, each corresponding to the older or newer years, as well as the distribution of estimated counterfactual resource shares in the older data.

### An example

-   Suppose in 2000, the woman's wage relative to all other wages mattered less for the sharing rule, i.e. $\hat\eta_t^{w_f,(2000)}$ is small.
-   In 2015, maybe $\hat\eta_t^{w_f,(2015)}$ is larger. This would mean that household allocations are more affected by the woman's wage relative to all other wages. Women’s human capital carries more weight in allocation than it used to
-   Applying the 2015 coefficients to 2000 households: even though 2000 women earned less on average, the same relative wage difference would buy them more bargaining power under the 2015 rule.
-   The identifying assumption is: conditional on $z_h$, the estimated sharing rule captures a stable mapping from characteristics to allocation, which is the same assumption Arduini (2025) relies on.

### Comparing the two years:

-   In general, the 2015 sharing rule applied to the 2000 data provides a narrower distribution for (necessarily) both sexes. Most estimated shares still fall within the unit interval.

```{r, fig.width=5, fig.height=8}
counterfactual_shares = plot_share_densities(counterfactual_within$data, 
                                            dev_type = "own")

broadleisure_shares = plot_share_densities(shares_fit_ownsex_2000$data, 
                                           dev_type = "own")

combined_plot =
  narrowleisure_shares + broadleisure_shares +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Estimated resource shares in the 2000 UKTUS",
    subtitle = "Using 2015 sharing rule (above) and 2000 sharing rule (below)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))
      
print(combined_plot)
```

-   Here are some plots with the marginal contributions of each household characteristic to the estimated sharing rule. This one is for the merged data (2000 and 2015). Predictably, women’s resource shares are higher in households with high female hourly pay and low male hourly pay. These effects are about equal in magnitude.
-   An increase in the age gap in the direction I've specified has a negative effect on the female resource share.
-   What's interesting is that the education effects are kind of the opposite of what was found for the couples without children. Increasing male education has a small positive effect and increasing female education has a relatively larger negative effect.
-   Being in a wealthier region generally means female resource shares are smaller.
-   The only ones that are statistically significant are the wage coefficients though, and the wage effects are consistent across both years.

```{r}
marginal_impacts_merged = marginal_impacts(shares_fit_ownsex_r_merged, 
                                           sharing_est_data_merged, rows_within)

kable(marginal_impacts_merged) |>
  kable_styling("striped", full_width = F)

```

-   Now, I look at the marginal effects for both years of the UKTUS. In 2000, both female wage and female education outweigh the male wage and male education effects, so that women in households with higher household budgets have higher resource shares.

```{r}
marginal_impacts_2000 = marginal_impacts(shares_fit_ownsex_r_2000, 
                                         data_2000, rows_within)

kable(marginal_impacts_2000) |>
  kable_styling("striped", full_width = F)
```

-   We see that opposite education effect in the 2015 data, where female education has a negative impact on female resource shares and male education has a negative impact on male resource shares. The female education effect is statistically significant, unlike in the 2000 data. I don't really know why this would happen, but in a setting with children, increasing your educational attainment means delaying having a child, spending less time with your child once you do have one, or maybe both. This is true for both parents, but especially for mothers as they're the ones who actually have to give birth and stuff. Also, because takeup for educational attainment has been growing faster for females than males, the returns to female education could be lower (since females are increasing their educational attainment yet facing a comparatively unchanging group of potential matches)

```{r}
marginal_impacts_2015 = marginal_impacts(shares_fit_ownsex_r_2015, 
                                         data_2015, rows_within)
kable(marginal_impacts_2015) |>
  kable_styling("striped", full_width = F)
```

-   By the way, the wage standard deviations are very high because there are some extremely high earners in the data (or people who have high earnings but work few hours). Something else I'll do is to remove these people and see what my estimates are afterwards.
-   Just to be clear, here is the regression output for the two years separately:

```{r}
# without coefficient restrictions
print(make_sysfit_kable_df(fit_res_within_2000, rows_within))
# with coefficient restrictions
print(make_sysfit_kable_df(fit_res_within_2015, rows_within))
```


## Kind of sort of not testing assumptions:

If private leisure C-D SAP holds then the proportion of household budget spent on overall private leisure (the sum of expenditure on female private leisure and expenditure on male private leisure) should not vary with the distribution of bargaining power

```{r}
# check budget share on private leisure correlated w/ estimated resource shares

# merge with sharing_est_data
sharing_est_data_merged_shares = inner_join(shares_fit_ownsex_r_merged$data,
                                sharing_est_data_merged,
                                by=c('serial', 'is_weekend')) |>
  mutate(share_budget_leisure_r = (private_leisure_exp_r_f + 
           private_leisure_exp_r_m) / y,
         share_budget_leisure = (private_leisure_exp_f + 
           private_leisure_exp_m) / y)

# correlation of estimated resource shares (broad leisure definition)
cor.test(sharing_est_data_merged_shares$share_budget_leisure, 
         sharing_est_data_merged_shares$shareown_etahat_f)

# (narrow leisure definition)
cor.test(sharing_est_data_merged_shares$share_budget_leisure_r, 
         sharing_est_data_merged_shares$shareown_etahat_f)
```
The share of household budget spent on private leisure is slightly negatively correlated with the woman's estimated resource share, but this correlation loses significance once I include personal care and sleep to match Arduini's definition of leisure.

The SRAT assumption allows preference levels to vary between different people and different household categories, while requiring that the ratio of the parameters is constant across household compositions. There is no particular reason to suspect women’s and men’s preferences change in a different manner when transitioning from living alone to living in a couple. The assumption is stronger for households with children, since the idea that single men and women’s preferences change similarly when they have children is less realistic.

Using only time use data requires the Public Goods C-D SAP assumption that preference parameters over all public goods sum to the same quantity for men and women. With time use data, I don't have information on market goods. I don't particularly want to clean an entire expenditure survey dataset, so I think what I can do here is see (at least indirectly) whether public-good SAP plausibly fails specifically through childcare or maybe through domestic work (similarly valued at wage). I do have extensive time use data about domestic work.

## an empirical exercise


Data I should use: probably individual-level data_working_parents_2015 and data_working_parents_2000

Variables I should use:
1. `is_weekend`: weekend vs weekday diaries
2. `dgorpaf`: region the household lives in 
3. `serial`: household ID
4. `sample`: which year the data is from (2000 or 2014)
5. `total_private_leisure`: leisure expenditure of the respondent, including eating, drinking, sleeping, and other personal care
6. `total_private_leisure_r`: leisure expenditure of the respondent, excluding eating, drinking, sleeping, and other personal care
7. The respondent's time spent working that day (In estimating resource shares, earnings and working hours not measured for secondary jobs, only main job)
8. Some measure of childcare need
9. `male`: Dummy variable for sex (`== 1` when respondent is male)

# thought jail slash random archive

Lewbel and Pendakur (2008) and Bargain and Donni (2012) rely on the ‘SAT’ assumption (similarity across household types), that preferences for the assignable good are stable across household compositions, allowing for identification of the preference parameter from singles data

[Lise and Seitz (2011)](https://www.jstor.org/stable/23015857): restricting the bargaining process so that men and women with equal earnings potential share resources equally

[Browning, Chiappori and Lewbel (2013)](https://academic.oup.com/restud/article-abstract/80/4/1267/1584142): BCL

\[Dunbar, Lewbel, and Pendakur (2013)\]\[https://www.sfu.ca/\~pendakur/CollectiveChildren30-with-appendices.pdf\]: Identifies how total household resources are divided up among household members, by observing how each family member’s expenditures on a single private good like clothing varies with income and family size. DLP permits identification of resource shares of both adults and children, where children are treated as having their own utility functions and welfare. This is in contrast to most of the empirical collective household literature, where expenditures on children are modelled just as public goods in the adults’ utility functions. Identification assumes that resource shares do not vary with total expenditure. DLP and LPW use the SAP assumption (similarity across people) that women and men in the same household composition have the same preference parameter for the assignable good.

[Lechene, Pendakur, and Wolf (2022)](https://www.journals.uchicago.edu/doi/10.1086/717892): LPW shows that, building on the identification result of DLP, it is possible to estimate resource shares by linear regression

\[Blundell et al (2019)\]\[https://www.sfu.ca/\~pendakur/papers/BKLP16_longtitle.pdf\]

\[Hyee and Robledo (2010)\]\[https://www.nottingham.ac.uk/economics/documents/discussion-papers/10-06.pdf\]

\[Gitter and Barham (2010)\]\[https://academic.oup.com/wber/article-abstract/22/2/271/1736679?redirectedFrom=fulltext&login=false\]

Calvi (2002): Under some mild assumptions, there exists a monotonic correspondence between Pareto weights and resource shares. Consequently, resource shares may also be interpreted as a measure of intra-household bargaining power. In effect, these are more tractable measures of power, as they are invariant to unobservable cardinalisations of the utility functions.

Calvi (2020) estimates resource shares and poverty at the individual level in India and finds that women—especially older women—have lower resource shares than men. This then implies that older Indian women have much higher poverty rates than previously thought. Calvi shows that these higher poverty rates (driven by lower resource shares) among older women can explain the finding of Anderson and Ray (2010) that Indian women over the age of 45 have higher mortality rates than do Indian men (a phenomenon they call “missing women”).

\[Andrew et al (2021)\]\[https://ifs.org.uk/inequality/wp-content/uploads/2021/03/BN-The-careers-and-time-use-of-mothers-and-fathers.pdf\] show that heteronormative households, mothers are more likely to leave the labour force after having children regardless of the relative wages between mother and father before childbirth.

\[Dunbar, Lewbel, and Pendakur (2017)\]\[https://www.bankofcanada.ca/wp-content/uploads/2017/10/swp2017-45.pdf\]: Generalizes DLP in two important ways. One is the identification of resource shares without imposing the BCL or DLP restrictions on similarity of preferences across households of different composition, or across people within a household. The other is by allowing for random variation in resource shares across households, presence of unobserved distribution factors. This reflects the fact that the relative bargaining power of individual household members, or the choice of social welfare function for the household, will in general vary across households for unobserved reasons. Assume that resource shares $η_{i,h}$ are independent of the household budget. Lise and Seitz (2011), Lewbel and Pendakur (2008), Bargain and Donni (2009, 2012) and DLP all use this restriction in their identification results, and supply some theoretical arguments for it.

[Bargain, Donni, and Hentati (2022)](https://academic.oup.com/jeea/article/20/6/2468/6608951?login=true#389007362): exclusive goods, that is, goods consumed by specific individuals in the household (e.g. adult male and female clothing), and an assumption of preference stability for these goods. Precisely, our identification hinges on the additional information provided by singles’ expenditure on these exclusive goods and on the assumption that adult preferences—for the exclusive goods only—do not depend on the marital or family status. If the presence of a child implies a decrease in the demand for the exclusive good for one of the spouses, then it is concluded that the latter has transferred a fraction of her/his resources to the child and/or that the shadow price of the composite good has increased. The fraction of child costs borne by each parent is retrieved for the first time. This is relatively important information for welfare evaluation, particularly to determine how reforms that channel resources to women—for example, “wallet to purse” reforms as in Lundberg, Pollak, and Wales (1997)—also improve the condition of children. We suggest a tractable approach whereby only a system of two budget share equations–adult male and female clothing–needs to be estimated. Empirical results show that children command a reasonably large share of resources. Resource shares are relatively balanced between spouses when there are no children. However, women contribute a lot more to children’s resources—their contribution is 36%–41% larger than that of men—which makes their share lower than their husband’s in large families.
