---
title: "Showing my data work (for supervisory purposes)"
format: html
editor: visual
---

```{r include=FALSE}
library(tidyverse)
library(systemfit)
library(ggplot2)
library(purrr)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(lmtest)
library(sandwich)
library(car)
library(patchwork)
library(fixest)
library(patchwork)
library(marginaleffects)

# actlines and directories
source("UKTUS_params.R")

# functions
source("functions.R")

# get all output
source("run_SUR.R")

# for display purposes
options(scipen=1)
options(digits=4)
```


# Estimating the sharing rule using household time use data

```{r}

summary = c("wage_f", "wage_m", "DVAge_f", "DVAge_m", "educ_f", "educ_m")

nonparents_summary = t(sapply(nonparents_est_data_merged_weekday[summary], summarytable))
parents_summary = t(sapply(parents_est_data_merged_weekday[summary], summarytable))

print("UKTUS working heterosexual couples (parents), pooled sample")
print(as.data.frame(parents_summary))
print("UKTUS working heterosexual couples (nonparents), pooled sample")
print(as.data.frame(nonparents_summary))
```

## Baseline results

```{r}
# deviations of characteristics compared to own sex
print("Regression results (sharing rule) of parents (excluding sleep and personal care)")
print(make_sysfit_kable_df(parents_res_r_within_merged, p_rows_within))

# deviations of characteristics compared to own sex
print("Regression results (sharing rule) of parents (including sleep and personal care)")
print(make_sysfit_kable_df(parents_res_within_merged, p_rows_within))
```

```{r}
# deviations of characteristics compared to own sex
print("Regression results (sharing rule) of non-parents (excluding sleep and personal care)")
print(make_sysfit_kable_df(nonparents_res_r_within_merged, np_rows_within))

# deviations of characteristics compared to own sex
print("Regression results (sharing rule) of non-parents (including sleep and personal care)")
print(make_sysfit_kable_df(nonparents_res_within_merged, np_rows_within))
```

```{r}
print(t.test(shares_nonparents_ownsex_merged$data$shareown_etahat_f))

print(t.test(shares_parents_ownsex_merged$data$shareown_etahat_f))

print(t.test(shares_parents_ownsex_merged$data$shareown_etahat_f,
             shares_nonparents_ownsex_merged$data$shareown_etahat_f))
```

```{r}
print(t.test(shares_nonparents_ownsex_r_merged$data$shareown_etahat_f))

print(t.test(shares_parents_ownsex_r_merged$data$shareown_etahat_f))

print(t.test(shares_parents_ownsex_r_merged$data$shareown_etahat_f,
             shares_nonparents_ownsex_r_merged$data$shareown_etahat_f))
```

```{r fig.width=10, fig.height=8}

narrowleisure_shares_p = plot_share_densities(parents_est_data_merged_shares |> filter(is_weekend == 1), 
                                            dev_type = "own",
                                            restrict_leisure = TRUE)

broadleisure_shares_p = plot_share_densities(parents_est_data_merged_shares |> filter(is_weekend == 1), 
                                           dev_type = "own",
                                           restrict_leisure = FALSE)

combined_plot_p =
  narrowleisure_shares_p + broadleisure_shares_p +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Estimated resource shares (parents)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))

narrowleisure_shares_np = plot_share_densities(nonparents_est_data_merged_shares, 
                                            dev_type = "own")

broadleisure_shares_np = plot_share_densities(nonparents_est_data_merged_shares, 
                                           dev_type = "own",
                                           restrict_leisure = FALSE)

combined_plot_np =
  narrowleisure_shares_np + broadleisure_shares_np +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Estimated resource shares (non-parents)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))

combined_plot_baseline = wrap_elements(combined_plot_p) | wrap_elements(combined_plot_np)

print(combined_plot_baseline)

```

```{r}
marginal_impacts_parents = marginal_impacts(shares_parents_ownsex_merged, 
                                           parents_est_data_merged, p_rows_within)

marginal_impacts_nonparents = marginal_impacts(shares_nonparents_ownsex_merged, 
                                           nonparents_est_data_merged, np_rows_within, 
                                           data_type = "nonparents")

marginal_impacts_parents_r = marginal_impacts(shares_parents_ownsex_r_merged, 
                                           parents_est_data_merged, p_rows_within)

marginal_impacts_nonparents_r = marginal_impacts(shares_nonparents_ownsex_r_merged, 
                                           nonparents_est_data_merged, np_rows_within, 
                                           data_type = "nonparents")


print("Interpretation of regression coefficients (excluding personal care and sleep, parents)")
print(marginal_impacts_parents_r)

print("Interpretation of regression coefficients (including personal care and sleep, parents)")
print(marginal_impacts_parents)

print("Interpretation of regression coefficients (excluding personal care and sleep, non-parents)")
print(marginal_impacts_nonparents_r)

print("Interpretation of regression coefficients (including personal care and sleep, non-parents)")
print(marginal_impacts_nonparents)


```
## Comparing narrow and broad leisure definitions:
```{r}
compare_shares = parents_est_data_merged_shares |>
  mutate(
    diff_share_f = shareown_etahat_f - shareown_etahat_r_f,
    abs_diff_share_f = abs(diff_share_f)
  )


print(compare_shares |> summarise(
    mean_diff = mean(diff_share_f),
    mean_abs_diff = mean(abs_diff_share_f),
    p25 = quantile(abs_diff_share_f, 0.25),
    p50 = quantile(abs_diff_share_f, 0.50),
    p75 = quantile(abs_diff_share_f, 0.75),
    corr = cor(shareown_etahat_f, shareown_etahat_r_f)
  ))

print(compare_shares |>
  mutate(
    q_narrow = ntile(shareown_etahat_r_f, 4),
    q_broad  = ntile(shareown_etahat_f, 4)
  ) |>
  count(q_narrow, q_broad))
```

## Comparing estimates across both UKTUS survey years:

```{r fig.width=10, fig.height=8}

narrowleisure_shares_2000 = plot_share_densities(parents_est_data_merged_shares |> filter(sample == "2000", is_weekend == 1), 
                                            dev_type = "own",
                                            restrict_leisure = TRUE)

broadleisure_shares_2000 = plot_share_densities(parents_est_data_merged_shares |> filter(sample == "2000", is_weekend == 1), 
                                           dev_type = "own",
                                           restrict_leisure = FALSE)

narrowleisure_shares_2015 = plot_share_densities(parents_est_data_merged_shares |> filter(sample == "2015", is_weekend == 1), 
                                            dev_type = "own",
                                            restrict_leisure = TRUE)

broadleisure_shares_2015 = plot_share_densities(parents_est_data_merged_shares |> filter(sample == "2015", is_weekend == 1),  
                                           dev_type = "own",
                                           restrict_leisure = FALSE)

combined_plot_2000 =
  narrowleisure_shares_2000 + broadleisure_shares_2000 +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Estimated resource shares in 2000 (parents)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))

combined_plot_2015 =
  narrowleisure_shares_2015 + broadleisure_shares_2015 +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Estimated resource shares in 2015 (parents)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))

combined_plot_years = wrap_elements(combined_plot_2000) | wrap_elements(combined_plot_2015)

print(combined_plot_years)
```
```{r fig.width=10, fig.height=8}

narrowleisure_shares_2000 = plot_share_densities(nonparents_est_data_merged_shares |> filter(sample == "2000"), 
                                            dev_type = "own",
                                            restrict_leisure = TRUE)

broadleisure_shares_2000 = plot_share_densities(nonparents_est_data_merged_shares |> filter(sample == "2000"),
                                           dev_type = "own",
                                           restrict_leisure = FALSE)

narrowleisure_shares_2015 = plot_share_densities(nonparents_est_data_merged_shares |> filter(sample == "2015"),
                                            dev_type = "own",
                                            restrict_leisure = TRUE)

broadleisure_shares_2015 = plot_share_densities(nonparents_est_data_merged_shares |> filter(sample == "2015"),
                                           dev_type = "own",
                                           restrict_leisure = FALSE)

combined_plot_2000 =
  narrowleisure_shares_2000 + broadleisure_shares_2000 +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Estimated resource shares in 2000 (non-parents)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))

combined_plot_2015 =
  narrowleisure_shares_2015 + broadleisure_shares_2015 +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Estimated resource shares in 2015 (non-parents)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))

combined_plot_years = wrap_elements(combined_plot_2000) | wrap_elements(combined_plot_2015)

print(combined_plot_years)
```

## Kind of testing assumptions:

- "If private leisure C-D SAP holds then the proportion of household budget spent on overall private leisure (the sum of expenditure on female private leisure and expenditure on male private leisure) should not vary with the distribution of bargaining power"
```{r}
# check budget share on private leisure correlated w/ estimated resource shares

print("test of correlations: estimated resource shares and total private leisure expenditure (including personal care and sleep")
# correlation of estimated resource shares (broad leisure definition)
cor.test(parents_est_data_merged_shares$share_budget_leisure, 
         parents_est_data_merged_shares$shareown_etahat_f)

print("test of correlations: estimated resource shares and total private leisure expenditure (excluding personal care and sleep")
# (narrow leisure definition)
cor.test(parents_est_data_merged_shares$share_budget_leisure_r, 
         parents_est_data_merged_shares$shareown_etahat_r_f)
```

## How private leisure responds to childcare need (constraints?)

- could be good to add a section about what preference parameter restriction on child human capital production youd need to identify the effect, and whether it can be tested with time use data only
- Individual private leisure expenditure responds to the individual budget generated by the sharing rule. Resource shares are based on household characteristics that don't usually change from day to day. Now, I want to see how this maps into state-dependent adjustments in leisure and childcare. Children of all ages mechanically spend more time outside the home during weekdays since most schools and childcare services are closed on weekends. This means that weekends shift the feasible time use set as children require more time from their parents. This extra childcare requirement must come out of somebody’s time through lower private leisure or lower market work. I have one weekday and one weekend activity diary for each member of each household, so I can see who is absorbing this extra childcare need. I propose using variation in time use generated through weekdays and weekends to see how households with different resource shares absorb short-run increases in childcare need.
- The weekend/weekday contrast shows how a fixed household bargaining arrangement responds to a tighter childcare constraint
- Earlier in this document, I wrote about the following relationships between leisure and childcare:
- Through one channel: If mothers already have low weekday private leisure, they may be at (or near) a constraint where further childcare would imply very high marginal disutility. The mother’s marginal utility loss from reducing $l_f$ further is very steep so the father’s leisure drops more when $\bar{h}$ rises. 
- Through another channel: If, say, female bargaining power is relatively high in some household, then when $\bar{h}$ rises the model predicts that fathers will absorb more of the childcare adjustment because the household places more weight on mother’s utility
- Through another channel: The male and female preference parameters for a child's human capital and how productive they are in providing childcare may differ
- The issue with these regressions is that my data cannot fully separate these channels, but it can test whether weekend adjustment is stronger in households with higher estimated female resource shares. What I can say is that households with different estimated female resource shares allocate childcare differently
- The estimated resource shares determine how household resources are allocated across members, but they do not by themselves identify whether lower individual expenditure on the private assignable good is driven by bargaining power, parent-specific preferences for child human capital, parent-specific productivity in childcare, or gendered expectations determining preference parameters.
- See the mean estimated female resource shares among households with above/below-median male/female wages, above/below-median male/female educational attainment, above/below-median number of children, above/below-median childcare time

See if women with lower or higher relative bargaining are already providing a lot on the weekdays.
Resource shares and child characteristics:
```{r}
# use parents_est_data_merged_shares find average of shareown_etahat_f at 
# above or below-median wage_f, educ_f, total_childcare_f 

wage_split = parents_est_data_merged_shares |>
  mutate(
    wage_f_high = wage_f > median(wage_f, na.rm = TRUE),
    wage_m_high = wage_m > median(wage_m, na.rm = TRUE)) |>
  group_by(wage_f_high, wage_m_high) |>
  summarise(
    mean_share_f = mean(shareown_etahat_f, na.rm = TRUE),
    .groups = "drop") |>
  mutate(
    wage_f_high = if_else(wage_f_high, "Above-median female wage", 
                          "Below-median female wage"),
    wage_m_high = if_else(wage_m_high, "Above-median male wage", 
                          "Below-median male wage")) |>
  pivot_wider(
    names_from = wage_m_high,
    values_from = mean_share_f)

child_split = parents_est_data_merged_shares |>
  mutate(
    wage_f_high = wage_f > median(wage_f, na.rm = TRUE),
    young_child = kid_age_min < 10) |>
  group_by(wage_f_high, young_child) |>
  summarise(
    mean_share_f = mean(shareown_etahat_f, na.rm = TRUE),
    .groups = "drop") |>
  mutate(
    wage_f_high = if_else(wage_f_high, "Above-median female wage", 
                          "Below-median female wage"),
    young_child = if_else(young_child, "Child under 10 in household", 
                          "No child under 10")) |>
  pivot_wider(
    names_from = young_child,
    values_from = mean_share_f)

num_split = parents_est_data_merged_shares |>
  mutate(
    wage_f_high = wage_f > median(wage_f, na.rm = TRUE),
    high_kids = num_kids_total > median(num_kids_total, na.rm=TRUE)) |>
  group_by(wage_f_high, high_kids) |>
  summarise(
    mean_share_f = mean(shareown_etahat_f, na.rm = TRUE),
    .groups = "drop") |>
  mutate(
    wage_f_high = if_else(wage_f_high, "Above-median female wage", 
                          "Below-median female wage"),
    high_kids = if_else(high_kids, "Above-median number of children", 
                        "Below-median number of children")) |>
  pivot_wider(
    names_from = high_kids,
    values_from = mean_share_f)

print(wage_split)
print(child_split)
print(num_split)
```

```{r}
# use parents_est_data_merged_shares find average of shareown_etahat_r_f at 
# above or below-median wage_f, educ_f, total_childcare_f 

wage_split_r = parents_est_data_merged_shares |>
  mutate(
    wage_f_high = wage_f > median(wage_f, na.rm = TRUE),
    wage_m_high = wage_m > median(wage_m, na.rm = TRUE)) |>
  group_by(wage_f_high, wage_m_high) |>
  summarise(
    mean_share_f = mean(shareown_etahat_r_f, na.rm = TRUE),
    .groups = "drop") |>
  mutate(
    wage_f_high = if_else(wage_f_high, "Above-median female wage", 
                          "Below-median female wage"),
    wage_m_high = if_else(wage_m_high, "Above-median male wage", 
                          "Below-median male wage")) |>
  pivot_wider(
    names_from = wage_m_high,
    values_from = mean_share_f)

child_split_r = parents_est_data_merged_shares |>
  mutate(
    wage_f_high = wage_f > median(wage_f, na.rm = TRUE),
    young_child = kid_age_min < 10) |>
  group_by(wage_f_high, young_child) |>
  summarise(
    mean_share_f = mean(shareown_etahat_r_f, na.rm = TRUE),
    .groups = "drop") |>
  mutate(
    wage_f_high = if_else(wage_f_high, "Above-median female wage", 
                          "Below-median female wage"),
    young_child = if_else(young_child, "Child under 10 in household", 
                          "No child under 10")) |>
  pivot_wider(
    names_from = young_child,
    values_from = mean_share_f)

num_split_r = parents_est_data_merged_shares |>
  mutate(
    wage_f_high = wage_f > median(wage_f, na.rm = TRUE),
    high_kids = num_kids_total > median(num_kids_total, na.rm=TRUE)) |>
  group_by(wage_f_high, high_kids) |>
  summarise(
    mean_share_f = mean(shareown_etahat_r_f, na.rm = TRUE),
    .groups = "drop") |>
  mutate(
    wage_f_high = if_else(wage_f_high, "Above-median female wage", 
                          "Below-median female wage"),
    high_kids = if_else(high_kids, "Above-median number of children", 
                        "Below-median number of children")) |>
  pivot_wider(
    names_from = high_kids,
    values_from = mean_share_f)

print(wage_split_r)
print(child_split_r)
print(num_split_r)
```

Table of time use data:
```{r}
# compute mean time use within each quartile
table_quartiles = timeuse_long |>
  group_by(share_quartile, day, sex, activity) |>
  summarise(
    mean_hours = mean(hours, na.rm = TRUE),
    sd_hours = sd(hours, na.rm = TRUE),
    n = n(),
    .groups = "drop") |>
  mutate(
    mean_hours = round(mean_hours, 2),
    sd_hours = round(sd_hours, 2)) |>
  select(day, sex, activity, share_quartile, mean_hours) |>
  pivot_wider(
    names_from = share_quartile,
    values_from = mean_hours) |>
  arrange(day, sex, activity) |>
  filter(activity %in% c(
    "Childcare",
    "Private leisure",
    "Private leisure (excluding sleep and personal care)",
    "Market work",
    "Domestic work excluding childcare"))

print(table_quartiles)
```

```{r}
# compute mean time use within each quartile restricted leisure shares
table_quartiles_r = timeuse_long |>
  group_by(share_quartile_r, day, sex, activity) |>
  summarise(
    mean_hours = mean(hours, na.rm = TRUE),
    sd_hours = sd(hours, na.rm = TRUE),
    n = n(),
    .groups = "drop") |>
  mutate(
    mean_hours = round(mean_hours, 2),
    sd_hours = round(sd_hours, 2)) |>
  select(day, sex, activity, share_quartile_r, mean_hours) |>
  pivot_wider(
    names_from = share_quartile_r,
    values_from = mean_hours) |>
  arrange(day, sex, activity) |>
  filter(activity %in% c(
    "Childcare",
    "Private leisure",
    "Private leisure (excluding sleep and personal care)",
    "Market work",
    "Domestic work excluding childcare"))

print(table_quartiles_r)
```

```{r}
men_weekday = reg_stacked |> filter(is_weekend == 0, male == 1)
men_weekend = reg_stacked |> filter(is_weekend == 1, male == 1)
women_weekday = reg_stacked |> filter(is_weekend == 0, male == 0)
women_weekend = reg_stacked |> filter(is_weekend == 1, male == 0)

summary(men_weekday$total_childcare)
print(sd(men_weekday$total_childcare))
summary(women_weekday$total_childcare)
print(sd(women_weekday$total_childcare))

summary(men_weekend$total_childcare)
print(sd(men_weekend$total_childcare))
summary(women_weekend$total_childcare)
print(sd(women_weekend$total_childcare))

summary(men_weekday$total_private_leisure)
print(sd(men_weekday$total_private_leisure))
summary(women_weekday$total_private_leisure)
print(sd(women_weekday$total_private_leisure))

summary(men_weekend$total_private_leisure)
print(sd(men_weekend$total_private_leisure))
summary(women_weekend$total_private_leisure)
print(sd(women_weekend$total_private_leisure))

summary(men_weekday$total_work)
print(sd(men_weekday$total_work))
summary(women_weekday$total_work)
print(sd(women_weekday$total_work))

summary(men_weekend$total_work)
print(sd(men_weekend$total_work))
summary(women_weekend$total_work)
print(sd(women_weekend$total_work))
```

```{r}
etable(feols(total_work ~ is_weekend * male * shareown_etahat_r_f
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))

etable(feols(total_childcare ~ is_weekend * male * shareown_etahat_r_f
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))

etable(feols(total_private_leisure ~ is_weekend * male * shareown_etahat_r_f
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))

etable(feols(total_private_leisure_r ~ is_weekend * male * shareown_etahat_r_f
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))

etable(feols(total_otherdomestic ~ is_weekend * male * shareown_etahat_r_f
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))
```

```{r}
etable(feols(total_work ~ is_weekend * male * shareown_etahat_f
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))

etable(feols(total_childcare ~ is_weekend * male * shareown_etahat_f
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))

etable(feols(total_private_leisure ~ is_weekend * male * shareown_etahat_f
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))

etable(feols(total_private_leisure_r ~ is_weekend * male * shareown_etahat_f
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))

etable(feols(total_otherdomestic ~ is_weekend * male * shareown_etahat_f
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))
```
```{r}
etable(feols(total_childcare ~ is_weekend * male * over_75p_fem_share_r
      | sample + dgorpaf,
      data = reg_stacked,
      cluster = "serial"))
```

```{r}
# example
eq_m = delta_total_private_leisure_m ~ 0 + shareown_etahat_f + num_kids_total +
  kid_age_min + avgage + agegap_m + has_childcare_help + has_otherdomestic_help + factor(sample) + factor(dgorpaf)
eq_f = delta_total_private_leisure_f ~ 0 + shareown_etahat_f + num_kids_total +
  kid_age_min + avgage + agegap_m + has_childcare_help + has_otherdomestic_help + factor(sample) + factor(dgorpaf)

summary(systemfit(list(mother = eq_m, father = eq_f), method = "SUR", data = delta_data))
```


# side quests

```{r}
# without coefficient restrictions

# deviations of characteristics compared to own sex
print("Leisure expenditure of parents (excluding sleep and personal care)")
print(make_sysfit_kable_df(parents_unres_r_within_merged, p_rows_within))

# deviations of characteristics compared to own sex
print("Leisure expenditure of parents (including sleep and personal care)")
print(make_sysfit_kable_df(parents_unres_within_merged, p_rows_within))
```

### How would households with characteristics $z_h$ be treated by a different sharing rule?
```{r}
# the 2015 sharing rule using the 2000 data
print("female resource shares, 2015 sharing rule using 2000 data, broad leisure")
summary(counterfactual_2015sharing_2000data$data$shareown_etahat_f)

# the main sharing rule (estimated using both years) using the 2000 data
print("female resource shares, main sharing rule using 2000 data, broad leisure")
summary(counterfactual_mainsharing_2000data$data$shareown_etahat_f)

# the 2000 sharing rule using 2000 data
print("female resource shares, 2000 sharing rule using 2000 data, broad leisure")
summary(shares_parents_ownsex_2000$data$shareown_etahat_f)

# the 2000 sharing rule using the 2015 data
print("female resource shares, 2000 sharing rule using 2015 data, broad leisure")
summary(counterfactual_2000sharing_2015data$data$shareown_etahat_f)

# the main sharing rule (estimated using both years) using the 2015 data
print("female resource shares, main sharing rule using 2015 data, broad leisure")
summary(counterfactual_mainsharing_2015data$data$shareown_etahat_f)

# the 2015 sharing rule using 2015 data
print("female resource shares, 2015 sharing rule using 2015 data, broad leisure")
summary(shares_parents_ownsex_2015$data$shareown_etahat_f)

# the 2000 sharing rule using all data
print("female resource shares, 2000 sharing rule using all data, broad leisure")
summary(counterfactual_2000sharing_alldata$data$shareown_etahat_f)

# the 2015 sharing rule using all data
print("female resource shares, 2015 sharing rule using all data, broad leisure")
summary(counterfactual_2015sharing_alldata$data$shareown_etahat_f)

```

```{r}
# the 2015 sharing rule using the 2000 data
print("female resource shares, 2015 sharing rule using 2000 data, narrow leisure")
summary(counterfactual_2015sharing_2000data_r$data$shareown_etahat_f)

# the main sharing rule (estimated using both years) using the 2000 data
print("female resource shares, main sharing rule using 2000 data, narrow leisure")
summary(counterfactual_mainsharing_2000data_r$data$shareown_etahat_f)

# the 2000 sharing rule using 2000 data
print("female resource shares, 2000 sharing rule using 2000 data, narrow leisure")
summary(shares_parents_ownsex_r_2000$data$shareown_etahat_f)

# the 2000 sharing rule using the 2015 data
print("female resource shares, 2000 sharing rule using 2015 data, narrow leisure")
summary(counterfactual_2000sharing_2015data_r$data$shareown_etahat_f)

# the main sharing rule (estimated using both years) using the 2015 data
print("female resource shares, main sharing rule using 2015 data, narrow leisure")
summary(counterfactual_mainsharing_2015data_r$data$shareown_etahat_f)

# the 2015 sharing rule using 2015 data
print("female resource shares, 2015 sharing rule using 2015 data, narrow leisure")
summary(shares_parents_ownsex_r_2015$data$shareown_etahat_f)

# the 2000 sharing rule using all data
print("female resource shares, 2000 sharing rule using all data, narrow leisure")
summary(counterfactual_2000sharing_alldata_r$data$shareown_etahat_f)

# the 2015 sharing rule using all data
print("female resource shares, 2015 sharing rule using all data, narrow leisure")
summary(counterfactual_2015sharing_alldata_r$data$shareown_etahat_f)

```

Sharing rules for each year separately:
```{r}
# deviations of characteristics compared to own sex
print("Regression results (sharing rule) of parents (including sleep and personal care, 2000)")
print(make_sysfit_kable_df(parents_res_within_2000, p_rows_within))

# deviations of characteristics compared to own sex
print("Regression results (sharing rule) of parents (including sleep and personal care, 2015)")
print(make_sysfit_kable_df(parents_res_within_2015, p_rows_within))
```
```{r}
# deviations of characteristics compared to own sex
print("Regression results (sharing rule) of parents (excluding sleep and personal care, 2000)")
print(make_sysfit_kable_df(parents_res_r_within_2000, p_rows_within))

# deviations of characteristics compared to own sex
print("Regression results (sharing rule) of parents (excluding sleep and personal care, 2015)")
print(make_sysfit_kable_df(parents_res_r_within_2015, p_rows_within))
```

```{r}
marginal_impacts_counterfactual_2000 = marginal_impacts(counterfactual_2000sharing_alldata, 
                                           parents_est_data_merged, p_rows_within)

marginal_impacts_counterfactual_2015 = marginal_impacts(counterfactual_2015sharing_alldata, 
                                           parents_est_data_merged, np_rows_within)

print("Interpretation of regression coefficients (combined sample, 2000 sharing rule, broad leisure)")
print(marginal_impacts_counterfactual_2000)

print("Interpretation of regression coefficients (combined sample, 2015 sharing rule, broad leisure)")
print(marginal_impacts_counterfactual_2015)
```
```{r}
marginal_impacts_counterfactual_2000 = marginal_impacts(counterfactual_2000sharing_alldata_r, 
                                           parents_est_data_merged, p_rows_within)

marginal_impacts_counterfactual_2015 = marginal_impacts(counterfactual_2015sharing_alldata_r, 
                                           parents_est_data_merged, np_rows_within)

print("Interpretation of regression coefficients (combined sample, 2000 sharing rule, narrow leisure)")
print(marginal_impacts_counterfactual_2000)

print("Interpretation of regression coefficients (combined sample, 2015 sharing rule, narrow leisure)")
print(marginal_impacts_counterfactual_2015)
```

## papers

Lewbel and Pendakur (2008) and Bargain and Donni (2012) rely on the ‘SAT’ assumption (similarity across household types), that preferences for the assignable good are stable across household compositions, allowing for identification of the preference parameter from singles data

[Lise and Seitz (2011)](https://www.jstor.org/stable/23015857): restricting the bargaining process so that men and women with equal earnings potential share resources equally

[Browning, Chiappori and Lewbel (2013)](https://academic.oup.com/restud/article-abstract/80/4/1267/1584142): BCL

\[Dunbar, Lewbel, and Pendakur (2013)\]\[https://www.sfu.ca/\~pendakur/CollectiveChildren30-with-appendices.pdf\]: Identifies how total household resources are divided up among household members, by observing how each family member’s expenditures on a single private good like clothing varies with income and family size. DLP permits identification of resource shares of both adults and children, where children are treated as having their own utility functions and welfare. This is in contrast to most of the empirical collective household literature, where expenditures on children are modelled just as public goods in the adults’ utility functions. Identification assumes that resource shares do not vary with total expenditure. DLP and LPW use the SAP assumption (similarity across people) that women and men in the same household composition have the same preference parameter for the assignable good.

[Lechene, Pendakur, and Wolf (2022)](https://www.journals.uchicago.edu/doi/10.1086/717892): LPW shows that, building on the identification result of DLP, it is possible to estimate resource shares by linear regression

\[Blundell et al (2019)\]\[https://www.sfu.ca/\~pendakur/papers/BKLP16_longtitle.pdf\]

\[Hyee and Robledo (2010)\]\[https://www.nottingham.ac.uk/economics/documents/discussion-papers/10-06.pdf\]

\[Gitter and Barham (2010)\]\[https://academic.oup.com/wber/article-abstract/22/2/271/1736679?redirectedFrom=fulltext&login=false\]

Calvi (2002): Under some mild assumptions, there exists a monotonic correspondence between Pareto weights and resource shares. Consequently, resource shares may also be interpreted as a measure of intra-household bargaining power. In effect, these are more tractable measures of power, as they are invariant to unobservable cardinalisations of the utility functions.

Calvi (2020) estimates resource shares and poverty at the individual level in India and finds that women—especially older women—have lower resource shares than men. This then implies that older Indian women have much higher poverty rates than previously thought. Calvi shows that these higher poverty rates (driven by lower resource shares) among older women can explain the finding of Anderson and Ray (2010) that Indian women over the age of 45 have higher mortality rates than do Indian men (a phenomenon they call “missing women”).

\[Andrew et al (2021)\]\[https://ifs.org.uk/inequality/wp-content/uploads/2021/03/BN-The-careers-and-time-use-of-mothers-and-fathers.pdf\] show that heteronormative households, mothers are more likely to leave the labour force after having children regardless of the relative wages between mother and father before childbirth.

\[Dunbar, Lewbel, and Pendakur (2017)\]\[https://www.bankofcanada.ca/wp-content/uploads/2017/10/swp2017-45.pdf\]: Generalizes DLP in two important ways. One is the identification of resource shares without imposing the BCL or DLP restrictions on similarity of preferences across households of different composition, or across people within a household. The other is by allowing for random variation in resource shares across households, presence of unobserved distribution factors. This reflects the fact that the relative bargaining power of individual household members, or the choice of social welfare function for the household, will in general vary across households for unobserved reasons. Assume that resource shares $η_{i,h}$ are independent of the household budget. Lise and Seitz (2011), Lewbel and Pendakur (2008), Bargain and Donni (2009, 2012) and DLP all use this restriction in their identification results, and supply some theoretical arguments for it.

[Bargain, Donni, and Hentati (2022)](https://academic.oup.com/jeea/article/20/6/2468/6608951?login=true#389007362): exclusive goods, that is, goods consumed by specific individuals in the household (e.g. adult male and female clothing), and an assumption of preference stability for these goods. Precisely, our identification hinges on the additional information provided by singles’ expenditure on these exclusive goods and on the assumption that adult preferences—for the exclusive goods only—do not depend on the marital or family status. If the presence of a child implies a decrease in the demand for the exclusive good for one of the spouses, then it is concluded that the latter has transferred a fraction of her/his resources to the child and/or that the shadow price of the composite good has increased. The fraction of child costs borne by each parent is retrieved for the first time. This is relatively important information for welfare evaluation, particularly to determine how reforms that channel resources to women—for example, “wallet to purse” reforms as in Lundberg, Pollak, and Wales (1997)—also improve the condition of children. We suggest a tractable approach whereby only a system of two budget share equations–adult male and female clothing–needs to be estimated. Empirical results show that children command a reasonably large share of resources. Resource shares are relatively balanced between spouses when there are no children. However, women contribute a lot more to children’s resources—their contribution is 36%–41% larger than that of men—which makes their share lower than their husband’s in large families.
