---
title: "Showing data work (for supervisory purposes)"
format: html
editor: visual
---

```{r echo=FALSE, include=FALSE}
library(tidyverse)
library(systemfit)
library(ggplot2)
library(purrr)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(lmtest)
library(sandwich)
library(car)
library(patchwork)
library(fixest)

# actlines and directories
source("UKTUS_params.R")

# functions
source("functions.R")

# get all output
source("run_SUR.R")

# for display purposes
options(scipen=1)
options(digits=4)
```


# My work so far

## Sharing in the collective household
In Development I, we discussed Rossella Calvi's 2020 paper on individual poverty rates in India. She proposes that unequal sharing of resources within the household can lead to higher mortality rates for older women. She estimates these resource shares structurally using some preference heterogeneity restrictions over private assignable goods in a collective household model, which definitely represents the spirit of much of the collective households literature in this area. In some later work, she finds that these resource shares are correlated with women's qualitative satisfaction with their availability of private leisure time.
I'm interested in the roles that childcare and gender norms may play in this structural setting. It could be that female empowerment shifts the total desired spending on public goods through a parenting/childcare channel (not just the composition of spending), then overall private/public split moves with bargaining power. Or, if one parent is already providing as much childcare as they can, the other parent must step in to provide whatever additional childcare is needed. If parents are working less to provide more childcare, this could then lead them to contribute less financially through realised earnings and then also have lower relative bargaining power, which lowers their leisure as well. It's possible that parents allocate resources or spend their time differently depending on their own preferences for sons or daughters (specifically, the Dahl and Moretti paper about shotgun marriages or whatever that was). 
Thinking about children as a public good: Tt's likely reasonable to say that parents intrinsically enjoy the process of caring for children as well as any resulting accomplishments and human capital produced by that parenting, although parents generally don't have to have the same preferences around childcare. In particular, mothers and fathers may differ in how much they believe they are entitled to a certain amount of private time, how much they are expected (by themselves or by those around them) to sacrifice leisure for the sake of childcare. I think it would be reasonable to argue as well that you can't outsource parenting to the extent that you can with other domestic production. 
For instance, suppose that parents value their consumption and leisure. Assume they also attach a value to their child which could depend on their child’s level of human capital $H$. There can be many households, but dropping most subscripts/indexing, we have adults $i \in \{f, m\}$ with individual utility functions $u_f$ and $u_m$. These two people have individual characteristics $\pi_f$ and $\pi_m$. The household's characteristics are $\zeta$. $f$ and $m$ consume $c_m$ or $c_f$ at price $p$ and decide spending on a public good $X$ at price $r$. Then, $f$ and $m$ will choose to spend their time on private leisure $l_f$/$l_m$ and parenting time $h_f$/$h_m$.
Parents are constrained in both their time and material investments. With their time and public good, the household produces $H = H(h_f, h_m, X, Z)$, where $Z$ is something like government/exogenous spending on children.  $f$ and $m$ have wages $w_f$ and $w_m$ with market work $m_i$ and non-labour income $y^{NL}$, such that $y_f = y^{NL}_f + w_f$ and $y_m = y^{NL}_m + w_m$. The household has time endowment normalised to 1. The household's total budget is $y = y_f + y_m = y^{NL}_f + w_f + y^{NL}_m + w_m$. 
Note that these are defined as "full" labour endowments and household budgets, as if both people in the household worked through their entire time endowment. This is because time use is endogenous, including choices for market work. This also values any non-market time use at the wage $w_i$. The household's true income is $ y^{NL}_f + w_f m_f + y^{NL}_m + w_m m_m = rX + p(c_f + c_m)$ and their time constraints are $h_i + l_i + m_i = 1$.
With Cobb-Douglas preferences and domestic production, I could substitute the production straight into the utility representation and get something like 

$$
U_i = \alpha_{i}^{c} \ln{(c_{i})} + \alpha_{i}^{l} \ln(l_{i}) + \alpha_{i}^{H}( \beta \ln(h_m) + \newline 
\beta^f \ln(h_f) + \beta_X \ln(X) + \beta^Z \ln(Z)) \newline 
= \alpha_{i}^{c} \ln{(c_{i})} + \alpha_{i}^{l} \ln(l_{i}) + \alpha_{i}^{X} \ln{(X)} + \newline 
\alpha_{i}^{H,i} \ln{(h_{i})} + \alpha_{i}^{H,-i} \ln{(h_{-i}) + \alpha_{i}^Z \ln(Z)}
$$

$f$ and $m$ will choose expenditure on market and time-produced (valued at wage) public goods $RQ = rX + w_f h_f + w_m h_m$, where $Q_h = (X, H)$ and $R$ is the price vector. $f$ and $m$ will also bargain over how to divide resources. For simplicity, I only have the public good entering utility through the human capital production function. There is a minimum amount of time required for the public good (i.e. minimum amount of time needed for childcare) which I'm leaving exogenous. $h_f$ and $h_m$ both show up in both utility functions because the men and women gain utility from human capital of the child, which can be supplied by both/either parent(s).
As in Francesca's setting, relative bargaining power will in general depend both on (i) market variables such as prices (including wages), and (ii) distribution factors, which enter the household’s optimisation problem only indirectly through the distribution of bargaining power (individual and household characteristics). The vector of variables that affect bargaining power is $z$, and the household maximises the weighted sum of individual utilities based on the Pareto weights of the members (Browning, Chiappori and Weiss, 2014). Households are solving the following:

$$
\text{max}_{\rho,M,d} \: \mu_{f}(z) u_{f}(c_f, l_f, H) + \newline
(1-\mu_{f}(z)) u_{m}(c_m, l_m, H), \newline
\text{subject to} \newline
y^{NL}_f + w_f m_f + y^{NL}_m + w_m m_m = rX + p(c_f + c_m), \newline
l_i + m_i + h_i = 1, \newline
h_m + h_f \geq \bar{h}, \newline
c_{i}, l_{i}, h_{i} \geq 0, \newline
H = H(h_f, h_m, X, Z)
$$
(Some rough work with the model that won't render correctly)
$$
\mathcal{L} = \mu U_f+(1-\mu)U_m + \lambda_y (y^{NL}_f + w_f m_f + y^{NL}_m + w_m m_m - rX - p(c_f + c_m)) + \lambda_f (1 - l_f - m_f - h_f) + \lambda_m (1 - l_m - m_m - h_m) + \lambda_h (h_f + h_m - \bar{h}) \newline 

\frac{\partial\mathcal{L}}{\partial c_f} = 0  \quad\Rightarrow\quad \mu\frac{\alpha_f^c}{c_f} = \lambda_y p, \newline

\frac{\partial\mathcal{L}}{\partial c_m} = 0 \quad\Rightarrow\quad (1-\mu)\frac{\alpha_m^c}{c_m}=\lambda_y p, \newline

\frac{\partial\mathcal{L}}{\partial l_f} = 0 \quad\Rightarrow\quad \mu\frac{\alpha_f^l}{l_f}= \lambda_f, \newline

\frac{\partial\mathcal{L}}{\partial l_m} = 0 \quad\Rightarrow\quad (1-\mu)\frac{\alpha_m^l}{l_m}=\lambda_m, \newline

\frac{\partial\mathcal{L}}{\partial m_f} = 0 \quad\Rightarrow\quad w_f \lambda_y = \lambda_f, \newline

\frac{\partial\mathcal{L}}{\partial m_m} = 0 \quad\Rightarrow\quad w_m \lambda_y = \lambda_m, \newline

\frac{\partial\mathcal{L}}{\partial X} = 0 \Rightarrow\quad \frac{\mu\alpha_f^X+(1-\mu)\alpha_m^X}{X}=\lambda_y r, \newline 

\frac{\partial\mathcal{L}}{\partial h_f} = 0 \Rightarrow\quad \frac{\mu\alpha_f^{Hf} + (1 - \mu)\alpha_m^{Hf}}{h_f} = \lambda_f - \lambda_h, \newline

\frac{\partial\mathcal{L}}{\partial h_m} = 0 \Rightarrow\quad \frac{\mu\alpha_f^{Hm} + (1 - \mu)\alpha_m^{Hm}}{h_m} = \lambda_m - \lambda_h.
$$

How much of the private assignable good does each person even get, and what determines this? With Cobb–Douglas preferences, homotheticity means a consumer will always allocate a fixed, constant share of their total budget to each good. Individuals spend a constant share of a portion $\rho_{i}$ of the household's resources on private leisure. From the household's problem that I was writing earlier, $\rho_{i}$ is the individual’s share of the household budget net of public goods expenditure:
$$
w_{i}l_{i} = \alpha^l_{i}\rho_{i}, \newline
\rho_{i} = \eta_{i}(y - RQ), \newline
w_{i}l_{i} = \alpha^l_{i}\eta_{i}(y - RQ)
$$

i feel like im explaining everything from the paper that was too obvious to actually wrte down

how much the objective improves if the constraint is relaxed: relaxing time constraint is valued at wage

the shadow cost of the childcare time constraint enters the same way through both the first order conditions. if the parents are up against the constraint (i.e. $\lambda_h \neq 0$), bargaining power shifts the marginal benefit of each parent's provision asymmetrically. the effect of relative bargaining power on who supplies required childcare depends on how the empowered spouse values maternal vs paternal childcare inputs in the child production/utility system, relative to the spouses’ wages

## Finding resource shares

Identification will rely on a private assignable good. A good is private and assignable if we observe that it is consumed by a specific person and we can attribute demand or expenditure for the good to that person. For instance, a lot of research in collective households uses clothing as it shows up in expenditure surveys. This is because clothing can be categorised into men's, women's, and children's, which attributes it to a specific member of the household. However, clothing is relatively durable so expenditure surveys may not capture any expenditure on it.
In this part, I was working out what the key identification issues would be and was initially writing everything out in the context of the more flushed-out model from Francesca's paper because I was trying to really understand how much restriction was needed for identification. 

### Working with time use data: preference restrictions on public goods expenditure


The SAP-type public goods preference restriction is that for each person in the household, the sum of Cobb-Douglas preference parameters is the same:

$$\alpha_{i}^{X} + \alpha_{i}^{Hf} + \alpha_{i}^{Hm} = a$$

With Cobb-Douglas utility, the share of the full budget spent on a block equals the sum of the preference parameters on that block, so the household spends a constant fraction $a$ of $y$ on public goods: $RQ = ay$. However, this assumption doesn't mean that the composition of public goods spending isn't allowed to change. It states that households may choose to spend on different public goods but the overall public-private composition is the same. For example, higher relative male bargaining power could lead to more spending on daycare and less spending on self-provided childcare, where spending would be valued at the person's wage. 

So public-goods C-D SAP gives you $(y - R Q) = a y$. For each person, this means their private leisure expenditure is given by $w_{i}l_{i} = \alpha^l_{i} \eta_{i} ay$

... still working through this one but basically if public-goods C-D SAP holds, total public spending will be a fixed fraction of $y$ for some household so that variation in $y$ maps cleanly into variation in the private goods part. Then, variation in leisure expenditure in $y$ is informative about $\eta$ (subject to the further private goods restriction we would need in order to deal with $\alpha^l$.

If it fails, $R Q$ (and hence $y-RQ$) can vary with distribution factors $z_h$. Then the slope on $y(z - \bar{z})$ conflates changes in $\eta_{i}$ and changes in the size of the private residual $(y-RQ)$, so the $\beta^z$’s no longer isolate sharing-rule responses. If bargaining power shifts how much the household “wants” to devote to public goods in general, then $(y-RQ)$ moves with $z$ which would mean resource shares are not identified. It could be that female empowerment shifts the total desired spending on public goods through a parenting/childcare channel (not just the composition of spending), which causes the overall private/public split to move with relative bargaining power.
In fact, using family data probably violates assumption that total expenditure on public goods does not vary with the distribution of bargaining power in the household. 

### Assumptions restricting preference heterogeneity on the private assignable good

For a household (h) in category (g), individual (t)’s resource share is approximated as

$$ \eta_{t,h} = \eta_{t,g}^0 + \sum_z \eta_{t,g}^z (z_h - \bar z)$$
For every household (h), $\sum_{t \in h} \eta_{t,h} = 1$

Plug in the approximation:

$$\sum_{t\in h}\eta_{t,g}^0 + \sum_z (z_h-\bar z)\sum_{t\in h}\eta_{t,g}^z =1$$

Since this must hold for all $z_h$, we need:

1. $$\sum_{t\in h}\eta_{t,g}^0 = 1$$

2. $$\sum_{t\in h}\eta_{t,g}^z = 0 \quad\forall z$$

With Cobb–Douglas preferences over private leisure and under the two-stage structure, individual $t$’s leisure expenditure satisfies

$$ w_{t,h} l_{t,h} = \alpha_{t,g}^l y_h (\eta_{t,g}^0 + \sum_{z} \eta_{t,g}^{z}(z_h - \bar{z})) \eta_{t,h} $$

This provides estimating equations of the form

$$ w_{t,h} l_{t,h} = \beta_{t,g}^0 y_h + \sum_z \beta_{t,g}^z , y_h (z_h-\bar z) + \varepsilon_{t,h}$$

where

$$\beta_{t,g}^0 = \alpha_{t,g}^l \eta_{t,g}^0 $$
and 

$$ \beta_{t,g}^z = \alpha_{t,g}^l \eta_{t,g}^z. $$

The SAP assumption states that for a given household category $g$, all individuals have the same Cobb–Douglas parameter for private leisure:
$$\alpha_{t,g}^l = \alpha_g^l \quad \forall t\in h$$
Under SAP, the slope coefficients become $\beta_{t,g}^z = \alpha_g^l \eta_{t,g}^z$ 

Now sum over household members:

$$ \sum_{t\in h}\beta_{t,g}^z = \alpha_g^l \sum_{t\in h}\eta_{t,g}^z$$

But from the adding-up condition above: 
$$\sum_{t\in h}\eta_{t,g}^z = 0 $$

Therefore:
$$ \sum_{t\in h}\beta_{t,g}^z = 0$$ 
Without SAP, this is

$$ \sum_t \beta_{t,g}^z = \sum_t \alpha_{t,g}^l \eta_{t,g}^z$$

which does not in general equal zero unless $\alpha_{t,g}^l$ is constant across (t).

If SAP fails, then for each $z$:

$$\beta_{f}^z = \alpha_f^l \eta_f^z,
\quad
\beta_{m}^z = \alpha_m^l \eta_m^z$$

We would observe $\beta_f^z, \beta_m^z$ and two equations with four unknowns:
$$\alpha_f^l,\ \alpha_m^l,\ \eta_f^z,\ \eta_m^z$$

With Cobb-Douglas utilities, men and women have the same preference parameter for leisure. They will spend the same share of their individual second-stage budget on leisure, but they may make different choices for the other public goods they consume.
Similarity across types: Slightly different assumption that some papers impose but this is also something I will touch on

### Linear estimation

-   Resource shares will generally depend on household budget, prices (including wages of all household members) and Pareto weights of household members: $\eta_{i,h}(y_h, p, r, w_h, µ_h)$.
-   Recall that Pareto weights $\mu_h$ are an unknown function of a vector of variables that determine relative bargaining power in the household, and that the collective model does not pin down a specific bargaining solution. Since $z_h$ in general includes the other variables which affect $\eta_{i,h}$, we can write $\eta_{i,h}(z_h)$
-   To estimate resource shares by linear regression, we linearly approximate them: $$\eta_{i,h} = \eta_{i,h}^0 + \sum_{z} \eta_{t,g}^{z} (z_h - \bar{z})$$
-   $\eta_{i,h}^0$ is the average resource share of a type (living in a specific household category). By definition, $\sum_{t \epsilon h} \eta_{t,g}^0 = 1$. For instance, a higher-than-average wage for the woman might increase the woman’s resource share, so that she would have a higher-than-average-for-women resource share.
-   $\eta_{t,g}^{z}$ captures the impact on sharing of a household’s characteristic $z_h$ deviating from the sample average $(z_h - \bar{z})$
-   An equation for each type of person in the form: 

$$w_{i,h}l_{i,h} = \alpha_{t,g}^{l} y_h (\eta_{t,g}^0 + \sum_{z} \eta_{t,g}^{z}(z_h - \bar{z})) + \epsilon_{i,h}$$
-   If $z_h$ includes all key characteristics that affect bargaining (or at least all those that are correlated with regressors) then estimates should be unbiased. The errors are negatively correlated within household (with a correlation weaker in magnitude than -1), so that a SURE estimation approach is recommended.
-   Equation for women: 
$$w_{i,h}l_{i,h} = \beta_{f,g}^0 y_h + \sum_{z} \beta_{f,g}^{z} y_h (z_h - \bar{z})) + \epsilon_{i,h}$$
-   Equation for men: 
$$w_{i,h}l_{i,h} = \beta_{m,g}^0 y_h + \sum_{z} \beta_{m,g}^{z} y_h (z_h - \bar{z})) + \epsilon_{i,h}$$
-   Then, $\beta_{t}^0 = \alpha_{t,g}^{l} \eta_{t,g}^0$ and $\beta_{t}^z = \alpha_{t,g}^{l} \eta_{t,g}^z$.
-   First, linear SURE regressions of leisure expenditure, one for men and one for women, with the restriction that $β_{f}^z + β_{m}^z = 0$. This is because $\beta_{t}^z = \alpha_{t}^{l} \eta_{t}^z$ and from private leisure C-D SAP and the definition of the linear approximation parameters, $\sum_{t \epsilon h} \beta_{t}^z = \alpha_{t}^{l} \sum_{t \epsilon h} \eta_{t}^z = 0$.
-   Next, estimate each type’s average resource share as $\hat{\eta}_{t}^0 = \frac{\hat{\beta_{t}^0}}{\hat{\beta_{f}^0} + \hat{\beta_{m}^0}}$
-   Next, estimate $\hat{\alpha_{t}^{l}} = \frac{\hat{\beta_{t}^0}}{\hat{\eta}_{t}^0}$, since $\beta_{t}^0 = \alpha_{t}^{l} \eta_{t}^0$.
-   Next, estimate $\hat{\eta_{t}^{z}} = \frac{\hat{\beta_{t}^z}}{\hat{\alpha_{t}^{l}}}$, since $\beta_{t}^z = \alpha_{t}^{l} \eta_{t}^z$.
-   Armed with the sharing rule, one can estimate individual-specific resource shares in the dataset used for estimation (the time-use dataset in this case): $\hat{\eta_{i,h}} = \hat{\eta}_{t}^0 + \sum_{z} \hat{\eta_{t}^{z}} (z_h - \bar{z})$.

# Estimating the sharing rule using household time use data

This part basically constitutes a replication of Francesca's work. I want to use this section to show some of the differences in how these structural estimates behave compared to the childless couple case and talk about why this could be happening and how identification could be impacted. Because of my later work with time diaries, I filtered for households where both partners/spouses filled out both days of the time diary. I have pooled samples of 704 couples without children and 635 couples with children across both UKTUS survey waves.  Wages are calculated using actual hours worked and actual earnings from a respondent's main job. Both employee earnings and self-employed earnings are used.
- I've never quite been able to replicate empirical results from any paper perfectly but my non-parent characteristics do differ slightly from Francesca's. I have not removed any of the very high earners from my data.
- The UKTUS does not contain enough data to distinguish between households who never had children and households with non-cohabiting children, so childless couples may tend to be older. As expected, the parents in my data occupy a smaller age range and are slightly younger. Both groups achieve similar average qualifications. Women are slightly more qualified than men and tend to earn slightly less. The wage standard deviation is high. This is due to relatively high earners with low working hours.
```{r echo=FALSE}

summary = c("wage_f", "wage_m", "DVAge_f", "DVAge_m", "educ_f", "educ_m")

nonparents_summary = t(sapply(nonparents_est_data_merged[summary], summarytable))
parents_summary = t(sapply(parents_est_data_merged[summary], summarytable))

print("UKTUS working heterosexual couples (parents), pooled sample")
print(as.data.frame(parents_summary))
print("UKTUS working heterosexual couples (nonparents), pooled sample")
print(as.data.frame(nonparents_summary))
```

## Baseline results

- Here are the results of SURE regressions of leisure expenditure on determinants of bargaining power $z$ among parents. By the linear estimating equations, the average point for a mother's resource share, for instance, is given by the estimated coefficients on household budget as follows: $\hat{\eta}^0_f = \frac{0.0272}{0.0327 + 0.0272} \approx 0.4541$. On average, fathers are slightly favoured using a definition of private leisure that excludes personal care and sleep, two categories that are less discretionary and more likely to have externalities on other people in the household. However, mothers are slightly favoured when these activities are included. We get estimated preference parameters on private leisure by dividing the household budget coefficient in the female regression by the average resource share calculated previously. 
- Estimated parameters fall once I restrict leisure (pretty predictably, as I'm now excluding a component of leisure that is very time-intensive)
  - Leisure preference parameter excluding personal care and sleep = 0.0272 / (0.0272 / (0.0272 + 0.0327)) = 0.0327 / (0.0327 / (0.0272 + 0.0327)) = 0.0599
  - Leisure preference parameter including personal care and sleep = 0.153 / (0.153 / (0.153 + 0.138)) = 0.138 / (0.138 / (0.138 + 0.153)) = 0.291
- By the preference heterogeneity restriction, these are constructed to be the same for males and females. Unit deviations from the mean of hourly wage (in 2025 GBP), qualification level (on the scale 0, 1, 2), the average age of the couple, and the age gap of the couple are calculated by dividing the coefficient on the corresponding interaction term by the estimated preference parameter. Only the coefficient on the household budget and the male and female wage deviations are significant determinants of bargaining power when using the broad leisure expenditure category. 

```{r echo=FALSE}
# deviations of characteristics compared to own sex
print("Leisure expenditure of parents (excluding sleep and personal care)")
print(make_sysfit_kable_df(parents_res_r_within_merged, rows_within))

# deviations of characteristics compared to own sex
print("Leisure expenditure of parents (including sleep and personal care)")
print(make_sysfit_kable_df(parents_res_within_merged, rows_within))
```

- Here are the regression results for the childless couples:
  - Leisure preference parameter excluding personal care and sleep = 0.0991 / (0.0991 / (0.0991 + 0.133)) = 0.133 / (0.133 / (0.133 + 0.0991)) = 0.2321
  - Leisure preference parameter including personal care and sleep = 0.162 / (0.162 / (0.162 + 0.232)) = 0.232 / (0.232 / (0.232 + 0.162)) = 0.394
- The preference parameters for the parents are lower than those for the childless couples, especially so for discretionary activities like entertainment and sport. This is consistent with individuals spending a significant portion of their daily time on leisure, with the presence of children constraining some of this time availability. Indeed, it's much more difficult to reduce sleep and personal care.
- Magnitude/sign/direction of the coefficients are similar to the previous regressions, but more are significant, aside from average age. It looks like in the less constrained setting, the set of effective bargaining instruments grows (or there's just more variation in labour-leisure choices than when time is more scarce)
```{r echo=FALSE}
# deviations of characteristics compared to own sex
print("Leisure expenditure of non-parents (excluding sleep and personal care)")
print(make_sysfit_kable_df(nonparents_res_r_within_merged, rows_within))

# deviations of characteristics compared to own sex
print("Leisure expenditure of non-parents (including sleep and personal care)")
print(make_sysfit_kable_df(nonparents_res_within_merged, rows_within))
```

-   When it comes to private leisure excluding sleep and personal care, sharing is pretty equal. This is not robust to adding sleep and personal care to my definition of leisure, with women generally commanding more of the overall share of resources through this channel. This seems weird to me, although the unrestricted regressions show the same thing. Here are some density plots that show resource share estimates. All estimates fall within the unit interval. This isn't something I can impose by construction, so the fact that this linear estimation method provides shares that can reasonably be interpreted as a percentage of resources so in this sense the model does appear to fit the data well.
- But actually though I don't think it's that simple because splitting up the two survey years shows quite different results in both years (will talk about this later)
```{r echo=FALSE, fig.width=5, fig.height=8}

narrowleisure_shares = plot_share_densities(shares_parents_ownsex_r_merged$data, 
                                            dev_type = "own",
                                            restrict_leisure = TRUE)

broadleisure_shares = plot_share_densities(shares_parents_ownsex_merged$data, 
                                           dev_type = "own",
                                           restrict_leisure = FALSE)

combined_plot =
  narrowleisure_shares + broadleisure_shares +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Estimated resource shares (parents)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))

print(combined_plot)
```
- (estimated resource shares for non-parents)
```{r echo=FALSE, fig.width=5, fig.height=8}

narrowleisure_shares = plot_share_densities(shares_nonparents_ownsex_r_merged$data, 
                                            dev_type = "own")

broadleisure_shares = plot_share_densities(shares_nonparents_ownsex_merged$data, 
                                           dev_type = "own",
                                           restrict_leisure = FALSE)

combined_plot =
  narrowleisure_shares + broadleisure_shares +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Estimated resource shares (non-parents)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))

print(combined_plot)
```

-   Here are some plots with the marginal contributions of each household characteristic to the estimated sharing rule. This one is for the merged data (2000 and 2015). Women’s resource shares are predicted higher in households with higher female hourly pay and lower male hourly pay, but the male wage effect is nearly twice as large.
-   An increase in the age gap in the direction I've specified means that the male age is relatively higher. This has a negative effect on the female resource share, as does an increase in the average age of the couple. 
-   What's interesting is that the education effects are kind of the opposite of what was found for the couples without children. Increasing male education has a smaller positive effect and increasing female education has a relatively larger negative effect.
-   Being in a wealthier region generally means female resource shares are smaller.

```{r echo=FALSE}
marginal_impacts_merged = marginal_impacts(shares_parents_ownsex_r_merged, 
                                           parents_est_data_merged, rows_within)

kable(marginal_impacts_merged) |>
  kable_styling("striped", full_width = F)

```

### How are people spending their time?
- See if women with lower or higher relative bargaining are already providing a lot on the weekdays.
```{r echo=FALSE}
reg_stacked = bind_rows(
  data_working_parents_2000 |>
    transmute(
      serial,
      sample = 2000,
      dgorpaf,
      pnum,
      is_weekend = as.integer(is_weekend), # dummy
      male = as.integer(male), # dummy
      wage, 
      total_work,
      total_private_leisure,
      total_private_leisure_r,
      total_childcare_nospouse,
      total_childcare,
      kid_age_min,
      num_kids_total,
      num_kids_male),
  data_working_parents_2015 |>
    transmute(
      serial,
      dgorpaf,
      pnum,
      sample = 2015,
      is_weekend = as.integer(is_weekend),
      male = as.integer(male),
      wage,
      total_work,
      total_private_leisure,
      total_private_leisure_r,
      total_childcare_nospouse,
      total_childcare,
      kid_age_min,
      num_kids_total,
      num_kids_male)) |>
  # time dummies
  mutate(dummy_2000 = ifelse(sample == 2000, 1, 0),
         dummy_2015 = ifelse(sample == 2015, 1, 0)) |>
  
  # only personal care
  mutate(total_private_personalcare = total_private_leisure - total_private_leisure_r) |>
  
  # individual identifiers
  group_by(sample, serial, pnum) |>
  mutate(individual_id = row_number()) |>
  ungroup() |>
  
  # child under five in household
  mutate(child_under_five = ifelse(kid_age_min <= 5, 1, 0),
         proportion_male_children = num_kids_male / num_kids_total) |>
  
  # merge with resource share estimates
  inner_join(shares_parents_ownsex_r_merged$data, by=c("serial"))


timeuse_long = reg_stacked |>
  filter(male %in% c(0, 1), is_weekend %in% c(0, 1)) |>
  transmute(
    shareown_etahat_f,
    day = if_else(is_weekend == 1, "Weekend", "Weekday"),
    sex = if_else(male == 1, "Men", "Women"),
    total_childcare_nospouse,
    total_private_leisure,
    total_private_leisure_r,
    total_childcare) |>
  pivot_longer(
    cols = c(total_childcare_nospouse, total_private_leisure, 
             total_private_leisure_r, total_childcare),
    names_to = "activity",
    values_to = "hours") |>
  mutate(
    activity = factor(
      activity,
      levels = c("total_childcare_nospouse", "total_private_leisure", 
                 "total_private_leisure_r", "total_childcare"),
      labels = c("Childcare (without spouse present)", 
                 "Private leisure",
                 "Private leisure (excluding sleep and personal care)",
                 "Childcare (general)")))
```

- Some descriptive visuals about what people are doing, but I don't think this is particularly informative? It just looks like women tend to supply more childcare in general
```{r echo=FALSE, warning = FALSE, message = FALSE}
# on the weekend, how do women spend their time scattered by estimated bargaining power
ggplot(timeuse_long |> filter(activity %in% c("Childcare (without spouse present)", 
                                              "Private leisure (excluding sleep and personal care)")),
       aes(x = shareown_etahat_f, y = hours, color = activity)) +
  geom_point(alpha = 1, shape = ".") +
  scale_x_continuous(limits = c(0, 1)) +
  facet_grid(day ~ sex) +
  labs(
    x = "Estimated Female Resource Share",
    y = "Hours",
    color = "Activity") +
  theme_minimal()
```


## Comparing estimates across both UKTUS survey years:

- These look super funky
```{r echo=FALSE, fig.width=5, fig.height=8}

narrowleisure_shares = plot_share_densities(shares_parents_ownsex_2000$data, 
                                            dev_type = "own")

broadleisure_shares = plot_share_densities(shares_parents_ownsex_2015$data, 
                                           dev_type = "own")

combined_plot =
  narrowleisure_shares + broadleisure_shares +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Estimated resource shares 2000 vs 2015",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11)))

print(combined_plot)
```

-   Now, I look at the marginal effects of bargaining power determinants for both years of the UKTUS. In 2000, both female wage and female education outweigh the male wage and male education effects, so that women in households with higher household budgets have higher resource shares.

```{r echo=FALSE}
marginal_impacts_2000 = marginal_impacts(shares_parents_ownsex_r_2000, 
                                         parents_est_data_2000, rows_within)

kable(marginal_impacts_2000) |>
  kable_styling("striped", full_width = F)
```

-   We see that opposite education effect in the 2015 data, where female education has a negative impact on female resource shares and male education has a negative impact on male resource shares. The female education effect is statistically significant, unlike in the 2000 data. I don't really know why this would happen, but in a setting with children, increasing your educational attainment means delaying having a child, spending less time with your child once you do have one, or maybe both. This is true for both parents, but especially for mothers as they're the ones who actually have to give birth and stuff. Also, because takeup for educational attainment has been growing faster for females than males, the returns to female education could be lower (since females are increasing their educational attainment yet facing a comparatively unchanging group of potential matches)

```{r echo=FALSE}
marginal_impacts_2015 = marginal_impacts(shares_parents_ownsex_r_2015, 
                                         parents_est_data_2015, rows_within)
kable(marginal_impacts_2015) |>
  kable_styling("striped", full_width = F)
```

## Kind of testing assumptions:

- If private leisure C-D SAP holds then the proportion of household budget spent on overall private leisure (the sum of expenditure on female private leisure and expenditure on male private leisure) should not vary with the distribution of bargaining power. 
- The share of household budget spent on private leisure is slightly negatively correlated with the woman's estimated resource share, but this correlation loses significance once I include personal care and sleep in total leisure expenditure to match Francesca's definition of leisure. Personal care and sleep are far less discretionary than other types of leisure, so the issues with private leisure C-D SAP might be coming from

```{r echo=FALSE}
# check budget share on private leisure correlated w/ estimated resource shares

# merge with sharing_est_data
parents_est_data_merged_shares = inner_join(shares_parents_ownsex_r_merged$data,
                                parents_est_data_merged,
                                by=c('serial')) |>
  mutate(share_budget_leisure_r = (private_leisure_exp_r_f + 
           private_leisure_exp_r_m) / y,
         share_budget_leisure = (private_leisure_exp_f + 
           private_leisure_exp_m) / y)

# correlation of estimated resource shares (broad leisure definition)
cor.test(parents_est_data_merged_shares$share_budget_leisure, 
         parents_est_data_merged_shares$shareown_etahat_f)

# (narrow leisure definition)
cor.test(parents_est_data_merged_shares$share_budget_leisure_r, 
         parents_est_data_merged_shares$shareown_etahat_f)
```

- The SRAT assumption allows preference levels to vary between different people and different household categories, while requiring that the ratio of the parameters is constant across household compositions. SRAT may be more defensible once I go to a setting with children, as parenthood plausibly shifts the level but not the ranking of leisure preferences between men and women. This could be more believable than men and women having the same level of preference for private leisure within household type. However, it is still a strong assumption as parenting likely changes preferences asymmetrically due to gender norms. Suppose motherhood reduces women’s leisure preference relative to men, that is, women put less weight on private leisure (or face guilt/constraints that behave like a lower effective preference parameter). If women have a lower taste for private leisure, they will choose less leisure even with the same resources. (Not really sure what I can do with this yet)
- Using only time use data requires the Public Goods C-D SAP assumption that preference parameters over all public goods sum to the same quantity for men and women. With time use data, I don't have information on market goods. I don't particularly want to clean an entire expenditure survey dataset, so I think what I can do here is see (at least indirectly) whether public-good SAP plausibly fails specifically through childcare or maybe through domestic work (similarly valued at wage). I do have extensive time use data about domestic work. (Not really sure what I can do with this yet)

## How private leisure responds to childcare need (constraints?)

- School-aged and daycare-aged children mechanically spend more time outside the home during weekdays since most schools/kindergartens/nurseries are closed on weekends. Weekends shift the feasible set as childcare needs at home rise. I have one weekday and one weekend diary for each member of each household, so I can see who is absorbing this extra childcare need. 
- Through one channel: If mothers already have low weekday private leisure, they may be at (or near) a constraint where further childcare would imply very high marginal disutility. The mother’s marginal utility loss from reducing $l_f$ further is very steep so the father’s leisure drops more when $\bar{h}$ rises. 
- Through another channel: If, say, female bargaining power is relatively high in some household, then when $\bar{h}$ rises the model predicts that fathers will absorb more of the childcare adjustment because the planner places more weight on mother’s utility.
- I still have to think about what else I could do with this! It might be useful to tie it back to the work from above... the weekend allows me to test how household production constraints shifts translate into private allocation

Simple regressions: Given the amount each parent works, do weekends reallocate leisure differently for males and females? I started off not conditioning on work hours
```{r echo=FALSE}
etable(
  feols(total_private_leisure_r ~ is_weekend * male |
          serial + sample + dgorpaf, data = reg_stacked),
  se = "cluster", cluster = "serial"
)

etable(
  feols(total_private_leisure_r ~ is_weekend * male + total_work |
          serial + sample + dgorpaf, data = reg_stacked),
  se = "cluster", cluster = "serial"
)
```

Without very young kids, weekends reduce mothers’ private (discretionary) leisure. Among households without young children, fathers lose less leisure than mothers on weekends. In households with young children, fathers lose about 1 hour more private leisure on weekends than mothers, relative to households without young children. Private leisure is not symmetrically adjustable across genders when childcare intensity changes. Either fathers having lower Pareto weights in that state or fathers having more slack (i.e. mothers are closer to a constraint). I could also try to see whether mothers with young children already have very low weekday private leisure and aren't at a point where they can even adjust the amount of childcare they provide. The note about the `child-under-five` being removed because of collinearity is due to the household fixed effects.
```{r echo=FALSE}
etable(feols(
  total_private_leisure_r ~ 
    is_weekend * male * child_under_five +
    total_work |
    serial + sample + dgorpaf,
  data = reg_stacked,
  cluster = "serial"
))
```

- I'll also try the same regression but with childcare. Again, that triple interaction coefficient is positive and significant, so fathers absorb weekend childcare shock. Actually, the 
```{r echo=FALSE}
etable(feols(
  total_childcare_nospouse ~ 
    is_weekend * male * child_under_five +
    total_work |
    serial + sample + dgorpaf,
  data = reg_stacked,
  cluster = "serial"
))
```

# side quests

### A closer look at deviations of average household characteristics

-   Francesca uses the following example for deviation: A higher-than-average wage for the woman might increase the woman’s resource share, so that she would have a higher-than-average-for-women resource share.
-   Bargaining power may come from strength relative to one's spouse or strength relative to a spouse's options within one's own gender. Say in 2006 there is a leisure gap when women contribute more to the household. This may result in smaller increases in women’s leisure than if it were men.
-   It's worth seeing the women-vs-women/men-vs-men and women-vs-all/men-vs-all comparison (think about someone's strength relative to their spouse's other options or relative to their sponse), but estimated resource shares using any of these definitions are very similar. I've checked that I haven't been calculating them incorrectly lol

Using just relaxing, socialising, sport, entertainment, and hobbies:
```{r echo=FALSE}
# deviations of characteristics compared to own sex
print(make_sysfit_kable_df(parents_res_r_within_merged, rows_within))

# deviations of characteristics across both sexes
print(make_sysfit_kable_df(parents_res_r_between_merged, rows_between))

# deviations of characteristics compared to opposite sex
print(make_sysfit_kable_df(parents_res_r_opposite_merged, rows_opposite))
```

Including eating, sleeping, and personal care:
```{r echo=FALSE}
print(make_sysfit_kable_df(parents_res_within_merged, rows_within))

print(make_sysfit_kable_df(parents_res_between_merged, rows_between))

print(make_sysfit_kable_df(parents_res_opposite_merged, rows_opposite))
```

- Also, here are the unconstrained SURE regressions for comparison. I find the magnitudes and signs of coefficients (and significance) are similar to the restricted regressions.
```{r echo=FALSE}
# without coefficient restrictions
print(make_sysfit_kable_df(parents_unres_r_within_merged, rows_within))
# with coefficient restrictions
print(make_sysfit_kable_df(parents_res_r_within_merged, rows_within))
```

### How would households with characteristics $z_h$ be treated by a different sharing rule? (I did do this but not sure my code is correct!)

-   Changes over time: take income and matchings from 2000, and take coefficients from 2015. apply the betas to the observed behaviour of 2015. Would I get different counterfactual leisures?
-   For two periods, I estimate $\beta_t^0 = \alpha_t^\ell \eta_t^0,$, $\beta_t^z = \alpha_t^\ell \eta_t^z$, average resource shares $\hat\eta_t^0 = \frac{\hat\beta_t^0}{\hat\beta_m^0 + \hat\beta_f^0}$, the leisure preference parameter $\hat\alpha_t^\ell = \frac{\hat\beta_t^0}{\hat\eta_t^0}$, and each marginal sharing rule $\hat\eta_t^z = \frac{\hat\beta_t^z}{\hat\alpha_t^\ell}$.
-   So for each year (say 2000 and 2015), I'd get: $$\hat\eta_{t,h}^{(2000)} = \hat\eta_t^{0,(2000)} + \sum_z \hat\eta_t^{z,(2000)} (z_h - \bar z), \newline
    \hat\eta_{t,h}^{(2015)} = \hat\eta_t^{0,(2015)} + \sum_z \hat\eta_t^{z,(2015)} (z_h - \bar z)$$
-   Then, why not take households observed in 2000 (their $z_h$), but apply the newer sharing rule: $$\hat\eta_{i,h}^{CF} = \hat\eta_t^{0,(2015)} + \sum_z \hat\eta_t^{z,(2015)} (z_h^{2000} - \bar z^{2015})$$
-   I'd want to compare $\hat\eta_{t,h}^{CF} \quad$ and $\quad \hat\eta_{t,h}^{(2000)}$.
-   If we have $\hat\eta_{f,h}^{CF} > \hat\eta_{f,h}^{(2000)}$ for the observed 2000 household characteristics, the sharing rule has shifted in a way that allocates a larger share of resources to women.
-   When the deviations of a household's characteristics are relative to averages across both males and females, I'd propose that a higher $\hat\eta\_t\^{z,(2015)}$ than $\hat\eta\_t\^{z,(2000)}$ for some $z$ is capturing increased importance of someone’s absolute wage/education/whatever is in determining allocation, more like “my wife’s human capital carries more weight in allocation than it used to” than “my wife's position relative to my other options carries more weight than it used to”.
-   When the deviations of a household's characteristics are relative to averages of only the respective gender associated with the characteristic (for instance, the deviation of the man's wage from the wage of other men), I'd propose that a higher $\hat\eta_t^{z,(2015)}$ than $\hat\eta_t^{z,(2000)}$ for some $z$ is capturing increased importance of someone’s absolute wage/education/whatever is in determining allocation, more like how they compare to the opposing spouse's outside options became more important for intrahousehold allocation. So, it's like the other way around from the previous point.
-   For a counterfactual analysis, I can provide tables with two sets of estimated average resource shares and marginal contributions, each corresponding to the older or newer years, as well as the distribution of estimated counterfactual resource shares in the older data.
-   Suppose in 2000, the woman's wage relative to all other wages mattered less for the sharing rule, i.e. $\hat\eta_t^{w_f,(2000)}$ is small.
-   In 2015, maybe $\hat\eta_t^{w_f,(2015)}$ is larger. This would mean that household allocations are more affected by the woman's wage relative to all other wages. Women’s human capital carries more weight in allocation than it used to
-   Applying the 2015 coefficients to 2000 households: even though 2000 women earned less on average, the same relative wage difference would buy them more bargaining power under the 2015 rule.
-   The identifying assumption is: conditional on $z_h$, the estimated sharing rule captures a stable mapping from characteristics to allocation, which is the same assumption Arduini (2025) relies on.

## random comments 

-   Generally, people take the approach of considering children as public goods or modelling them as individual members with their own preferences (like in the Dunbar, Lewbel, and Pendakur papers using expenditure data).
-   The main limitation of private leisure is that its price is only known for those who work, for whom it coincides with the observed wage. For non-participants like children, the price of time is higher than their (unobserved) wage and we would have to estimate that in order to use time-use data to identify sharing. In the Arduini paper, she focuses on working couples and leaves this extension for future work. However, I think children are even more complicated than non-working adults.
-   The way I'm estimating the resource shares suggests that children themselves don't exist as household members in their own right, since the estimated shares add up to one for the male and female spouse/partners, and so falls under the public goods that each spouse/partner provides. I think one of the limitations of child-specific data in general, whether on time use or expenditure of goods in the strict sense, is that children may not have a lot of control over how they spend their time even without parents or other household members present (e.g. specified break times in school). So not only do children not earn wages, they don't even have a reliable way to enforce their own preferences.
-   If I had a third equation linked to a child-assignable good, I'd use $\eta_m + \eta_f + \eta_c = 1$
-   Parents presumably get enjoyment and returns out of parenting and spending time with their children. This is even the case for child-specific activities like homework and a child's private leisure, since children benefit developmentally from them and parents then benefit from this development. I think it would be really challenging to come up with a) a private assignable good and b) a shadow price of private leisure for children.
-   This line of thinking could be extended to anyone's private leisure or even anyone's clothing. If one spouse takes assignable time for private leisure and this gives them something nice to talk about with the other, or if one spouse likes it when the other wears new clothing, then neither are really private.
-   Dauphin et al (2008) and Cherchye, De Rock and Vermeulen (2008) test whether observed household demand functions are consistent with children having separate utility functions, and find evidence in favor of this hypothesis, though they do not estimate children’s shares.
-   The collective household model has been proposed partly in response to the dissatisfaction with the unitary model. The former simply assumes that the family decision process, whatever its exact nature, leads to Pareto efficient outcomes. One interesting feature of the collective model is that it does not require the specification of the bargaining process. Evidently, the generality of the efficiency assumption comes at a price: the model is not very informative about which variables, other than those in the budget constraint, may influence the decision process. In particular, the model is totally silent as to the role children may play in the process.
-   Harbaugh et al. (2001) found that “at age 7 children’s choices about consumption goods show clear evidence of rationality, though also many inconsistencies. By age 11, choices by children... are as rational as choices by adults”. Harbaugh et al. (2003) further showed in another experiment that children display good bargaining skills as early as 7 years of age. Naturally, one may object to the inclusion of young children as decision-makers on several grounds. Within a bargaining framework, the child’s threat point could correspond to his level of well-being in a non-cooperating equilibrium with his parents. Parents could then resort to punishment to reduce their child’s bargaining power or to make it non existent.

## papers

Lewbel and Pendakur (2008) and Bargain and Donni (2012) rely on the ‘SAT’ assumption (similarity across household types), that preferences for the assignable good are stable across household compositions, allowing for identification of the preference parameter from singles data

[Lise and Seitz (2011)](https://www.jstor.org/stable/23015857): restricting the bargaining process so that men and women with equal earnings potential share resources equally

[Browning, Chiappori and Lewbel (2013)](https://academic.oup.com/restud/article-abstract/80/4/1267/1584142): BCL

\[Dunbar, Lewbel, and Pendakur (2013)\]\[https://www.sfu.ca/\~pendakur/CollectiveChildren30-with-appendices.pdf\]: Identifies how total household resources are divided up among household members, by observing how each family member’s expenditures on a single private good like clothing varies with income and family size. DLP permits identification of resource shares of both adults and children, where children are treated as having their own utility functions and welfare. This is in contrast to most of the empirical collective household literature, where expenditures on children are modelled just as public goods in the adults’ utility functions. Identification assumes that resource shares do not vary with total expenditure. DLP and LPW use the SAP assumption (similarity across people) that women and men in the same household composition have the same preference parameter for the assignable good.

[Lechene, Pendakur, and Wolf (2022)](https://www.journals.uchicago.edu/doi/10.1086/717892): LPW shows that, building on the identification result of DLP, it is possible to estimate resource shares by linear regression

\[Blundell et al (2019)\]\[https://www.sfu.ca/\~pendakur/papers/BKLP16_longtitle.pdf\]

\[Hyee and Robledo (2010)\]\[https://www.nottingham.ac.uk/economics/documents/discussion-papers/10-06.pdf\]

\[Gitter and Barham (2010)\]\[https://academic.oup.com/wber/article-abstract/22/2/271/1736679?redirectedFrom=fulltext&login=false\]

Calvi (2002): Under some mild assumptions, there exists a monotonic correspondence between Pareto weights and resource shares. Consequently, resource shares may also be interpreted as a measure of intra-household bargaining power. In effect, these are more tractable measures of power, as they are invariant to unobservable cardinalisations of the utility functions.

Calvi (2020) estimates resource shares and poverty at the individual level in India and finds that women—especially older women—have lower resource shares than men. This then implies that older Indian women have much higher poverty rates than previously thought. Calvi shows that these higher poverty rates (driven by lower resource shares) among older women can explain the finding of Anderson and Ray (2010) that Indian women over the age of 45 have higher mortality rates than do Indian men (a phenomenon they call “missing women”).

\[Andrew et al (2021)\]\[https://ifs.org.uk/inequality/wp-content/uploads/2021/03/BN-The-careers-and-time-use-of-mothers-and-fathers.pdf\] show that heteronormative households, mothers are more likely to leave the labour force after having children regardless of the relative wages between mother and father before childbirth.

\[Dunbar, Lewbel, and Pendakur (2017)\]\[https://www.bankofcanada.ca/wp-content/uploads/2017/10/swp2017-45.pdf\]: Generalizes DLP in two important ways. One is the identification of resource shares without imposing the BCL or DLP restrictions on similarity of preferences across households of different composition, or across people within a household. The other is by allowing for random variation in resource shares across households, presence of unobserved distribution factors. This reflects the fact that the relative bargaining power of individual household members, or the choice of social welfare function for the household, will in general vary across households for unobserved reasons. Assume that resource shares $η_{i,h}$ are independent of the household budget. Lise and Seitz (2011), Lewbel and Pendakur (2008), Bargain and Donni (2009, 2012) and DLP all use this restriction in their identification results, and supply some theoretical arguments for it.

[Bargain, Donni, and Hentati (2022)](https://academic.oup.com/jeea/article/20/6/2468/6608951?login=true#389007362): exclusive goods, that is, goods consumed by specific individuals in the household (e.g. adult male and female clothing), and an assumption of preference stability for these goods. Precisely, our identification hinges on the additional information provided by singles’ expenditure on these exclusive goods and on the assumption that adult preferences—for the exclusive goods only—do not depend on the marital or family status. If the presence of a child implies a decrease in the demand for the exclusive good for one of the spouses, then it is concluded that the latter has transferred a fraction of her/his resources to the child and/or that the shadow price of the composite good has increased. The fraction of child costs borne by each parent is retrieved for the first time. This is relatively important information for welfare evaluation, particularly to determine how reforms that channel resources to women—for example, “wallet to purse” reforms as in Lundberg, Pollak, and Wales (1997)—also improve the condition of children. We suggest a tractable approach whereby only a system of two budget share equations–adult male and female clothing–needs to be estimated. Empirical results show that children command a reasonably large share of resources. Resource shares are relatively balanced between spouses when there are no children. However, women contribute a lot more to children’s resources—their contribution is 36%–41% larger than that of men—which makes their share lower than their husband’s in large families.
